### 2017-08-16 doc/MSE.tex
added text and some changes in complex model related to the different perception of the manager

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}

\begin{document}

\title{MSE and dynamic state variable model with sigma function}

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Jan Jaap Poos}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{janjaap.poos@wur.nl}}

\abstract{
TBD

Use DSVM with a tuning parameter that measures how important is to be near optimal.

We use this DSVM to investigate how fishing locations change in response when fish biology is incorporated.}

\date{\today}

\keywords{word1; word2; word3; and word4}

\maketitle

\section{Introduction}
\label{sec1}

The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation to encourage more selective fishing practices. Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other.

Fisheries management typically focuses on the sustainable exploitation of single target species, but the potentially negative effects on, for example, reaching other species conservation objectives, should also be evaluated. Achieving single species MSY in complex and dynamic fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017}. There is no unique way of translating the single-species MSY objective to the multispecies case. Maximisation of yield from one stock will generally require different strategies than maximisation of yield from another.  Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures. Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, on the operation of a mixed fishery and on the fished stocks.

MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}. At the same time being simple enough to build and calibrate reliably with the available data. More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.

There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2017,Batsleer2015, Dowling2011, Poos2010} have been developed. These models generally predict the effect in the short term (within a year) by optimizing an objective function and determine which area best suits this behaviour given the set of incentives that exist (which may also depend on factors such as size, home port, distance to fishing grounds and expected catch rates). Effects of e.g. relocation costs of marine protected areas, the implementation of the landings obligation, have been modelled under these approaches. In contrast, to enable longer term prediction accounting for the impact of these choices on the populations and consequently on the future yields, this fleet dynamic model should be coupled to a biological dynamic model where the feedback between fleet and stock dynamics are explicity modelled.

A key module in any fisheries MSE, and our primary concern here, is the fleet-dynamics model combined with a biological dynamic model, i.e. if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2017}. Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}. The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel. The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.

Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.

%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.

\section{Methods}
\label{sec2}

\subsection{Stock structure and dynamics}
\label{sec2.1}

For purposes of model simplicity, we consider a hypothetical mixed fishery on two species. The fishery is modelled in a individual based dynamic state variable model. In order to allow for fishing effort allocation, the model is seasonally and spatially explicit. The dynamics of the two fish species are governed by a simple age-structured model that is also seasonally and spatially explicit. In the age structured model for the fish species, individuals are assumed to be born at age 0, and in season 1. The number of fish of age $a$ at year $y$, in season $s$ and area $p$ is written as $N (a, y, s, p)$. The number of new-born individuals in the model is independent of the size of the adult population. Individuals can be born in each of the areas, such that  
\begin{equation*}
N (0, y, 1, p) = R (p)
\end{equation*}

Mortality in the model can occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens. However, for simplicity we assume that his natural mortality is negligible. Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species. The decrease in population numbers is thus the result of the catches, which are in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.

\begin{equation*}
C (a, y, s, p) = q(a) * E(y,s,p)
\end{equation*}

and 
\begin{equation*}
N (a, y, s, p) = N (a, y, s, p) - C (a, y, s, p) 
\end{equation*}

Migration for each species is defined by a matrix that defines the migration to and from each of the areas in the model. Its dimensions are thus equal to the number of areas in the model.

the natural mortality at age $a$ in seson $s$ and $F (a, y, s, q)$ is the fishing mortality at age $a$ in year $y$, season $s$ and area $q$. $p (a,y,s,i, q)$ is the proportion of the population available or vulnerable to fishing after migration occurs. Life history traits inputted are: (i) each season less than 1\% of specie 1 and 2 die of natural causes, (ii)  during the first season, specie 1 and 2 spawn in ``A'', (iii) every season, 20\% of specie 1 and 2 migrate from area ``A'' to ``B'' and from ``B'' to ``A'', (iv) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (v) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (vi) at the beginning of each year the fishing recruitment is stablish in 100.

The purpose of the operating model is not to reproduce the entire complexity of a real mixed fishery, but to define plausible hypothesis about population dynamics and then to implement the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.

\subsection{The dynamic-state variable model}
\label{sec2.2}

We develop models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming (Clark and Mangel, 2000; Houston and McNamara, 1999; Poos et al., 2010; Dowling et al., 2011; Batsleer et al., 2015 and Alzorriz et al., 2017).

In order to calculate the optimal state dependent choices during the year, we start by defining the contribution of the annual fines for exceeding landings quotas at the end of the year to the utility is: NEED TO CHECK LANDINGS or CATCH!!
\begin{equation*}
\Phi (L_i, Q_i^L, F_i^L) = -\sum_{i} (max(0, (L_i - Q_i^L)\times F_i^L),
\end{equation*}

where $L_i$ are the cumulative annual landings, for species i. $Q_i^L$ are the annual quota for landings. $F_i^L$ are the fines per unit weight for exceeding landings quota. Hence, starting with $V (L, Q, F, t) = \Phi (L, Q, F)$ we assigned a probability to each region patch proportional to its profit associated at that point in the season (t), given the quota remaining. Following Dowling et al., 2011, if $V^\ast$ is the profit at the optimal location for a given Q and t, we set
\begin{equation*}
\Delta (Q_i^L, t) = V^\ast (L, Q, F, t) - V (L, Q, F, t),
\end{equation*}

and then define the probability of fishing particular area as
\begin{equation*}
P (Q_i^L, t) = \frac{e^-\Delta (Q_i^L, t)/\sigma} {\sum_{i}(e^-\Delta (Q_i^L, t)/\sigma)},
\end{equation*}

where $\sigma$ is a tuning parameter that measures how important is to be near optimal.  A high value of sigma would mean that probabilities are uniform and vessels are distributed uniformly along the different fishing areas, while a low sigma value would mean that vessels concentrate in the optimal location. For computations, we use $\sigma = 40$, to introduce some error in the decision making.

Then, we can iterate backwards in time and fine the optimal choice in terms of location and discarding behaviour for all possible states, combinig the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding anuual quota.

\begin{table}[!h]
\centering
\caption{Fishery parameterization.}
\label{t:fishery}
\resizebox{0.4\textwidth}{!}{%
\begin{tabular}{@{}lc@{}} 
\hline
\textbf{model}                              &   \\ \hline
\\
$\sigma $                                         & 40\\
number of areas & 2 \\
number of seasons                                           & 6\\
\\
\hline
\textbf{Variable costs}                              &\\ \hline
\\
fuel costs (Euro day\textsuperscript{-1})       & 1240\\
gear maintenance (Euro day\textsuperscript{-1}) & 287 \\
crew share                                           & 35\%\\
landing costs (Euro t\textsuperscript{-1})      & 121\\
\\
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}&  \\ \hline
                      \\
sp1                  & 1000 \\
sp2                  & 1000   \\

\\
\hline 
\end{tabular}
}
\end{table}

- Prices of fish have been assumed to be constant over fish age and time.

%We assume high sigma values  to introduce some error in the decision making. A high sigma value is settled to run the model, therefore, we are forcing the model to have uniform probabilities along the different fishing patches. However, with this approach, there is a model time where vessels concentrate in the different locations in the same manner. Therefore, if at that moment we introduce a low sigma value, we force the model to concentrate probabilities in the optimal location.

%If we run each scenario 100 times, we will be able to reproduce the confidence intervals. Providing reference points for management purposes and providing information on the contribution of a fixed number of individuals to the spawning component of the population (spawning stock per recruitment model).

\subsection{Endogenously determined quotas}
\label{sec2.3}

Quota is determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeds Fmsy. This requires that fishing mortality for each stock must not exceed Fmsy, which is a limit reference point. An Optimum Yield (OY) is determined based on a MSY that must prevent overfishing, and that can be reduced from the MSY limit to take into account social, economic and precautionary considerations.  Therefore, quota is determined by a set of harvest control rule parameters consisting of the exploitaion rate at MSY.
related to the OY, which must be risk averse.

\begin{equation*}
 U_{T+1} = U_{MSY}, 
\end{equation*}
\begin{equation*}
 Q_{i, T+1}^{L} = (U_{T+1}\times B_{T+1}),
\end{equation*}

%Given the landing parameters associated

%hr<- catches.n/ (catches.n+ pop)
%landings.ratio <- landings.n/ cathces.n
%OY <- 
%hr_MSY <- OY[OY_landings==max(OY_landingslandings),]_hr

From an economic perspective, a linear increase towards maximum TAC is the optimal strategy under plausible assumptions on market demand for fish and harvesting cost, as shown in Froese et al., 2010. The economic rationale is that, beyond Blim, the stock is within safe biological limits, and as it becomes more and more productive, it is able to supply the market to an increasing extent.

A linear reduction in fishing mortality has been
implemented in Australia (DAFF 2007) a
nd is common in the USA. The differences between
a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in
Appendix S4.

estimating the hr Fisheries are managed by a total allowable catch (TAC). A
maximum TAC is set for each stock so that the respective target biomass is maintained
on average. This maximum TAC may be taken as long as biomass fluctuations remain
above 
B
msy
.

- No quota constrained

- Through a yield per recruit model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.  Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield. The principal purpose of yield per recruit models is to get an idea of the effect of selection pattern and fishing mortality on the yield from a fixed number of individuals that enters the fishery.

Therefore if we have a harvest rate of 0.32, we will need to reduce the harvest in a 68\%. Reduce quota in next run in 68\%. From this reduction in quota and the subsequent reduction in IQ, this scenario will be evaluated allowing and not allowing the fleet to discard.

\subsection{Scenarios analysed}
\label{sec2.4}

Simulations have been performed under four different scenarios. All compared to an unmanaged fishery,  with unconstrained quota for both species.

\begin{enumerate}

\item
Scenario I,  decrease the quota on both species to reach the Fmsy.
Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).

\item 
Scenario II, decrease the quota on sp 1 and unconstrained quota for species 2.
CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.

\item
Scenario III, same as scenario II but allowing to discard only species 2.

\item
Scenario IV, Same as scenario I but allowing to discard sp2.
\end{enumerate}

\section{Results}
\label{sec3}

\subsection{Simulated catch decisions and location choice}
\label{sec3.1}

The largest yield (or catch) that can be taken from a species' stock over an indefinite period. Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely. Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small. At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate. At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals. Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity. At this point, there are no surplus individuals to be harvested and yield drops to zero. The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield

There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983). This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource. 
For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis. Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.

As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing. It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere. Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g. Panayotou 1988).

The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced. This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g. the tuna fishery for the canning industry.

The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition. Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points. Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.

In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points. In evaluating investment projects, including resource management, economists discount the future value of any commodity. Discount rates may be in the order of 10\%. In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested. Long-lived species with slow growth rates, such as whales, clearly fall into this category. The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).

\section{Discussion}
\label{sec4}
tbd
prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative. Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future. This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.

\begin{notes}[Acknowledgements]
The content of this paper does not reflect the official opinion of the European Commission. Responsibility for the information and views expressed in this paper relies entirely with the authors.
\end{notes}

\bibliographystyle{plain}
\bibliography{mylib}

\end{document}

### 2017-08-16 complex model/complexmodel.R

@@ -121,7 +121,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
     season        <- 1:6
     areas         <- c("a", "b")
     stab.model    <- 10
-    NUMRUNS       <- 70
+    NUMRUNS       <- 80
     MPstart       <- 40
     SIMNUMBER     <- 700 #pos
     SIGMA         <- 40 #sig 
@@ -181,7 +181,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
     sp3Price <- sp4Price <- sp5Price <- array(c(0), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
     #---effort and prices used (note that now c is removed (but that if other runs, then make sure to fix/remove code that removes "c" option)                                                                                         
     
-    control     <- DynState.control(spp1LndQuota= 200,  spp2LndQuota=200, spp1LndQuotaFine= 3e6, spp2LndQuotaFine= 3e6, fuelUse = 1, fuelPrice = 50.0, landingCosts= 0,gearMaintenance= 0, addNoFishing= TRUE, increments= 25, spp1DiscardSteps= SPP1DSCSTEPS, spp2DiscardSteps= SPP2DSCSTEPS, sigma= SIGMA, simNumber= SIMNUMBER, numThreads= 20)
+    control     <- DynState.control(spp1LndQuota= 200,  spp2LndQuota=200, spp1LndQuotaFine= 3e6, spp2LndQuotaFine= 3e6, fuelUse = 1, fuelPrice = 150.0, landingCosts= 0,gearMaintenance= 0, addNoFishing= TRUE, increments= 25, spp1DiscardSteps= SPP1DSCSTEPS, spp2DiscardSteps= SPP2DSCSTEPS, sigma= SIGMA, simNumber= SIMNUMBER, numThreads= 20)
     
     #this is where our loop starts, after we set up stable population
     for(yy in (stab.model):(stab.model+NUMRUNS)){
@@ -259,8 +259,13 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
[…]

### 2017-08-16 doc/MSE.tex
added fisures to the text

@@
 \documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
-
+\usepackage{array,multirow,graphicx}
 \begin{document}
@@
 
+\begin{table}[!h]
+\centering
+\caption{}
+\label{scenarios}
+\begin{tabular}{|c |c | c |}
+\hline
+ \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Naive manager}}}} 
+ & &  \\
+ & Discarding allowed & \\
+ & Manager only perceives $L$ &\\
+ & assuming that $C = L$  &\\
+ & YPR based in $F_L$  & \\
+ & assuming $L/C=1$    & No discarding allowed\\
+ & & $C = L$ \\\cline{1-2}
+  \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Smart manager}}}} 
+ & &  Advice and quota\\
+ & Discarding allowed & based on $C = L$\\
+ & Manager perceives $L$ and $D$ & \\
+ & assuming that $C = L + D$ &\\
+ & YPR based in $F_C$  &\\
+ & assuming $L/C \neq 1$ &\\
+  & &  \\\hline
+
+ & \textbf{NO LO}              & \textbf     {LO}                                  \\ \hline
+\end{tabular}
+\end{table}
+
 \begin{enumerate}
@@
 \item
-Scenario I,  decrease the quota on both species to reach the Fmsy.
+Scenario I: \textbf{NO LO + Naive manager},  decrease the quota on both species to reach the Fmsy.
 Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).
@@
 \item 
-Scenario II, decrease the quota on sp 1 and unconstrained quota for species 2.
+Scenario II: \textbf{NO LO + Smart manager}, decrease the quota on sp 1 and unconstrained quota for species 2.
 CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.
@@
 \item
-Scenario III, same as scenario II but allowing to discard only species 2.
+Scenario III: \textbf{LO}, same as scenario II but allowing to discard only species 2.
 
-\item
-Scenario IV, Same as scenario I but allowing to discard sp2.
+%\item
+%Scenario IV, Same as scenario I but allowing to discard sp2.
 \end{enumerate}
@@
 
+Scenario III:
+%----------------------------------------------------------------------------------------------------------------------------------------
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Catch_LO.eps} 
+\caption{}
+\label{fig:catch_lo}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Effort_LO.eps} 
+\caption{}
+\label{fig:effort_lo}
+\end{figure}
+%----------------------------------------------------------------------------------------------------------------------------------------
+
 \section{Discussion}

### 2017-10-03 doc/MSE.tex
changes in text and scenarios

--- 2017-08-16 doc/MSE.tex
@@
 \documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
+\usepackage{array,multirow,graphicx}
+\usepackage{pdflscape}% for landscape
 
@@
 
-The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation to encourage more selective fishing practices.
+The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices.
 Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other.
+LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation.
 
@@
 Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
+If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
+Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.
 
@@
 Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species.
-The decrease in population numbers is thus the result of the catches, which are in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
+The decrease in population numbers is thus the result of the catches ($C (a, y, s, p)$), which are in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
 
@@
 
-Migration for each species is defined by a matrix that defines the migration to and from each of the areas in the model. Its dimensions are thus equal to the number of areas in the model.
+Migration for each species is defined by a matrix that defines the migration to ($top$) and from ($fromp$) each of the areas in the model.
+Its dimensions are thus equal to the number of areas in the model.
+The migration in population numbers ($Mig (a, 1, s, top, fromp)$) is thus the proportion ($p (a, y, s, i, fromq, top)$) of the population available or vulnerable to fishing after migration occurs.
+\begin{equation*}
+Mig (a, 1, s, top, fromp) = Mig(a, 1, s, top, fromp) + N (a, y, s, fromp) * p (a, y, s, i, fromq, top)
+\end{equation*}
 
-the natural mortality at age $a$ in seson $s$ and $F (a, y, s, q)$ is the fishing mortality at age $a$ in year $y$, season $s$ and area $q$.
-$p (a,y,s,i, q)$ is the proportion of the population available or vulnerable to fishing after migration occurs.
-Life history traits inputted are: (i) each season less than 1\% of specie 1 and 2 die of natural causes, (ii)  during the first season, specie 1 and 2 spawn in ``A'', (iii) every season, 20\% of specie 1 and 2 migrate from area ``A'' to ``B'' and from ``B'' to ``A'', (iv) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (v) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (vi) at the beginning of each year the fishing recruitment is stablish in 100.
+and then population numbers
+\begin{equation*}
+N (a, y, s, top) = N (a, y, s, top) +  Mig (a, 1, s, top, fromp)
+\end{equation*}
+
+Life history traits inputted are: (i)  during the first season, specie 1 spawns in ``A'', while specie 2 spawns in ``B'', (ii) every season, 20\% of specie 1 and 2 migrate from area ``A'' to ``B'' and from ``B'' to ``A'', (iii) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (iv) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (v) at the beginning of each year the fishing recruitment is stablish in 100.
 
@@
 
-We develop models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming (Clark and Mangel, 2000; Houston and McNamara, 1999; Poos et al., 2010; Dowling et al., 2011; Batsleer et al., 2015 and Alzorriz et al., 2017).
+We develop models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2017, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
 
-In order to calculate the optimal state dependent choices during the year, we start by defining the contribution of the annual fines for exceeding landings quotas at the end of the year to the utility is: NEED TO CHECK LANDINGS or CATCH!!
+In order to calculate the optimal state dependent choices during the year, we start by defining the contribution of the annual fines for exceeding landings or catches quotas at the end of the year to the utility is:
 \begin{equation*}
-\Phi (L_i, Q_i^L, F_i^L) = -\sum_{i} (max(0, (L_i - Q_i^L)\times F_i^L),
+\Phi (C_i, Q_i^C, F_i^C) = -\sum_{i} (max(0, (C_i - Q_i^C)\times F_i^C),
 \end{equation*}
 
-where $L_i$ are the cumulative annual landings, for species i.
-$Q_i^L$ are the annual quota for landings.
-$F_i^L$ are the fines per unit weight for exceeding landings quota.
-Hence, starting with $V (L, Q, F, t) = \Phi (L, Q, F)$ we assigned a probability to each region patch proportional to its profit associated at that point in the season (t), given the quota remaining.
-Following Dowling et al., 2011, if $V^\ast$ is the profit at the optimal location for a given Q and t, we set
+where $C_i$ are the cumulative annual catches, for species i.
+$Q_i^C$ are the annual quota for catches.
+$F_i^C$ are the fines per unit weight for exceeding catch quota.
+Hence, starting with $V (C, Q, F, t) = \Phi (C, Q, F)$ we assigned a probability to each region patch proportional to its profit associated at that point in the season (t), given the quota remaining.
+Following \cite{Dowling2011}, if $V^\ast$ is the profit at the optimal location for a given Q and t, we set
 \begin{equation*}
-\Delta (Q_i^L, t) = V^\ast (L, Q, F, t) - V (L, Q, F, t),
+\Delta (Q_i^C, t) = V^\ast (C_i, Q_i^C, F_i^C, t) - V (C_i, Q_i^C, F_i^C, t),
 \end{equation*}
@@
 \begin{equation*}
-P (Q_i^L, t) = \frac{e^-\Delta (Q_i^L, t)/\sigma} {\sum_{i}(e^-\Delta (Q_i^L, t)/\sigma)},
+P (Q_i^C, t) = \frac{e^-\Delta (Q_i^C, t)/\sigma} {\sum_{i}(e^-\Delta (Q_i^C, t)/\sigma)},
 \end{equation*}
@@
 A high value of sigma would mean that probabilities are uniform and vessels are distributed uniformly along the different fishing areas, while a low sigma value would mean that vessels concentrate in the optimal location.
-For computations, we use $\sigma = 40$, to introduce some error in the decision making.
+For computations, we use $\sigma = 40$, to introduce some error in the decision making (Table \ref{t:fishery}).
 
 Then, we can iterate backwards in time and fine the optimal choice in terms of location and discarding behaviour for all possible states, combinig the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding anuual quota.
+Prices of fish have been assumed to be constant over fish age and time.
+The variable costs include fuel costs ($\sim{40\%}$), crew share ($\sim{35\%}$) and gear mantainance ($\sim{4\%}$)).
+Fuel costs depend on gear used, trip effort and fuel price.
+Fuel costs were set to 150 \texteuro day\textsuperscript{-1}.
+Gear maintenance costs were assumed proportional to fishing effort, landing costs proportional to the total weight landed, and crew share variable costs proportional to the income from catches (Table~\ref{t:fishery}).
 
@@
 \label{t:fishery}
-\resizebox{0.4\textwidth}{!}{%
-\begin{tabular}{@{}lc@{}} 
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{@{}|lc|lc|lc|@{}} 
 \hline
-\textbf{model}                              &   \\ \hline
-\\
-$\sigma $                                         & 40\\
-number of areas & 2 \\
-number of seasons                                           & 6\\
-\\
-\hline
-\textbf{Variable costs}                              &\\ \hline
-\\
-fuel costs (Euro day\textsuperscript{-1})       & 1240\\
-gear maintenance (Euro day\textsuperscript{-1}) & 287 \\
-crew share                                           & 35\%\\
-landing costs (Euro t\textsuperscript{-1})      & 121\\
-\\
-\hline
-\textbf{Market value (Euro kg\textsuperscript{-1})}&  \\ \hline
-                      \\
-sp1                  & 1000 \\
-sp2                  & 1000   \\
-
-\\
-\hline 
+\textbf{model}                              &               & \textbf{Variable costs}                        &      &\textbf{Market value}&  \\ \hline
+& & & & & \\
+number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  & sp1 (Euro kg\textsuperscript{-1})                 & 1000 \\ 
+$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 287  & sp2 (Euro kg\textsuperscript{-1})                 & 1000   \\
+number of areas                             & 2             &crew share                                      & 35\% & & \\
+number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & &\\
+intial quota $Q_sp1$ and $Q_sp2$            & 200           &                                                &      & & \\
+& & & & & \\\hline
 \end{tabular}
 }
+%\hline
+%\textbf{model}                              &   \\ \hline
+%\\
+%number of vessels                                 &700\\
+%$\sigma $                                         & 40\\
+%number of areas                                   & 2 \\
+%number of seasons                                 & 6\\
+%intial quota Q1 and Q2                            & 200\\
+%\\\hline
+%\textbf{Variable costs}                              &\\ \hline
+%\\
+%fuel costs (Euro day\textsuperscript{-1})       & 1240\\
+%gear maintenance (Euro day\textsuperscript{-1}) & 287 \\
+%crew share                                      & 35\%\\
+%landing costs (Euro t\textsuperscript{-1})      & 121\\
+%\\\hline
+%\textbf{Market value (Euro kg\textsuperscript{-1})}&  \\ \hline
+%\\
+%sp1                  & 1000 \\
+%sp2                  & 1000   \\
+%\\\hline 
+%\end{tabular}
+%}
 \end{table}
 
-- Prices of fish have been assumed to be constant over fish age and time.
+% Prices of fish have been assumed to be constant over fish age and time.
 
@@
 
+prel.quota <-  sum(sweep((hr1wanted[1]/mean(hr1[,yy,]))* hr1[,yy,]*landings.ratio1[,yy,]*apply(pop1[,yy+1,,],c(1,2), sum) ,1,wts,"*"))/SIMNUMBER
+        tac.constrained <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
+        quota1[,yy+1,,] <- max(min(prel.quota, tac.constrained[2]), tac.constrained[1])
+
 From an economic perspective, a linear increase towards maximum TAC is the optimal strategy under plausible assumptions on market demand for fish and harvesting cost, as shown in Froese et al., 2010.
@@
 
-A linear reduction in fishing mortality has been
-implemented in Australia (DAFF 2007) a
-nd is common in the USA. The differences between
-a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in
-Appendix S4.
+A linear reduction in fishing mortality has been implemented in Australia (DAFF 2007) a nd is common in the USA.
+The differences between a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in Appendix S4.
 
-estimating the hr Fisheries are managed by a total allowable catch (TAC). A
-maximum TAC is set for each stock so that the respective target biomass is maintained
-on average. This maximum TAC may be taken as long as biomass fluctuations remain
-above 
-B
-msy
-.
+estimating the hr Fisheries are managed by a total allowable catch (TAC). A maximum TAC is set for each stock so that the respective target biomass is maintained
+on average. This maximum TAC may be taken as long as biomass fluctuations remain above Bmsy.
 
@@
 
-Simulations have been performed under four different scenarios. All compared to an unmanaged fishery,  with unconstrained quota for both species.
+Simulations have been performed under three different scenarios. All compared to an unmanaged fishery, with unconstrained quota for both species.
+
+\begin{table}[!h]
+\centering
+\caption{}
+\label{scenarios}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{|c |c | c | c |}
+\hline
+ \textbf{Scenario I}            & \textbf{Scenario II}                & \textbf{Scenario III}               & \textbf{ScenarioIV}           \\ 
+ \textbf{Fulll avoidance of discards} & \textbf{Landings selectivity} & \textbf{Catch selectivity}          & \textbf{LO not fully enforced} \\ \hline
+ & &  &\\
+ No discarding allowed          & Discarding allowed                  & Discarding allowed                  & Discarding ocurred but not perceived \\
+ Manager perceives $C$          & Manager perceives $L$ and $D$       & Manager perceives $L$ and $D$       & Manager only perceives $L$ \\
+ assuming that $C = L$          & assuming that $C = L + D$           & assuming that $C = L + D$           & assuming that $C = L$  \\
+ YPR based in $F_C$             & YPR based in $F_L$                  & YPR based in $F_C$                  &  YPR based in $F_L$  \\
+ assuming $L/C=1$               & assuming $L/C \neq 1$               & assuming $L/C \neq 1$               & assuming $L/C=1$\\
+  & &  &\\\hline
+\end{tabular}
+}
+\end{table}
+
+% \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Naive manager}}}} 
+% & &  \\
+ %& Discarding allowed & \\
+% & Manager only perceives $L$ &\\
+% & assuming that $C = L$  &\\
+% & YPR based in $F_L$  & \\
+% & assuming $L/C=1$    & No discarding allowed\\
+% & & $C = L$ \\\cline{1-2}
+%  \parbox[c]{2mm}{\multirow{7}{*}{\rotatebox[origin=c]{90}{\textbf{Smart manager}}}} 
+% & &  Advice and quota\\
+% & Discarding allowed & based on $C = L$\\
+% & Manager perceives $L$ and $D$ & \\
+% & assuming that $C = L + D$ &\\
+% & YPR based in $F_C$  &\\
+% & assuming $L/C \neq 1$ &\\
+%  & &  \\\hline
+% & \textbf{NO LO}              & \textbf     {LO}                                  \\ \hline
 
@@
 \item
-Scenario I,  decrease the quota on both species to reach the Fmsy.
-Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).
+\textbf{Scenario I: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
+%decrease the quota on both species to reach the Fmsy.
+%Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).
 
 \item
-Scenario II, decrease the quota on sp 1 and unconstrained quota for species 2.
-CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.
+\textbf{Scenario II: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
+Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
 
 \item
-Scenario III, same as scenario II but allowing to discard only species 2.
+\textbf{Scenario III: Catch selectivity},  under this scenario the fishery is under a full catch selectivity, but all discards are landed and contribute to yield.
+Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
+This results generally in higher $F_{MSY}$.
+%decrease the quota on both species to reach the Fmsy.
+%Reduce the IQs in the same percentage as the change in HR to reach the equilibrium (MSY).
 
 \item 
-Scenario IV, Same as scenario I but allowing to discard sp2.
+\textbf{Scenario IV: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding.
+However the manager could think that is fully enforced and assumed that only what is landed contribute to yield.
+This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality.
+Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.
+en vez de plantearlo como que no es LO, podria plantearse como que realmente esta bajo el LO, pero no hay compliance y el manager no lo sabe. Por tanto los hr se estiman como si el C fueran L...
+%decrease the quota on sp 1 and unconstrained quota for species 2.
+%CONSTRAINED fishery for sp1 and UNCONSTRAINED sp2 (FIXING QUOTA SP2 to 200). Only sp1 is forced to try to reach the Fmsy. However discarding is allowed for both species.
+
+%same as scenario II but allowing to discard only species 2.
+
+%\item
+%Scenario IV, Same as scenario I but allowing to discard sp2.
 \end{enumerate}
@@
 \label{sec3}
+
+\newpage
+\begin{landscape}
+Scenario I:
+%----------------------------------------------------------------------------------------------------------------------------------------
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO.eps} 
+\caption{}
+\label{fig:catch_lo}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO.eps} 
+\caption{}
+\label{fig:effort_lo}
+\end{figure}
+\end{landscape}
+
+%----------------------------------------------------------------------------------------------------------------------------------------
+\newpage
+\begin{landscape}
+Scenario II:
+%----------------------------------------------------------------------------------------------------------------------------------------
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_disc_YieldLand.eps} 
+\caption{}
+\label{fig:catch_lo}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_disc_YieldLand.eps} 
+\caption{}
+\label{fig:effort_lo}
+\end{figure}
+\end{landscape}
+
+%----------------------------------------------------------------------------------------------------------------------------------------
+\newpage
+\begin{landscape}
+Scenario III:
+%----------------------------------------------------------------------------------------------------------------------------------------
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_smartmanager.eps} 
+\caption{}
+\label{fig:catch_lo}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_smartmanager.eps} 
+\caption{}
+\label{fig:effort_lo}
+\end{figure}
+\end{landscape}
+
+%----------------------------------------------------------------------------------------------------------------------------------------
+\newpage
+\begin{landscape}
+Scenario IV:
+%----------------------------------------------------------------------------------------------------------------------------------------
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Catch_LO_nocompliance.eps}  
+\caption{}
+\label{fig:catch_lo}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textheight,height=6cm]{Figures/Effort_LO_nocompliance.eps} 
+\caption{}
+\label{fig:effort_lo}
+\end{figure}
+\end{landscape}
+%----------------------------------------------------------------------------------------------------------------------------------------
 
@@
 
+\newpage
 \bibliographystyle{plain}

### 2017-12-06 doc/MSE_OCT.tex
last model version

\documentclass[12pt,oneline,a4paper,numbib]{ouparticle}
\usepackage{array,multirow,graphicx}
\usepackage{pdflscape}% for landscape

\usepackage{adjustbox}     % adjusting table (too wide)
\usepackage{rotating}      % sidewaytables
\usepackage{array}
\usepackage{stackengine}

\usepackage{amsmath} % \numberwithin{equation} doesn't exist without this package.
\numberwithin{equation}{subsection} % This line resets equation numbering when starting a new section.
\renewcommand{\theequation}{Eq. \thesubsection.\arabic{equation}} % This line ads "Eq." in front of your equation numbering.

\begin{document}

\title{MSE and dynamic state variable model with sigma function}

\author{%
\name{Nekane Alzorriz}
\address{European Commission, Joint Research Centre (JRC), Sustainable Resources Directorate, Water and Marine Resources Unit, Via Enrico Fermi 2749, 21027 Ispra, Italy.}
\email{nekane.alzorriz@gmail.com}
\and
\name{Jan Jaap Poos}
\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
\email{janjaap.poos@wur.nl}}

\abstract{
TBD
Farcas 2016:The case of fisheries management illustrates how the inherent structural instability of ecosystems can have deep-running policy implications. We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses complex food webs in its operating model. 
Use DSVM with a tuning parameter that measures how important is to be near optimal.
We use this DSVM to investigate how fishing locations change in response when fish biology is incorporated.}

\date{\today}

\keywords{word1; word2; word3; and word4}

\maketitle

\section{Introduction}
\label{sec1}

The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices. Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other. LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation.

Fisheries management typically focuses on the sustainable exploitation of single target species, but the potentially negative effects on, for example, reaching other species conservation objectives, should also be evaluated. Achieving single species MSY in complex and dynamic fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017}. There is no unique way of translating the single-species MSY objective to the multispecies case. Maximisation of yield from one stock will generally require different strategies than maximisation of yield from another.  Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures. Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty.

MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}. At the same time being simple enough to build and calibrate reliably with the available data. More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.

There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010} have been developed. These models generally predict the effect in the short term (within a year) by optimizing an objective function and determine which area best suits this behaviour given the set of incentives that exist (which may also depend on factors such as size, home port, distance to fishing grounds and expected catch rates). Effects of e.g. relocation costs of marine protected areas, the implementation of the landings obligation, have been modelled under these approaches. In contrast, to enable longer term prediction accounting for the impact of these choices on the populations and consequently on the future yields, this fleet dynamic model should be coupled to a biological dynamic model where the feedback between fleet and stock dynamics are explicity modelled.

We review the potential of MSE, emphasazing the role of individual harvester decision-making and socioeconomic drivers on management effectiveness, i.e. if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}. Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}. The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel. The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.

Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.

\section{Methods}
\label{sec2}
The MSE approach is based upon a set of model of the "true" population dynamics of the species (called "operating model"; section \ref{sec2.1}). The operating model aims capturing the key processes in the dynamics of the fish population given the best knowledge available, and can be thought of as a minimum realistic model \cite{Punt1995}. The next step in the MSE is to simulate the process of monitoring the stock (called "observation model"; section \ref{sec2.2}). This resulted in simulated measurements such as individual harvester decision-making and the consequent biomass (including error). The observation data are then passed to the "management model" (section \ref{sec2.3}). The management model encompasses the harvest control rules (HCR). The HCR references to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level, changes in gear or spatial and temporal reallocations or restrictions of fishing effort. Management actions are rarely implemented without error. This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g. when and where harvest is taken) are not accounted for in the HCR. By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.

\subsection{Stock structure and dynamics}
\label{sec2.1}

For purposes of model simplicity, we considered a hypothetical mixed fishery on two species. The fishery was modelled in a individual based dynamic state variable model. In order to allowed for fishing effort allocation, the model was seasonally and spatially explicit. The dynamics of the two fish species were governed by a simple age-structured model that is also seasonally and spatially explicit. In the age structured model for the fish species, individuals were assumed to be born at age 0, and in season 1. The number of fish of age $a$ at year $y$, in season $s$ and area $p$ was written as $N (a, y, s, p)$. The number of new-born individuals in the model was independent of the size of the adult population. Individuals could be born in each of the areas, such that  
\begin{equation}
N (0, y, 1, p) = R (p)
\end{equation}

Mortality in the model could occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens. However, for simplicity we assumed that his natural mortality was negligible. Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species. The decrease in population numbers was thus the result of the catches ($C (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area. 
\begin{equation}
C (a, y, s, p) = q(a) * E(y,s,p),
\end{equation}
and 
\begin{equation}
N (a, y, s, p) = N (a, y, s, p) - C (a, y, s, p). 
\end{equation}

It was assumed that fishing mortality ocurred before migration ocurred. Migration for each species was defined by a matrix $Mig$ that defines the impact of migration (.e. immigration-emigration) on a given stock was also dependent on the relative sizes of the stocks. The size of that matrix  was defined by the number of age classes, seasons and number of immigrants into area $p$ ($im$) and emigrants leaving area $p$ ($em$) in the model. The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($p (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs.   
\begin{equation}
Mig (a, 1, s, im, em) = Mig(a, 1, s, im, em) + N (a, y, s, em) * p (a, y, s, i, em, im),
\end{equation}
and then population numbers
\begin{equation}
N (a, y, s, im) = N (a, y, s, im) +  Mig (a, 1, s, im, em).
\end{equation}

Life history traits inputted were: (i)  during the first season, specie 1 spawned in $A$, while specie 2 spawned in $B$, (ii) every season, 20\% of specie 1 and 2 emigrated and immigrated between areas, (iii) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (iv) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (v) at the beginning of each year the fishing recruitment was stablish in 100.

The purpose of the operating model was not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.

\subsection{The dynamic-state variable model}
\label{sec2.2}

To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}. A dynamic state variable model \cite{Houston1999, ClarkandMangel2000} was used to model location choice, extending the model structure in \cite{Batsleer2015}. In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations. The model had annual fines for exceeding landings or catch quota as in \cite{Alzorriz2018}). In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings or catch quotas at the end of the year:
\begin{equation}
\Phi (C_i, Q_i, F_i)= -\sum_i (max( 0, (C_i - Q_i))* F_i),
\end{equation}
where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel. These cumulative landings defined the state of the individual. $Q_i$ was the annual individual quota for landings or catches for the quota species. Individual quotas were not transferable. In order to reduce computation time, we set quota only for two species. $F_i$ was the fine per unit weight for exceeding individual landings or catch quotas was set to $3 x 10^3$ Euro day\textsuperscript{-1}. These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.

The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, s)= \Phi (C_i, Q_i, F_i)$. For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$ (A, B and including area 0: “staying in port”), and to keep or discard any combination of the age classes caught of the quota species. This discarding was defined by a matrix $d$. The size of that matrix was defined by the number of species under quota constraints and the number of age classes. Each element could take the value 0 (discard) or 1 (keep on board and land). The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
\begin{equation}
V (C_i, Q_i, F_i, s) = max_{p,d}( R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1])),
\end{equation}
where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices). Prices of fish were assumed to be dependant over fish age and time (subsection \ref{sec2.2.1}). $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table ~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices. The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$. The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.

Rather than assuming that each individual always made the optimal choice, we assigned a probability to each choice proportional to its expected utility, following \cite{Dowling2011}. The expected utility for any choice was
\begin{equation}
U (C_i, Q_i, F_i, s) = R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]).
\end{equation}

If $U^*$ was the expected utility at the optimal choice for a given t, we set
\begin{equation}
\Delta_{p, d}(C_i, Q_i, F_i, s) =  U^* (C_i, Q_i, F_i, s) - U (C_i, Q_i, F_i, s),
\end{equation}
and then defined the probability of a choice for a given area and discarding as	
\begin{equation} \label{eq:solve}
P_{p, d}(C_i, Q_i, F_i, s) = \frac
                % Nominator
                {e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}}
                % Denominator
                {\sum_p \sum_d e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}},
\end{equation}
where $\sigma$ was a tuning parameter that measures how important was to be near optimal. A high value of sigma resulted in uniform probabilities of choices, and that vessels were distributed uniformly among the different fishing areas. In contrast, if sigma was small then all vessels concentrated in the optimal location (but note that $\sigma$ should be $> 0$). For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 1.1 x 10^{11}$, but was generally of the order of $1.1 x 10^5$ in magnitude).

The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location and discarding behaviour for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota. Further details for this procedure can be found in \cite{Alzorriz2018, Batsleer2016} and \cite{Dowling2011}.

Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in ~\ref{eq:solve}. This was done for 700 individual vessels.

\begin{table}[!h]
\centering
\caption{Fishery parameterization.}
\label{t:fishery}
\resizebox{0.7\textwidth}{!}{%
\begin{tabular}{@{}lc lc @{}} 
\hline
\textbf{model}                              &               & \textbf{Variable costs}                        &      \\\hline%&\textbf{Market value}&  \\ \hline
& & & \\%& & \\
number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  \\%& sp1 (Euro kg\textsuperscript{-1})                 & 1000 \\ 
$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0   \\% & sp2 (Euro kg\textsuperscript{-1})                 & 1000   \\
number of areas                             & 2             &crew share                                      & 35\% \\%& & \\
number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    \\%& &\\
intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
& & & \\\hline%& & \\\hline
\end{tabular}
}
\end{table}

\subsubsection{Size dependant pricing}
\label{sec2.2.1}
% Prices of fish have been assumed to be constant over fish age and time.
Common knowledge and case studies indicated that size and weight of a fish could be important attributes when determining its economic value. However, so far there has been no broader assessment whether the size-dependence of price really is prevalent and significant in fisheries. The results from \cite{Zimmermann2011, Zimmermann2013} underline that size-dependent pricing is common in Norwegian fisheries, and is therefore also of high relevance for decisions in fisheries management. Therefore, the resulting data of mean weights, average prices and total yield per weight classes and year were used for further analysis. We fitted linear mixed models (Bates and Maechler, 2010) and general additive models (Wood, 2010) in R (R: Development Core Team, 2012), using mean price and mean body weight as the response variable and covariate, respectively. Thus, the statistical models took the form  
\begin{equation}
 p_{i} \sim c_{0} + c_{1} w_{i} + f (y_{i}), 
\end{equation}
where p is price of observation i, wi is the mean weight of the corresponding weight class, and the term f(yi) corresponds to the effect of year, treated either as a random effect or as a smooth term. Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models. Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results. All models were fitted independently to each stock.
\begin{table}[!h]
\centering
\caption{Size dependant pricing.}
\label{t:pricing}
\resizebox{0.5\textwidth}{!}{%
\begin{tabular}{@{}lcccccc@{}} 
\hline
\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{6}{c}{season} \\ 
age                      & 1     & 2     & 3    & 4    & 5    & 6    \\ \hline
                      \\
1                       &    &    &   &   & & \\
2                       &    &    &   &   & &\\
3                       &    &    &   &   & &\\
4                       &    &    &   &   & &\\
\\
\hline 
\end{tabular}
}
\end{table}

\subsection{Endogenously determined quotas}
\label{sec2.3}

The three management strategies we considered differ by the formulae used to determine exploitation rates (HCR). These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan. For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.

\begin{table}[!h]
\centering
\caption{}
\label{scenarios}
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c  c }
\hline
                                & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
                                & \textit{(landings selectivity)}  & \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
& & &  \\
Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
Manager catch perception        & $C = L + D$                    &        $C = L$                       &  $C = L$  \\
Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\
Selectivity ($L/C$)             &  $ \neq 1$                     & $=1$                                 & $=1$\\
&  & &  \\\hline
\end{tabular}
}
\end{table}

\begin{enumerate}

\item
\textbf{Scenario I: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield. Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.

\item
\textbf{Scenario II: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.

\item 
\textbf{Scenario III: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding. However the manager could think that is fully enforced and assumed that only what is landed contribute to yield. This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality. Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.

\end{enumerate}

HCRs in the form of Quotas were determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeded Fmsy. This required that fishing mortality for each stock must not exceed Fmsy, which was a limit reference point. An Optimum Yield (OY) was determined based on a MSY that must prevented overfishing, and that could be reduced from the MSY limit to take into account social, economic and precautionary considerations.  Therefore, quota was determined by a set of harvest control rule parameters consisting of the exploitaion rate at MSY.

The resulted biomass harvested rate in a year $y$, depended in the manager catch perception and the standing biomass:
\begin{equation}
 H_{a, y, s} =\frac
                % Nominator
                {C_i numbers(a, y, s, p)}
                % Denominator
                {C_i numbers(a, y, s, p) \times B_i numbers(a, y, s, p)},
\end{equation}

\begin{equation}
 lratio (a, y, s) = \frac
                % Nominator
                {L_i numbers(a, y, s, p)}
                % Denominator
                {C_i numbers(a, y, s, p)},
\end{equation}

\begin{equation}
 H_{MSY} = YIELD CURVE..., 
\end{equation}

\begin{equation}
 Q_{MSY} = \frac
                % Nominator
                {\sum( [\frac{H_{MSY}}{\bar{H}} \times H (a, s) \times lratio (a, s) \times B numbers(a, s)] \times wts) }
                % Denominator
                {number of vessels}, 
\end{equation}
\begin{equation}
 Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85)), 
\end{equation}

%Given the landing parameters associated

%hr<- catches.n/ (catches.n+ pop)
%landings.ratio <- landings.n/ cathces.n
%OY <- 
%hr_MSY <- OY[OY_landings==max(OY_landingslandings),]_hr

%prel.quota <-  sum(sweep((hr1wanted[1]/mean(hr1[,yy,]))* hr1[,yy,]*landings.ratio1[,yy,]*apply(pop1[,yy+1,,],c(1,2), sum) ,1,wts,"*"))/SIMNUMBER
%        tac.constrained <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
%        quota1[,yy+1,,] <- max(min(prel.quota, tac.constrained[2]), tac.constrained[1])

%From an economic perspective, a linear increase towards maximum TAC is the optimal strategy under plausible assumptions on market demand for fish and harvesting cost, as shown in Froese et al., 2010. The economic rationale is that, beyond Blim, the stock is within safe biological limits, and as it becomes more and more productive, it is able to supply the market to an increasing extent.

%A linear reduction in fishing mortality has been implemented in Australia (DAFF 2007) a nd is common in the USA. The differences between a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in Appendix S4.

%estimating the hr Fisheries are managed by a total allowable catch (TAC). A maximum TAC is set for each stock so that the respective target biomass is maintained
%on average. This maximum TAC may be taken as long as biomass fluctuations remain above Bmsy.

%- No quota constrained

%- Through a yield per recruit model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.  Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield. The principal purpose of yield per recruit models is to get an idea of the effect of selection pattern and fishing mortality on the yield from a fixed number of individuals that enters the fishery.

%Therefore if we have a harvest rate of 0.32, we will need to reduce the harvest in a 68\%. Reduce quota in next run in 68\%. From this reduction in quota and the subsequent reduction in IQ, this scenario will be evaluated allowing and not allowing the fleet to discard.

\section{Results}
\label{sec3}

\newpage
%\begin{landscape}
Scenario I:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand.eps} 
\end{minipage}
\caption{}
\end{figure}
%\end{landscape}

%----------------------------------------------------------------------------------------------------------------------------------------

Scenario II:
%----------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO.eps} 
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO.eps} 
\end{minipage}
\caption{}
\end{figure}

%----------------------------------------------------------------------------------------------------------------------------------------

Scenario III:

\begin{figure}[!h]
\centering
\begin{minipage}{.45\textwidth}
\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_nocompliance.eps}  
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\centering
\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_nocompliance.eps} 
\end{minipage}
\caption{}
\label{fig:effort_lo}
\end{figure}

%----------------------------------------------------------------------------------------------------------------------------------------

\subsection{Simulated catch decisions and location choice}
\label{sec3.1}

The largest yield (or catch) that can be taken from a species' stock over an indefinite period. Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely. Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small. At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate. At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals. Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity. At this point, there are no surplus individuals to be harvested and yield drops to zero. The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield

There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983). This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource. 
For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis. Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.

As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing. It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere. Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g. Panayotou 1988).

The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced. This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g. the tuna fishery for the canning industry.

The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition. Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points. Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.

In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points. In evaluating investment projects, including resource management, economists discount the future value of any commodity. Discount rates may be in the order of 10\%. In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested. Long-lived species with slow growth rates, such as whales, clearly fall into this category. The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).

\newpage
\section{Discussion}
\label{sec4}
tbd
prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative. Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future. This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.

\begin{notes}[Acknowledgements]
The content of this paper does not reflect the official opinion of the European Commission. Responsibility for the information and views expressed in this paper relies entirely with the authors.
\end{notes}

\newpage
\bibliographystyle{plain}
\bibliography{mylib}

\end{document}

### 2017-12-06 RDynState5NAsigmaseason6_age/src/RDynState.cpp

@@ -518,10 +518,11 @@ Rprintf("Start of DynStateF\n");
   PITYPE thePriceParms      = (PITYPE) matalloc(sizeof(float), (void *)0, 3, MAXHORIZON, NOSPEC, NOSIZES);
     
   int theEffortArray[320000][MAXHORIZON];
-  double *theIncrementArray = (double *) malloc((size_t)NOSPEC * sizeof (double));
-  double *theShortTermGains = (double *) malloc((size_t)kNPatch * sizeof (double));
-  double *theShortTermCosts = (double *) malloc((size_t)kNPatch * sizeof (double));
-  double *theShortTermEcon  = (double *) malloc((size_t)kNPatch * sizeof (double));
+  double *theIncrementArray     = (double *) malloc((size_t)NOSPEC * sizeof (double));
+  double *theShortTermGains     = (double *) malloc((size_t)kNPatch * sizeof (double));
+  double *theShortTermCosts     = (double *) malloc((size_t)kNPatch * sizeof (double));
+  double *theShortTermCrewShare = (double *) malloc((size_t)kNPatch * sizeof (double));
+  double *theShortTermEcon      = (double *) malloc((size_t)kNPatch * sizeof (double));
 
   if (theLndParms == 0l) Rprintf("error in memory allocation theLndParms \n");
   if (theDisParms == 0l) Rprintf("error in memory allocation theDisParms \n");
@@ -739,7 +740,8 @@ Rprintf("Start of DynStateF\n");
     for (int i = 0; i < kNPatch; i++){     
       theShortTermGains[i] = shortTermGains(t, noInc, theLndParms, i, thePriceParms);
       theShortTermCosts[i] = shortTermCosts(t, i, kPriceEffort, theEffortArray);     //*, noInc, theLndParms, thePriceParms);
-      theShortTermEcon[i]  =  theShortTermGains[i] - theShortTermCosts[i];
+      theShortTermCrewShare[i] = theShortTermGains[i]*0.35;     //*, noInc, theLndParms, thePriceParms);
+      theShortTermEcon[i]  =  theShortTermGains[i] - theShortTermCosts[i] - theShortTermCrewShare[i];
     }
 
     maxspp[t]  = accumulate(whatRangeLT,whatRangeLT+t,0) +2;                         // calc max number of increments to loop over for this timest

### 2017-12-06 complex model/complexmodel.R

@@ -154,14 +154,14 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
     effort <- array(c(1), dim=c(length(areas), length(season)), dimnames=list(option =areas,season=as.character(season)))
     
     
-    pop1  <- pop2   <-array(0, dim=c(length(ages),endy+1,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:(endy+1)), season=as.character(season), option =areas))
-    catches.n.dsvm1      <- catches.n.dsvm2      <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
+    pop1  <- pop2         <-array(0, dim=c(length(ages),endy+1,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:(endy+1)), season=as.character(season), option =areas))
+    catches.n.dsvm1       <- catches.n.dsvm2       <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
     landings.n.dsvm1      <- landings.n.dsvm2      <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
-    catches.wt.dsvm1     <- catches.wt.dsvm2     <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
+    catches.wt.dsvm1      <- catches.wt.dsvm2      <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
[…]

### 2017-12-06 complex model/complexmodel_OCT.R

@@ -1,102 +1,122 @@
+options(width=200)
+library(RDynState5NAsigmaseason6Age)
+library(ggplot2)
+library(reshape2)
+library(plyr)
 library(scales)
+
 setwd("~/Dropbox/BoB/MSE/Git/Nekane/complex model/")
 source("functions.R")
+
+
 ##############################################################################
 # POPULATION DYNAMICS
 ##############################################################################
+
 population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
-#pop[age,year, season, area]
-MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
-for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))
-for (ss in (1:length(season))){
-# move time ---------------------
-if (ss ==1){
-for(age in 2:(dim(pop)[1])){
-pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
+  #pop[age,year, season, area]                                                                                                        
+  MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
+  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
+    for (ss in (1:length(season))){
+      # move time ---------------------                                                             
+      if (ss ==1){
+        for(age in 2:(dim(pop)[1])){
+          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
+        }
+      } else {
+        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
+      }
+      
+      # birth/recruitment ---------------------                                                     
+      if (ss ==1){
+        pop[1,as.character(y),as.character(ss),] <- recruitment;
+      }
[…]

### 2017-12-20 doc/previous versions/MSE_DEC.tex
changes scenarios MP

--- 2017-12-06 doc/MSE_OCT.tex
@@
 TBD
-Farcas 2016:The case of fisheries management illustrates how the inherent structural instability of ecosystems can have deep-running policy implications.
-We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses complex food webs in its operating model.
-Use DSVM with a tuning parameter that measures how important is to be near optimal.
-We use this DSVM to investigate how fishing locations change in response when fish biology is incorporated.}
+We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses a dynamic state variable model including errors in decision-making in its operating model.}
 
@@
 
-\keywords{word1; word2; word3; and word4}
+\keywords{management stategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield}
 
@@
 
-MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}.
-At the same time being simple enough to build and calibrate reliably with the available data.
-More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.
-
 There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
@@
 
+MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}.
+At the same time being simple enough to build and calibrate reliably with the available data.
+More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.
+In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Kell2007, Dichmont2013}.
+
 We review the potential of MSE, emphasazing the role of individual harvester decision-making and socioeconomic drivers on management effectiveness, i.e.
@@
 
-Using a MSE, the main objective of this work is to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
+Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 \label{sec2}
-The MSE approach is based upon a set of model of the "true" population dynamics of the species (called "operating model"; section \ref{sec2.1}).
-The operating model aims capturing the key processes in the dynamics of the fish population given the best knowledge available, and can be thought of as a minimum realistic model \cite{Punt1995}.
-The next step in the MSE is to simulate the process of monitoring the stock (called "observation model"; section \ref{sec2.2}).
-This resulted in simulated measurements such as individual harvester decision-making and the consequent biomass (including error).
-The observation data are then passed to the "management model" (section \ref{sec2.3}).
-The management model encompasses the harvest control rules (HCR).
-The HCR references to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level, changes in gear or spatial and temporal reallocations or restrictions of fishing effort.
-Management actions are rarely implemented without error.
+
+The essential elements of the MSE framework that we followed is summarized in Figure \ref{fig:MSE}.
+The operating model (OM) aimed capturing the key processes in the dynamics of the fish population given the best knowledge available, and can be thought of as a minimum realistic model \cite{Punt1995}.
+The OM was built using a set of models of the dynamics of fish stocks under fishing pressure and coupled with the effect on fishers of variations in stock status and availability (sections \ref{sec2.1} and \ref{sec2.2} respectively).
+This resulted in simulated measurements (called "observation model") such as individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
+The observation data were then passed to the "management model" (section \ref{sec2.3}).
+The management model encompassed the harvest control rules (HCR).
+The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures (section \ref{sec2.3}).
+
+% Management actions are rarely implemented without error.
 This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
@@
 By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=0.7\textwidth]{Figures/MSE_6.eps} 
+\caption{The conceptual framework and how it was mapped.}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=0.6\textwidth]{Figures/Areadynamics3.eps} 
+\caption{The conceptual framework and how it was mapped.}
+\label{fig:MSE}
+\end{figure}
 
@@
 The fishery was modelled in a individual based dynamic state variable model.
-In order to allowed for fishing effort allocation, the model was seasonally and spatially explicit.
+In order to allowed for fishing effort allocation, the model was seasonally (6 seasons) and spatially (areas A and B) explicit.
 The dynamics of the two fish species were governed by a simple age-structured model that is also seasonally and spatially explicit.
-In the age structured model for the fish species, individuals were assumed to be born at age 0, and in season 1.
-The number of fish of age $a$ at year $y$, in season $s$ and area $p$ was written as $N (a, y, s, p)$.
-The number of new-born individuals in the model was independent of the size of the adult population.
-Individuals could be born in each of the areas, such that
+The age structured model for the fish species was based on 4 cohorts, and individuals were assumed to be born at age 0, and in season 1.
+The number of fish $i$ of age $a$ at year $y$, in season $s$ and area $p$ was written as $N_i (a, y, s, p)$.
+The number of new-born individuals in the model was independent of the size of the adult population, but was function of area $p$ ($R_i (p)$).
+Numbers of recruits were set to 100 individuals during the first season, and spawning population of species 1 ocurred in area $A$, while species 2 did so in area B.
+Thus, individuals could be born in each of the areas, such that:
 \begin{equation}
-N (0, y, 1, p) = R (p)
+N_i (0, y, 1, p) = R_i (p)
 \end{equation}
 
+Moreover, combining the weight-based von Bertalanffy growth equation ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
 Mortality in the model could occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens.
@@
 Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species.
-The decrease in population numbers was thus the result of the catches ($C (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
+The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
 \begin{equation}
-C (a, y, s, p) = q(a) * E(y,s,p),
+C_i (a, y, s, p) = q(a) * E(y,s,p),
 \end{equation}
@@
 \begin{equation}
-N (a, y, s, p) = N (a, y, s, p) - C (a, y, s, p). 
+N_i (a, y, s, p) = N_i (a, y, s, p) - C (a, y, s, p). 
 \end{equation}
@@
 The size of that matrix  was defined by the number of age classes, seasons and number of immigrants into area $p$ ($im$) and emigrants leaving area $p$ ($em$) in the model.
-The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($p (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs.
+A 20 \% constant migration during any season was set between areas.
+The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs.
 \begin{equation}
-Mig (a, 1, s, im, em) = Mig(a, 1, s, im, em) + N (a, y, s, em) * p (a, y, s, i, em, im),
+Mig (a, 1, s, im, em) = Mig(a, 1, s, im, em) + N (a, y, s, em) * P (a, y, s, i, em, im),
 \end{equation}
@@
 \end{equation}
-
-Life history traits inputted were: (i)  during the first season, specie 1 spawned in $A$, while specie 2 spawned in $B$, (ii) every season, 20\% of specie 1 and 2 emigrated and immigrated between areas, (iii) growth following the von Berttalanffy equation (with the following parameters: $L\infty= 20$, $k= 0.093$), (iv) age-weight relationship ($W= L\infty*(1-e^{-k*a})$), (v) at the beginning of each year the fishing recruitment was stablish in 100.
-
-The purpose of the operating model was not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.
 
@@
 where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
-Prices of fish were assumed to be dependant over fish age and time (subsection \ref{sec2.2.1}).
+Prices of fish were assumed to be dependant over fish age (subsection \ref{sec2.2.1}).
 $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table ~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
@@
 In contrast, if sigma was small then all vessels concentrated in the optimal location (but note that $\sigma$ should be $> 0$).
-For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 1.1 x 10^{11}$, but was generally of the order of $1.1 x 10^5$ in magnitude).
+For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
 
@@
 \centering
-\caption{Fishery parameterization.}
+\caption{Model and fishery conditioning.}
 \label{t:fishery}
-\resizebox{0.7\textwidth}{!}{%
-\begin{tabular}{@{}lc lc @{}} 
+\resizebox{0.9\textwidth}{!}{%
+\begin{tabular}{@{}lc lc lc @{}} 
 \hline
-\textbf{model}                              &               & \textbf{Variable costs}                        &      \\\hline%&\textbf{Market value}&  \\ \hline
+\textbf{model}                              &               & \textbf{Fishery variable costs}                &      &\textbf{Fishery }&  \\ \hline
 & & & \\%& & \\
-number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  \\%& sp1 (Euro kg\textsuperscript{-1})                 & 1000 \\ 
-$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0   \\% & sp2 (Euro kg\textsuperscript{-1})                 & 1000   \\
-number of areas                             & 2             &crew share                                      & 35\% \\%& & \\
-number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    \\%& &\\
+number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  & q                 & 0.0005 \\ 
+$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0    & effort (p,s)      & 1   \\
+number of areas                             & 2             &crew share                                      & 35\% & CPUE & mean($pop*q*wts(yy-2:yy)$)\\
+number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & CPUE dev& $0.08*CPUE$\\
 intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
@@
 \label{sec2.2.1}
-% Prices of fish have been assumed to be constant over fish age and time.
-Common knowledge and case studies indicated that size and weight of a fish could be important attributes when determining its economic value.
-However, so far there has been no broader assessment whether the size-dependence of price really is prevalent and significant in fisheries.
-The results from \cite{Zimmermann2011, Zimmermann2013} underline that size-dependent pricing is common in Norwegian fisheries, and is therefore also of high relevance for decisions in fisheries management.
-Therefore, the resulting data of mean weights, average prices and total yield per weight classes and year were used for further analysis.
-We fitted linear mixed models (Bates and Maechler, 2010) and general additive models (Wood, 2010) in R (R: Development Core Team, 2012), using mean price and mean body weight as the response variable and covariate, respectively.
-Thus, the statistical models took the form
+Prices of fish have been assumed to be fixed over time but influenced by the size of the catch.
+Common knowledge and case studies indicate that size and weight of a fish could be important attributes when determining its economic value \cite{Zimmermann2011, Zimmermann2013}.
+Therefore, the resulting data of mean weights and average prices were used for further analysis.
+We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively.
+Thus, following \cite{Zimmermann2011}, the statistical models took the form:
 \begin{equation}
- p_{i} \sim c_{0} + c_{1} w_{i} + f (y_{i}), 
+ pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
 \end{equation}
-where p is price of observation i, wi is the mean weight of the corresponding weight class, and the term f(yi) corresponds to the effect of year, treated either as a random effect or as a smooth term.
-Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models.
+where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes.
+$\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$.
+We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
+ %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models.
 Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results.
@@
 \centering
-\caption{Size dependant pricing.}
+\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
 \label{t:pricing}
 \resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}lcccccc@{}} 
+\begin{tabular}{@{}lcccc@{}} 
 \hline
-\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{6}{c}{season} \\ 
-age                      & 1     & 2     & 3    & 4    & 5    & 6    \\ \hline
+\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
+                         & 1     & 2     & 3    & 4   \\ \hline
                       \\
-1                       &    &    &   &   & & \\
-2                       &    &    &   &   & &\\
-3                       &    &    &   &   & &\\
-4                       &    &    &   &   & &\\
+species 1                &  778  & 1354   & 1780  & 2096  \\
+species 2                &  778  & 1354   & 1780  & 2096  \\
 \\
@@
                                 & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
-                                & \textit{(landings selectivity)}  & \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
+                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
 & & &  \\
 Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
-Manager catch perception        & $C = L + D$                    &        $C = L$                       &  $C = L$  \\
+Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)                      &  $C = L$ ($D \neq 0$) \\
+Selectivity ($l_{ratio}$)         &  $ \neq 1$                     & $=1$                                 & $=1$\\
 Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\
-Selectivity ($L/C$)             &  $ \neq 1$                     & $=1$                                 & $=1$\\
+
 &  & &  \\\hline
@@
 
-HCRs in the form of Quotas were determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeded Fmsy.
+HCRs in the form of quotas were determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeded Fmsy.
 This required that fishing mortality for each stock must not exceed Fmsy, which was a limit reference point.
@@
 
-The resulted biomass harvested rate in a year $y$, depended in the manager catch perception and the standing biomass:
+The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table ~\ref{t:fishery}) and the standing biomass:
 \begin{equation}
@@
                 % Nominator
-                {C_i numbers(a, y, s, p)}
+                {C_{i (numbers)}(a, y, s, p)}
                 % Denominator
-                {C_i numbers(a, y, s, p) \times B_i numbers(a, y, s, p)},
+                {C_{i (numbers)}(a, y, s, p) \times B_{i (numbers)}(a, y, s, p)},
 \end{equation}
@@
 \begin{equation}
- lratio (a, y, s) = \frac
+ l_{ratio} (a, y, s) = \frac
                 % Nominator
-                {L_i numbers(a, y, s, p)}
+                {L_{i (numbers)}(a, y, s, p)}
                 % Denominator
-                {C_i numbers(a, y, s, p)},
+                {C_{i (numbers)}(a, y, s, p)},
 \end{equation}
+
+Through a yield per recruitment model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.
 
@@
                 % Nominator
-                {\sum( [\frac{H_{MSY}}{\bar{H}} \times H (a, s) \times lratio (a, s) \times B numbers(a, s)] \times wts) }
+                {\sum( [\frac{H_{MSY}}{\bar{H}} \times H (a, s) \times l_{ratio} (a, s) \times B_{(numbers)}(a, s)] \times wts) }
                 % Denominator
-                {number of vessels}, 
+                {number\  of\  vessels}, 
 \end{equation}
 \begin{equation}
- Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85)), 
+ Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85), 
 \end{equation}
@@
 
-%- No quota constrained
-
-%- Through a yield per recruit model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.
-Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.
-The principal purpose of yield per recruit models is to get an idea of the effect of selection pattern and fishing mortality on the yield from a fixed number of individuals that enters the fishery.
-
-%Therefore if we have a harvest rate of 0.32, we will need to reduce the harvest in a 68\%.
-Reduce quota in next run in 68\%.
-From this reduction in quota and the subsequent reduction in IQ, this scenario will be evaluated allowing and not allowing the fleet to discard.
-
 \section{Results}
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand1.eps} 
 \end{minipage}
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand1.eps} 
 \end{minipage}
@@
 
+\newpage
+\section{Discussion}
+\label{sec4}
+tbd
+prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative.
+Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future.
+This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.
+
 The largest yield (or catch) that can be taken from a species' stock over an indefinite period.
@@
 
-\newpage
-\section{Discussion}
-\label{sec4}
-tbd
-prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative.
-Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future.
-This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.
-
 \begin{notes}[Acknowledgements]

### 2017-12-20 doc/previous versions/MSE_OCT.tex
changes scenarios MP

--- 2017-12-06 doc/MSE_OCT.tex
@@
 By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
+
+\begin{figure}[!h]
+\centering
+\begin{minipage}{.4\textwidth}
+\includegraphics[width=\textwidth]{Figures/Areadynamics.eps} 
+\end{minipage}
+\hfill
+\begin{minipage}{.5\textwidth}
+\includegraphics[width=\textwidth]{Figures/MSE.eps} 
+\end{minipage}
+\caption{The conceptual framework and how it was mapped.}
+\end{figure}
 
@@
 \label{sec2.2.1}
-% Prices of fish have been assumed to be constant over fish age and time.
-Common knowledge and case studies indicated that size and weight of a fish could be important attributes when determining its economic value.
-However, so far there has been no broader assessment whether the size-dependence of price really is prevalent and significant in fisheries.
-The results from \cite{Zimmermann2011, Zimmermann2013} underline that size-dependent pricing is common in Norwegian fisheries, and is therefore also of high relevance for decisions in fisheries management.
-Therefore, the resulting data of mean weights, average prices and total yield per weight classes and year were used for further analysis.
-We fitted linear mixed models (Bates and Maechler, 2010) and general additive models (Wood, 2010) in R (R: Development Core Team, 2012), using mean price and mean body weight as the response variable and covariate, respectively.
-Thus, the statistical models took the form
+Prices of fish have been assumed to be fixed over time but influenced by the size of the catch.
+Common knowledge and case studies indicated that size and weight of a fish could be important attributes when determining its economic value \cite{Zimmermann2011, Zimmermann2013}.
+Therefore, the resulting data of mean weights and average prices were used for further analysis.
+We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively.
+Thus, following \cite{Zimmermann2011}, the statistical models took the form:
 \begin{equation}
- p_{i} \sim c_{0} + c_{1} w_{i} + f (y_{i}), 
+ pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
 \end{equation}
-where p is price of observation i, wi is the mean weight of the corresponding weight class, and the term f(yi) corresponds to the effect of year, treated either as a random effect or as a smooth term.
-Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models.
+where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes.
+$\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$.
+ %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models.
 Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results.
@@
 \resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}lcccccc@{}} 
+\begin{tabular}{@{}lcccc@{}} 
 \hline
-\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{6}{c}{season} \\ 
-age                      & 1     & 2     & 3    & 4    & 5    & 6    \\ \hline
+\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
+                         & 1     & 2     & 3    & 4   \\ \hline
                       \\
-1                       &    &    &   &   & & \\
-2                       &    &    &   &   & &\\
-3                       &    &    &   &   & &\\
-4                       &    &    &   &   & &\\
+species 1                &  778  & 1354   & 1780  & 2096  \\
+species 2                &  778  & 1354   & 1780  & 2096  \\
 \\
@@
                                 & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
-                                & \textit{(landings selectivity)}  & \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
+                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
 & & &  \\
@@
 \begin{equation}
- Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85)), 
+ Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85), 
 \end{equation}

### 2017-12-20 RDynState5NAsigmaseason6Age/src/RDynState.cpp

@@ -1,0 +1,875 @@
+#include <Rdefines.h>
+#include <Rinternals.h>
+#include <R.h>
+
+#include <vector>
+#include <math.h>
+#include <stdlib.h>
+#include <time.h>   
+#include <algorithm>
+#include <numeric>
+#include "matalloc.h"
+#include <omp.h>
+
+using namespace std;
+
+int             outofbounds_int;
+bool            outofbounds_bool;
+double          outofbounds_double;
+
+// 160 by 98  works fine
+#define SPP1CAPACITY      950    // max number of first choke sp (larger than MAXNOINC * MAXHORIZON *NOSIZES= 64 * 5 * 2 = 640) 
+#define SPP2CAPACITY      950    // max number of second choke sp (MAXNOINC * MAXHORIZON *NOSIZES= 64* 5* 2 = 640)
+#define NOSPEC              5    // number of species used in the analysis
+#define NOSIZES             4    // number of size classes used in the analysis
+#define MAXNOINC           30    // max number of increments
+#define MAXHORIZON          6    // number of seasons to fish (= time)
+
+typedef unsigned UFINT;
+
+typedef float         (**FTYPE); /* make pointer called FTYPE for FF0Star and FF1 arrays (that only has states and stores V for best choices) */
+typedef float         (***FCTYPE); /* make pointer called FCTYPE for FF0 array (that has states and patch, storing V for all patches/choices)_  */
+typedef float         (****ITYPE); /* make pointer called ITYPE optimal choice array*/
+typedef float         (*****PTYPE);/* make pointer called PTYPE for information on statistical distribution prey*/
+typedef float         (****ATYPE);/* make pointer called ATYPE for information on statistical distribution prey aggregated over sizes*/
+typedef float         (***PITYPE); /* make pointer called PITYPE for price information*/
+
+
+/* random number generator*/
+float ranf ()
+{
+  return ((float)rand()/RAND_MAX); //genereert random getal tussen 0 en 1
+}
+
[…]

### 2017-12-20 RDynState5NAsigmaseason6_age/src/RDynState.cpp

@@ -1,875 +1,0 @@
-#include <Rdefines.h>
-#include <Rinternals.h>
-#include <R.h>
-
-#include <vector>
-#include <math.h>
-#include <stdlib.h>
-#include <time.h>   
-#include <algorithm>
-#include <numeric>
-#include "matalloc.h"
-#include <omp.h>
-
-using namespace std;
-
-int             outofbounds_int;
-bool            outofbounds_bool;
-double          outofbounds_double;
-
-// 160 by 98  works fine
-#define SPP1CAPACITY      750    // max number of first choke sp (larger than MAXNOINC * MAXHORIZON *NOSIZES= 64 * 5 * 2 = 640) 
-#define SPP2CAPACITY      750    // max number of second choke sp (MAXNOINC * MAXHORIZON *NOSIZES= 64* 5* 2 = 640)
-#define NOSPEC              5    // number of species used in the analysis
-#define NOSIZES             4    // number of size classes used in the analysis
-#define MAXNOINC           30    // max number of increments
-#define MAXHORIZON          6    // number of seasons to fish (= time)
-
-typedef unsigned UFINT;
-
-typedef float         (**FTYPE); /* make pointer called FTYPE for FF0Star and FF1 arrays (that only has states and stores V for best choices) */
-typedef float         (***FCTYPE); /* make pointer called FCTYPE for FF0 array (that has states and patch, storing V for all patches/choices)_  */
-typedef float         (****ITYPE); /* make pointer called ITYPE optimal choice array*/
-typedef float         (*****PTYPE);/* make pointer called PTYPE for information on statistical distribution prey*/
-typedef float         (****ATYPE);/* make pointer called ATYPE for information on statistical distribution prey aggregated over sizes*/
-typedef float         (***PITYPE); /* make pointer called PITYPE for price information*/
-
-
-/* random number generator*/
-float ranf ()
-{
-  return ((float)rand()/RAND_MAX); //genereert random getal tussen 0 en 1
-}
-
[…]

### 2017-12-20 complex model/complexmodel_DEC.R

@@ -123,9 +123,10 @@ areas         <- c("a", "b")
 stab.model    <- 10
 NUMRUNS       <- 80
 MPstart       <- 40
+MPstartLO     <- 65
 SIMNUMBER     <- 700 #pos
 SIGMA         <- 40 #sig 
-SPP1DSCSTEPS  <- 0
+SPP1DSCSTEPS  <- 2
 SPP2DSCSTEPS  <- 0
 endy          <- stab.model + NUMRUNS
 Linf          <- 20
@@ -134,9 +135,9 @@ wts           <- Linf*(1-exp(-K*ages))
 q             <- 0.0005
 natmortality  <- 0.0001
 migconstant   <- 0.2
-scenario      <- "1" 
-sp1price      <- slope1price <- 150
-sp2price      <- slope2price <- 150
+sp1price      <- sp2price      <- 150
+slope1price <- 150
+slope2price <- 150
 # scenario I: discarding is not allowed, YPR based in C (C=L)
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
 # scenario III: discarding ocurred but not perceived, YPR based in L, hr wanted based in landings
@@ -188,15 +189,14 @@ sp1Price <- array(c(sp1price+ slope1price*((wts-mean(wts)/mean(wts)))), dim=c(le
 sp2Price <- array(c(sp2price+ slope2price*((wts-mean(wts)/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 sp3Price <- sp4Price <- sp5Price <- array(c(0), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 #---effort and prices used (note that now c is removed (but that if other runs, then make sure to fix/remove code that removes "c" option)                                                                                         
-
[…]

### 2018-04-16 doc/MSE_MAR_18.tex
march

--- 2017-12-20 doc/previous versions/MSE_DEC.tex
@@
 More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.
-In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Kell2007, Dichmont2013}.
+In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Dichmont2013, Kell2007}.
 
@@
 
-Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependat behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
+Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{Figures/MSE_6.eps} 
+\includegraphics[width=0.7\textwidth]{Figures/MSE.eps} 
 \caption{The conceptual framework and how it was mapped.}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{Figures/Areadynamics3.eps} 
+\includegraphics[width=0.6\textwidth]{Figures/Areadynamics.eps} 
 \caption{The conceptual framework and how it was mapped.}
@@
 To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
-A dynamic state variable model \cite{Houston1999, ClarkandMangel2000} was used to model location choice, extending the model structure in \cite{Batsleer2015}.
+A dynamic state variable model \cite{ClarkandMangel2000, Houston1999} was used to model location choice, extending the model structure in \cite{Batsleer2015}.
 In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations.
@@
 where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
-Prices of fish were assumed to be dependant over fish age (subsection \ref{sec2.2.1}).
-$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table ~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
+Prices of fish were assumed to be dependent over fish age (subsection \ref{sec2.2.1}).
+$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
 The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$.
@@
 
-Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in ~\ref{eq:solve}.
+Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}.
 This was done for 700 individual vessels.
@@
 & & & \\%& & \\
-number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 150  & q                 & 0.0005 \\ 
+number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 1240  & q                 & 0.0005 \\ 
 $\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0    & effort (p,s)      & 1   \\
@@
 
-\subsubsection{Size dependant pricing}
+\subsubsection{Size dependent pricing}
 \label{sec2.2.1}
@@
 \resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}lcccc@{}} 
+\begin{tabular}{@{}llcccc@{}} 
 \hline
 \textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
-                         & 1     & 2     & 3    & 4   \\ \hline
+&                         & 1     & 2     & 3    & 4   \\ \hline
                       \\
-species 1                &  778  & 1354   & 1780  & 2096  \\
-species 2                &  778  & 1354   & 1780  & 2096  \\
+Scenario I &              &       &        &       &     \\                       
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  778  & 1354   & 1780  & 2096  \\
+                      \\
+Scenario II &              &       &        &       &     \\ 
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  464  & 752    & 965   & 1123 \\
 \\
@@
 
-The three management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
+The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
 These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
@@
 \label{scenarios}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{l c c  c }
+\resizebox{0.8\textwidth}{!}{%
+\begin{tabular}{l c c}
 \hline
-                                & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
-                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
-& & &  \\
-Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
-Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)                      &  $C = L$ ($D \neq 0$) \\
-Selectivity ($l_{ratio}$)         &  $ \neq 1$                     & $=1$                                 & $=1$\\
-Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\
+                                & \textbf{MP}            & \textbf{LO}                  \\ 
+                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
+& &   \\
+Discarding                      & Allowed                        & Not allowed                 \\
+Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
+Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
+Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
 
-&  & &  \\\hline
+&  &   \\\hline
 \end{tabular}
@@
 
-\item 
-\textbf{Scenario III: LO not fully enforced}, if LO is not fully enforced, therefore, there will be the incentive to continue discarding.
-However the manager could think that is fully enforced and assumed that only what is landed contribute to yield.
-This is the typicall situation of a choke species, where the management target such as $F_{MSY}$ cannot be easily achieved, because reductions in TACS do not necessarily translate into the expected reduction in fishing mortality.
-Rather, they may translate into increased discards if the fishery is mixed and continues fishing for other species beyond the exhaustion of the TAC of the species in question.
-
 \end{enumerate}
@@
 
-The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table ~\ref{t:fishery}) and the standing biomass:
+The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table~\ref{t:fishery}) and the standing biomass:
 \begin{equation}
@@
 
+We simulated the effects of implementing a Management Plan (MP) and the Landing Obligation (LO) without any exceptions or flexibility.
+To be able to analyse the results of this complex actions of implementing the Mp and the LO, the implementations should start simultaneously for all stocks and fleets.
+Under the MP, the Harvest Control Rule for the TACs is to advise $F_{MSY}$ unless the biomass falls below a trigger biomass.
+If the biomass falls below this value, the advised TAC is based on the biomass trends.
+To check for trends, the latest biomass levels are compared to previous ones.
+If there is a decrease of a 20\% or more, the TAC is reduced by 15\%.
+If the comparision shows an increase of more then 20 \% the TAC is increased by 15\%.
+
 \newpage
 %\begin{landscape}
-Scenario I:
+Scenario I: prices of the two species are equal (a truly mixed fishery)
 %----------------------------------------------------------------------------------------------------------------------------------------
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_disc_YieldLand1.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioI.eps} 
 \end{minipage}
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_disc_YieldLand1.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioI.eps} 
 \end{minipage}
@@
 
-Scenario II:
+Scenario II: prices of the species 2 are 50\% of species 1 prices
 %----------------------------------------------------------------------------------------------------------------------------------------
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioII.eps} 
 \end{minipage}
@@
 \begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO.eps} 
+\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioII.eps} 
 \end{minipage}
@@
 
-%----------------------------------------------------------------------------------------------------------------------------------------
+\subsection{ catch decisions and location choice}
+\label{sec3.1}
+When discarding was allowed, but TACs changed by 15\% based on previous year but trying approximate to Fmsy, TACS showed a reduction and discards increased.
 
-Scenario III:
-
-\begin{figure}[!h]
-\centering
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_nocompliance.eps}  
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-\centering
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_nocompliance.eps} 
-\end{minipage}
-\caption{}
-\label{fig:effort_lo}
-\end{figure}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
-
-\subsection{Simulated catch decisions and location choice}
-\label{sec3.1}
+The implementation of LO caused an increase in TACs and revenues were higher than when discarding was allowed.
+Some fishing opportunities are always lost under the landing obligation; the final fishing mortalities are lower than the target fishing mortalities (at least for one species).
+However, this does not necessarily imply that the catches will be lower with landing obligation than without it.
+Catches will be reduced in the initial years of landing obligation implementation.
+However, at Spawning stock biomass increases with LO, the catches will increase too.
+Although the fishing opportunities are lost due to the choke effect, the biomass of some stocks might increase; after a few years, this extra abundance could be converted into more catches.
+This implies that are fleets that, in the mid-term, could benefit from the landing obligation.
 

### 2018-04-16 doc/MSE_MAR_18.tex
changes commented by skype

@@
 Thus, individuals could be born in each of the areas, such that:
+
 \begin{equation}
@@
 The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
+
 \begin{equation}
@@
 \end{equation}
+
 and
+
 \begin{equation}
@@
 It was assumed that fishing mortality ocurred before migration ocurred.
-Migration for each species was defined by a matrix $Mig$ that defines the impact of migration (.e.
+Migration for each species was defined by a matrix $M$ that defines the impact of migration (.e.
 immigration-emigration) on a given stock was also dependent on the relative sizes of the stocks.
@@
 A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs.
+The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+
 \begin{equation}
-Mig (a, 1, s, im, em) = Mig(a, 1, s, im, em) + N (a, y, s, em) * P (a, y, s, i, em, im),
+M (a, 1, s, im, em) = M (a, 1, s, im, em) + N (a, y, s, em) * P (a, y, s, i, em, im),
 \end{equation}
+
 and then population numbers
+
 \begin{equation}
-N (a, y, s, im) = N (a, y, s, im) +  Mig (a, 1, s, im, em).
+N (a, y, s, im) = N (a, y, s, im) +  M (a, 1, s, im, em).
 \end{equation}
@@
 \end{equation}
+
 where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel.
@@
 \end{equation}
+
 where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
@@
 If $U^*$ was the expected utility at the optimal choice for a given t, we set
+
 \begin{equation}
@@
 \end{equation}
+
 and then defined the probability of a choice for a given area and discarding as
+
 \begin{equation} \label{eq:solve}
@@
 \label{t:fishery}
-\resizebox{0.9\textwidth}{!}{%
-\begin{tabular}{@{}lc lc lc @{}} 
+\resizebox{0.6\textwidth}{!}{%
+\begin{tabular}{@{}lc@{}} 
 \hline
-\textbf{model}                              &               & \textbf{Fishery variable costs}                &      &\textbf{Fishery }&  \\ \hline
-& & & \\%& & \\
-number of vessels                           &700            &fuel costs (Euro day\textsuperscript{-1})       & 1240  & q                 & 0.0005 \\ 
-$\sigma $                                   & 40            &gear maintenance (Euro day\textsuperscript{-1}) & 0    & effort (p,s)      & 1   \\
-number of areas                             & 2             &crew share                                      & 35\% & CPUE & mean($pop*q*wts(yy-2:yy)$)\\
-number of seasons                           & 6             &landing costs (Euro t\textsuperscript{-1})      & 0    & CPUE dev& $0.08*CPUE$\\
-intial quota $Q_{sp1}$ and $Q_{sp2}$  & 200&                 &      \\%& & \\
-& & & \\\hline%& & \\\hline
+\textbf{model}                              &               \\ \hline
+&  \\%& & \\
+number of vessels                           &700            \\ 
+$\sigma $                                   & 40           \\
+number of areas                             & 2            \\
+number of seasons                           & 6            \\
+intial quota $Q_{sp1}$ and $Q_{sp2}$        & 200 \\%& & \\
+&  \\
+\textbf{Fishery variable costs}             &               \\ \hline   
+&  \\%& & \\
+fuel costs (Euro day\textsuperscript{-1})       & 1240  \\ 
+gear maintenance (Euro day\textsuperscript{-1}) & 0     \\ 
+crew share                                      & 35\% \\ 
+landing costs (Euro t\textsuperscript{-1})      & 0    \\ 
+&                 \\  
+\textbf{Fishery }&                 \\ \hline   
+&  \\%& & \\
+q                 & 0.0005 \\ 
+effort (p,s)      & 1  \\ 
+CPUE & mean($pop*q*wts(yy-2:yy)$)\\ 
+CPUE dev& $0.08*CPUE$\\ 
+&                 \\  \hline%& & \\\hline
 \end{tabular}
@@
 \resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}llcccc@{}} 
+\begin{tabular}{@{}lcccc@{}} 
 \hline
 \textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
-&                         & 1     & 2     & 3    & 4   \\ \hline
+                         & 1     & 2     & 3    & 4   \\ \hline
                       \\                      
-Scenario I &              &       &        &       &     \\                       
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  778  & 1354   & 1780  & 2096  \\
-                      \\
-Scenario II &              &       &        &       &     \\ 
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  464  & 752    & 965   & 1123 \\
+species 1                &  778  & 1354   & 1780  & 2096  \\
+species 2                &  778  & 1354   & 1780  & 2096  \\
+
 \\
@@
 \item
-\textbf{Scenario I: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
+\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
 Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
@@
 \item
-\textbf{Scenario II: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
+\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
 
@@
 We simulated the effects of implementing a Management Plan (MP) and the Landing Obligation (LO) without any exceptions or flexibility.
-To be able to analyse the results of this complex actions of implementing the Mp and the LO, the implementations should start simultaneously for all stocks and fleets.
+To be able to analyse the results of this complex actions of implementing the MP and the LO, the implementations should start simultaneously for all stocks and fleets.
 Under the MP, the Harvest Control Rule for the TACs is to advise $F_{MSY}$ unless the biomass falls below a trigger biomass.
@@
 
-\newpage
-%\begin{landscape}
-Scenario I: prices of the two species are equal (a truly mixed fishery)
-%----------------------------------------------------------------------------------------------------------------------------------------
-\begin{figure}[!h]
-\centering
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioI.eps} 
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioI.eps} 
-\end{minipage}
-\caption{}
-\end{figure}
-%\end{landscape}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
-
-Scenario II: prices of the species 2 are 50\% of species 1 prices
-%----------------------------------------------------------------------------------------------------------------------------------------
-\begin{figure}[!h]
-\centering
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioII.eps} 
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioII.eps} 
-\end{minipage}
-\caption{}
-\end{figure}
-
-\subsection{ catch decisions and location choice}
+\subsection{Catch decisions and location choice}
 \label{sec3.1}
@@
 This implies that are fleets that, in the mid-term, could benefit from the landing obligation.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.69\textwidth]{Figures/figurecatches.eps} 
+\caption{Caches and harvest rates- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.69\textwidth]{Figures/figurecatches2.eps} 
+\caption{Change in selectivity- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=.82\textwidth]{Figures/figureffort.eps} 
+\caption{Effort- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/figurerevenue.eps} 
+\caption{Revenues- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+%---
 
@@
 
-\newpage
 \bibliographystyle{plain}
@@
 
+\newpage
+\begin{notes}[Apendix]
+
+Different scenarios, considering wether prices of the two species are equal or not.
+
+\begin{table}[!h]
+\centering
+\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
+\label{t:pricing}
+\resizebox{0.8\textwidth}{!}{%
+\begin{tabular}{@{}llcccc@{}} 
+\hline
+\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
+&                         & 1     & 2     & 3    & 4   \\ \hline
+                      \\
+Scenario I &              &       &        &       &     \\                       
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  778  & 1354   & 1780  & 2096  \\
+                      \\
+Scenario II &              &       &        &       &     \\ 
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  621  & 1053    & 1373   & 1610 \\
+                      \\
+Scenario III &              &       &        &       &     \\ 
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  464  & 752    & 965   & 1123 \\
+                      \\
+Scenario IV &              &       &        &       &     \\ 
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  307  & 451    & 558   & 637 \\
+
+\\
+\hline 
+\end{tabular}
+}
+\end{table}
+
+%----------------------------------------------------------------------------------------------------------------------------------------
+\newpage
+\begin{landscape}
+\begin{figure}[!h]
+\centering
+\includegraphics[.45\textwidth]{Figures/Catch_scenarioI.eps} 
+\includegraphics[.45\textwidth]{Figures/Effort_scenarioI.eps}
+\caption{Scenario I: Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+\end{landscape}
+
+\newpage
+\begin{landscape}
+\begin{figure}[!h]
+\centering
+\includegraphics[.45\textwidth]{Figures/Catch_scenarioII.eps} 
+\includegraphics[.45\textwidth]{Figures/Effort_scenarioII.eps}
+\caption{Scenario II: prices of the species 2 are 75\% of species 1 prices)}
+\end{figure}
+\end{landscape}
+
+\newpage
+\begin{landscape}
+\begin{figure}[!h]
+\centering
+\includegraphics[.45\textwidth]{Figures/Catch_scenarioIII.eps} 
+\includegraphics[.45\textwidth]{Figures/Effort_scenarioIII.eps}
+\caption{Scenario III: prices of the species 2 are 50\% of species 1 prices)}
+\end{figure}
+\end{landscape}
+
+\newpage
+\begin{landscape}
+\begin{figure}[!h]
+\centering
+\includegraphics[.45\textwidth]{Figures/Catch_scenarioIV.eps} 
+\includegraphics[.45\textwidth]{Figures/Effort_scenarioIV.eps}
+\caption{Scenario IV: prices of the species 2 are 25\% of species 1 prices)}
+\end{figure}
+\end{landscape}
+
+%----------------------------------------------------------------------------------------------------------------------------------------
+
+%\begin{figure}[!h]
+%\centering
+%\begin{minipage}{.45\textwidth}
+%\includegraphics[width=0.9\textheight]{Figures/Catch_scenarioII.eps} 
+%\end{minipage}
+%\hfill
+%\begin{minipage}{.45\textwidth}
+%\includegraphics[width=0.9\textheight]{Figures/Effort_scenarioII.eps} 
+%\end{minipage}
+%\caption{Scenario II: prices of the species 2 are 50\% of species 1 prices}
+%\end{figure}
+
+\end{notes}
+
 \end{document}

### 2018-04-23 doc/MSE_MAR_18.tex
last modifications

@@
 
-The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices.
+The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices \cite{CFP2013}.
 Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other.
-LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation.
+LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation \cite{Alzorriz2016, Alzorriz2018}.
 
@@
 Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures.
-Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty.
+Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
 
@@
 A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($Mig (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+The migration in population numbers ($M (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
 
 \begin{equation}
-M (a, 1, s, im, em) = M (a, 1, s, im, em) + N (a, y, s, em) * P (a, y, s, i, em, im),
+M (a, 1, s, im, em) = N (a, y, s, em) * P (a, y, s, i, em, im),
 \end{equation}
@@
                       \\                      
-species 1                &  778  & 1354   & 1780  & 2096  \\
-species 2                &  778  & 1354   & 1780  & 2096  \\
-
+species 1 and 2            &  778  & 1354   & 1780  & 2096  \\
 \\

### 2018-04-23 doc/MSE_MAR_18.tex
started rewrite intro doc (had to remove appendix)

@@
 
-The new Common Fisheries Policy marks a number of key changes to European fisheries management, including the introduction of multi-annual management plans aimed at achieving the Maximum Sustainable Yield (MSY) for all stocks and the gradual introduction of a landings obligation (LO) to encourage more selective fishing practices \cite{CFP2013}.
-Their implementation cannot be seen in isolation from each other as strategies aimed at one objective may have consequences for achieving the other.
-LO and catch quotas should provide incentives for change in the fisheries, including adoption through taking up selective gears and spatiotemporal effort allocation \cite{Alzorriz2016, Alzorriz2018}.
-
-Fisheries management typically focuses on the sustainable exploitation of single target species, but the potentially negative effects on, for example, reaching other species conservation objectives, should also be evaluated.
-Achieving single species MSY in complex and dynamic fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017}.
-There is no unique way of translating the single-species MSY objective to the multispecies case.
-Maximisation of yield from one stock will generally require different strategies than maximisation of yield from another.
-Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures.
-Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
-
+The European fisheries management reform of 2013 included the introduction of multi-annual management plans aimed at achieving Maximum Sustainable Yield (MSY) for commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
+The LO  should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}.
+Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017, Batsleer2013}.
+Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
 There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
@@
 relocation costs of marine protected areas, the implementation of the landings obligation, have been modelled under these approaches.
-In contrast, to enable longer term prediction accounting for the impact of these choices on the populations and consequently on the future yields, this fleet dynamic model should be coupled to a biological dynamic model where the feedback between fleet and stock dynamics are explicity modelled.
+Using spatial and temporal distributions of catch rates within a single year, these models forecast  high costs in mixed fisheries in the short term as a result of the LO.
+These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{ Alzorriz2018, Batsleer2013,Batsleer2015}.
+This will result in fishig effort reallocation and early closures of fisheries once quota have been reached.
+However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies
+
+To enable longer term prediction of the effects of changes in fisheires management, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
+Such a link allow to explicilty model the feedback that exists between the fleet dynamics and the populaiton dynamics.
+Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures.
+Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
 
@@
 
-We review the potential of MSE, emphasazing the role of individual harvester decision-making and socioeconomic drivers on management effectiveness, i.e.
+We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e.
 if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}.
@@
 
-The essential elements of the MSE framework that we followed is summarized in Figure \ref{fig:MSE}.
-The operating model (OM) aimed capturing the key processes in the dynamics of the fish population given the best knowledge available, and can be thought of as a minimum realistic model \cite{Punt1995}.
-The OM was built using a set of models of the dynamics of fish stocks under fishing pressure and coupled with the effect on fishers of variations in stock status and availability (sections \ref{sec2.1} and \ref{sec2.2} respectively).
+A Managament Strategy Evaluation framework was used to forecast the dynamics of the mixed fishery.
+The essential elements of the framework are summarized in Figure \ref{fig:MSE}.
+The framework consists of an Operating Model (OM) and a Management Procedure (MP).
+The OM captured key processes in the dynamics of the fish populations, and can be thought of as a minimum realistic model \cite{Punt1995}.
 This resulted in simulated measurements (called "observation model") such as individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
@@
 
-\newpage
-\begin{notes}[Apendix]
-
-Different scenarios, considering wether prices of the two species are equal or not.
-
-\begin{table}[!h]
-\centering
-\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
-\label{t:pricing}
-\resizebox{0.8\textwidth}{!}{%
-\begin{tabular}{@{}llcccc@{}} 
-\hline
-\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
-&                         & 1     & 2     & 3    & 4   \\ \hline
-                      \\
-Scenario I &              &       &        &       &     \\                       
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  778  & 1354   & 1780  & 2096  \\
-                      \\
-Scenario II &              &       &        &       &     \\ 
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  621  & 1053    & 1373   & 1610 \\
-                      \\
-Scenario III &              &       &        &       &     \\ 
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  464  & 752    & 965   & 1123 \\
-                      \\
-Scenario IV &              &       &        &       &     \\ 
-&species 1                &  778  & 1354   & 1780  & 2096  \\
-&species 2                &  307  & 451    & 558   & 637 \\
-
-\\
-\hline 
-\end{tabular}
-}
-\end{table}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
-\newpage
-\begin{landscape}
-\begin{figure}[!h]
-\centering
-\includegraphics[.45\textwidth]{Figures/Catch_scenarioI.eps} 
-\includegraphics[.45\textwidth]{Figures/Effort_scenarioI.eps}
-\caption{Scenario I: Prices of the two species are equal (a truly mixed fishery)}
-\end{figure}
-\end{landscape}
-
-\newpage
-\begin{landscape}
-\begin{figure}[!h]
-\centering
-\includegraphics[.45\textwidth]{Figures/Catch_scenarioII.eps} 
-\includegraphics[.45\textwidth]{Figures/Effort_scenarioII.eps}
-\caption{Scenario II: prices of the species 2 are 75\% of species 1 prices)}
-\end{figure}
-\end{landscape}
-
-\newpage
-\begin{landscape}
-\begin{figure}[!h]
-\centering
-\includegraphics[.45\textwidth]{Figures/Catch_scenarioIII.eps} 
-\includegraphics[.45\textwidth]{Figures/Effort_scenarioIII.eps}
-\caption{Scenario III: prices of the species 2 are 50\% of species 1 prices)}
-\end{figure}
-\end{landscape}
-
-\newpage
-\begin{landscape}
-\begin{figure}[!h]
-\centering
-\includegraphics[.45\textwidth]{Figures/Catch_scenarioIV.eps} 
-\includegraphics[.45\textwidth]{Figures/Effort_scenarioIV.eps}
-\caption{Scenario IV: prices of the species 2 are 25\% of species 1 prices)}
-\end{figure}
-\end{landscape}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
-
-%\begin{figure}[!h]
-%\centering
-%\begin{minipage}{.45\textwidth}
-%\includegraphics[width=0.9\textheight]{Figures/Catch_scenarioII.eps} 
-%\end{minipage}
-%\hfill
-%\begin{minipage}{.45\textwidth}
-%\includegraphics[width=0.9\textheight]{Figures/Effort_scenarioII.eps} 
-%\end{minipage}
-%\caption{Scenario II: prices of the species 2 are 50\% of species 1 prices}
-%\end{figure}
-
-\end{notes}
-
 \end{document}

### 2018-04-24 doc/MSE_MAR_18.tex
continued editing text, methods section

@@
 
-The European fisheries management reform of 2013 included the introduction of multi-annual management plans aimed at achieving Maximum Sustainable Yield (MSY) for commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
-The LO  should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}.
+Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Bunnefeld2011}) seeks to study the implications of management strategies using simulation\cite{Punt2016}.
+Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}.
+Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable.
+As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy.The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
+
+The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}.
+Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
+New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed.
+These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual.
+The effect of a choice on the utility depends on the economic environment,   such as the home port of the vessel and the distance to fishing grounds, and the bioological enviroment, such as the spatial distribution of the resources.
+Using such dynamic variable state models, effects of e.g.
+relocation costs of marine protected areas have been modelled \cite{Dowling2011}.
+
+Dynamic state variables have also been used to describe discarding behaviour \cite{Gillis1995, Batsleer2013, Batsleer2015}.
+These studies may shed light on the potential outcomes of the European fisheries management reform of 2013.
+That reform included using the use of the Maximum Sustainable Yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
+Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO.
+These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{ Alzorriz2018}.
+This will result in fishig effort reallocation and early closures of fisheries once quota have been reached.
+However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
+Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
+Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
 Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017, Batsleer2013}.
 Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
-There is a substantial literature addressing the question of effort allocation in fisheries and the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
-New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010} have been developed.
-These models generally predict the effect in the short term (within a year) by optimizing an objective function and determine which area best suits this behaviour given the set of incentives that exist (which may also depend on factors such as size, home port, distance to fishing grounds and expected catch rates).
-Effects of e.g.
-relocation costs of marine protected areas, the implementation of the landings obligation, have been modelled under these approaches.
-Using spatial and temporal distributions of catch rates within a single year, these models forecast  high costs in mixed fisheries in the short term as a result of the LO.
-These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{ Alzorriz2018, Batsleer2013,Batsleer2015}.
-This will result in fishig effort reallocation and early closures of fisheries once quota have been reached.
-However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies
 
-To enable longer term prediction of the effects of changes in fisheires management, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
-Such a link allow to explicilty model the feedback that exists between the fleet dynamics and the populaiton dynamics.
-Whatever the management regime, the performance of management strategies is conditioned by population dynamics, but also by exploitation dynamics, and particularly by the response of the fleets to management measures.
-Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Venables2009}) seeks to study the likely implications of potential harvest policies, or strategies, by evaluating management scenarios on the operation of a mixed fishery and on the fished stocks by including stakeholders and by assessing its robustness to uncertainty \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
-
-MSE involves building a simulation model for the entire process under study, complex enough both to address the current knowledge and uncertainty on the dynamics of fish stocks under fishing pressure, the effect on fishers of variations in stock status and availability, and their responses to those changes and those in management regimes \cite{Venables2009}.
-At the same time being simple enough to build and calibrate reliably with the available data.
-More comprehensive MSEs may also take the effects of the fishery on the broader environment, and economic and social effects into account \cite{Dichmont2008, Fulton2007}.
-In such modelling framework, advice and management systems can be quantitatively evaluated through simulation testing \cite{Dichmont2013, Kell2007}.
-
+To enable longer term prediction of the effects of changes in fisheries management, and thus complete the Management Strategy Evaluation, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
 We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e.
@@
 
-A Managament Strategy Evaluation framework was used to forecast the dynamics of the mixed fishery.
+The Managament Strategy Evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
+The framework consists of an Operating Model (OM) and a Management Procedure (MP).
+In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality.
+(ii) A management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by FMSY reference points.
+(iii) a fleet of individual fishers who aim to make the best use of their annual quota.
+Within this mixed fishery, individual vessels make adaptive choices on fishing location and discarding that depend on the distribution of the resources, and the quota that they are allocated.
+Each of these elements is discussed in more detail below.
 The essential elements of the framework are summarized in Figure \ref{fig:MSE}.
-The framework consists of an Operating Model (OM) and a Management Procedure (MP).
-The OM captured key processes in the dynamics of the fish populations, and can be thought of as a minimum realistic model \cite{Punt1995}.
-This resulted in simulated measurements (called "observation model") such as individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
-The observation data were then passed to the "management model" (section \ref{sec2.3}).
-The management model encompassed the harvest control rules (HCR).
-The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
-Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures (section \ref{sec2.3}).
 
-% Management actions are rarely implemented without error.
-This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
-when and where harvest is taken) are not accounted for in the HCR.
-By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
+The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
+The operating model thus includes  individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
+The  management body used the MP to make its decision on how to respond to the state of the resource.
+This management procedure thus includes the data collection from the fishery, how these are interpreted, and the Harvest Control Rule that dictates the limits that are set to the fishery.
 
@@
 
-For purposes of model simplicity, we considered a hypothetical mixed fishery on two species.
-The fishery was modelled in a individual based dynamic state variable model.
-In order to allowed for fishing effort allocation, the model was seasonally (6 seasons) and spatially (areas A and B) explicit.
-The dynamics of the two fish species were governed by a simple age-structured model that is also seasonally and spatially explicit.
-The age structured model for the fish species was based on 4 cohorts, and individuals were assumed to be born at age 0, and in season 1.
-The number of fish $i$ of age $a$ at year $y$, in season $s$ and area $p$ was written as $N_i (a, y, s, p)$.
-The number of new-born individuals in the model was independent of the size of the adult population, but was function of area $p$ ($R_i (p)$).
-Numbers of recruits were set to 100 individuals during the first season, and spawning population of species 1 ocurred in area $A$, while species 2 did so in area B.
-Thus, individuals could be born in each of the areas, such that:
+In order to allow the fishery to make spatial and temporal choices, the model was seasonally (six seasons) and spatially (two areas) explicit.
+The dynamics of two fish stocks were modelled by an age-structured model that is also seasonally and spatially explicit.
+The maximum age in the age-structured fish model was 4 years.
+The number of fish of stock $i$ of age $a$ at year $y$, in season $s$, and area $p$ was written as $N_i (a, y, s, p)$.
+The two stocks are named "stock 1" and "stock 2".
+and the areas are named "area a" and "area b".
+Individuals were born at age 0, at the start of each year $y$, in season 1.
+The number of new-born individuals in the model was independent of the size of the adult population, but was function of area $p$, and written as $R_i (p)$.
+Hence,
 
@@
 
-Moreover, combining the weight-based von Bertalanffy growth equation ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
-Mortality in the model could occur from catches in the fishery, and natural causes such as predation, diseases, and scenesens.
-However, for simplicity we assumed that his natural mortality was negligible.
-Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimics long-lived species.
-The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$), which were in turn a function of age dependent catchability $q(a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
+In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks.
+The difference between the stocks is that all new-borns of stock 1 were born in "area a", while all new-borns of stock 2 were born in "area b".
+
+Mortality in the model resulted only from catches in the fishery.
+Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible in comparison.
+Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
+The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$), which were in turn a function of age dependent catchability $q_i (a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
+Thus,
 
 \begin{equation}
-C_i (a, y, s, p) = q(a) * E(y,s,p),
+C_i (a, y, s, p) = N_i (a, y, s, p) * q(a) * E(y,s,p).
 \end{equation}
 
-and
+The mortality reduces the population numbers as for seasons 2 to 6 as
 
 \begin{equation}
-N_i (a, y, s, p) = N_i (a, y, s, p) - C (a, y, s, p). 
+N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p), 
 \end{equation}
 
-It was assumed that fishing mortality ocurred before migration ocurred.
-Migration for each species was defined by a matrix $M$ that defines the impact of migration (.e.
+while in season 1, the population numbers were defined as
+
+\begin{equation}
+N_i (a+1, y, 1, p) = N_i (a, y, 6, p) - C_i (a, y, 6, p). 
+\end{equation}
+
+Migration for each species was defined by a matrix $M_i$ that defines the impact of migration (.e.
 immigration-emigration) on a given stock was also dependent on the relative sizes of the stocks.
@@
 A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($M (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+The migration in population numbers ($M_i (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
 
 \begin{equation}
-M (a, 1, s, im, em) = N (a, y, s, em) * P (a, y, s, i, em, im),
+M_i (a, 1, s, im, em) = N_i (a, y, s, em) * P (a, y, s, i, em, im),
 \end{equation}
@@
 \end{equation}
+
+Moreover, combining the weight-based von Bertalanffy growth equation ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
 
@@
 
+The management model encompassed the harvest control rules (HCR).
+The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures (section \ref{sec2.3}).
+
+% Management actions are rarely implemented without error.
+This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
+when and where harvest is taken) are not accounted for in the HCR.
+By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
+
 \section{Results}

### 2018-05-02 doc/MSE_MAR_18.tex
worked on methods
'finished' popdyn and DSVM

@@
 \label{sec1}
+
+We have to implement alpha and beta, and make sure that age in VBLG is seasonal
 
@@
 
-\subsection{Stock structure and dynamics}
+\subsection{Population dynamics}
 \label{sec2.1}
 
-In order to allow the fishery to make spatial and temporal choices, the model was seasonally (six seasons) and spatially (two areas) explicit.
-The dynamics of two fish stocks were modelled by an age-structured model that is also seasonally and spatially explicit.
-The maximum age in the age-structured fish model was 4 years.
+In order to allow the fishery to make spatial and temporal choices, the model needed to be seasonally and spatially explicit.
+The dynamics of the fish stocks were modelled using an age-structured model that was spatially explicit, with seasonal time steps.
+In each seasonal time step, fish grow, migrate, and die.
 The number of fish of stock $i$ of age $a$ at year $y$, in season $s$, and area $p$ was written as $N_i (a, y, s, p)$.
-The two stocks are named "stock 1" and "stock 2".
-and the areas are named "area a" and "area b".
+The ages in the model range between age 0 and age $A$, the maximum age in the model.
+the seasons range between season 1 and season $S$, the last season within each year.
 Individuals were born at age 0, at the start of each year $y$, in season 1.
-The number of new-born individuals in the model was independent of the size of the adult population, but was function of area $p$, and written as $R_i (p)$.
-Hence,
+The number of new-born individuals $R_i (p)$ in the model was a function of the area $p$ and independent of the size of the adult population.
+The population numbers at age 0 are thus
 
 \begin{equation}
-N_i (0, y, 1, p) = R_i (p)
+N_i (0, y, 1, p) = R_i (p).
 \end{equation}
-
-In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks.
-The difference between the stocks is that all new-borns of stock 1 were born in "area a", while all new-borns of stock 2 were born in "area b".
 
@@
 Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
-The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$), which were in turn a function of age dependent catchability $q_i (a)$, and fishing effort $E (y,s,p)$ in any time, season, and area.
-Thus,
+The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$).
+
+The mortality caused by the catches of the fishery reduce the population numbers.
+Because the model was seasonally structured, the population numbers for seasons 2 to $S$ were dependent on the previous season,
 
 \begin{equation}
-C_i (a, y, s, p) = N_i (a, y, s, p) * q(a) * E(y,s,p).
+N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p). 
 \end{equation}
 
-The mortality reduces the population numbers as for seasons 2 to 6 as
+the population numbers in season 1 depended on the numbers in season $S$ of the previous year,
 
 \begin{equation}
-N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p), 
+N_i (a+1, y, 1, p) = N_i (a, y, S, p) - C_i (a, y, S, p). 
 \end{equation}
 
-while in season 1, the population numbers were defined as
+Migration for each species was defined by an arrayx $M_i (a, s, em, im)$ that defined  immigration and emigration on a given stock relative to the stock sizes.
+The size of that array was defined by the number of age classes, seasons and number of areas.
+emigrants leave area $em$ and move to area $im$.
+The emigrated part of the population is then substracted from each of the areas, so that that population numbers per year and season remain unaffected by migration.
+
+Indivdual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight.
+The stock weights are thus calculated from the population numbers as
 
 \begin{equation}
-N_i (a+1, y, 1, p) = N_i (a, y, 6, p) - C_i (a, y, 6, p). 
+W_i(a,y,s,p) =  \alpha * (N_i (a, y, S, p) * L_{\infty_i} * (1-\exp^{(-K * a)}))^{\beta},
 \end{equation}
+ and the same is done for the catch weights.
 
-Migration for each species was defined by a matrix $M_i$ that defines the impact of migration (.e.
-immigration-emigration) on a given stock was also dependent on the relative sizes of the stocks.
-The size of that matrix  was defined by the number of age classes, seasons and number of immigrants into area $p$ ($im$) and emigrants leaving area $p$ ($em$) in the model.
-A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($M_i (a, 1, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model.
+The Dynamic State Variable model inputs consist of the expected individual catch rates, which are a random variable.
+This, random variable is normally distributed, and its mean $\hat c_i (a, y, s, p)$  is a function of population size, age dependent catchability $q_i (a)$, and fishing effort $E (y,s,p)$ in any year, season, and area.
 
 \begin{equation}
-M_i (a, 1, s, im, em) = N_i (a, y, s, em) * P (a, y, s, i, em, im),
+\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a) * E(y,s,p).
 \end{equation}
-
-and then population numbers
-
-\begin{equation}
-N (a, y, s, im) = N (a, y, s, im) +  M (a, 1, s, im, em).
-\end{equation}
-
-Moreover, combining the weight-based von Bertalanffy growth equation ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
 
@@
 
-To simulate the process of monitoring the stock, we developed models for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
-A dynamic state variable model \cite{ClarkandMangel2000, Houston1999} was used to model location choice, extending the model structure in \cite{Batsleer2015}.
-In order to accommodate each individual vessels set of choices, the model may be constrained by individual vessels quota for the individual species and will respond by changing their fishing pattern in terms of: (i) whether or not to go fishing in one of the two areas ($A$ and $B$), and (ii) whether to discard one or more species and age class combinations.
-The model had annual fines for exceeding landings or catch quota as in \cite{Alzorriz2018}).
-In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings or catch quotas at the end of the year:
+To simulate a fleet of individual fishing vessels,we used a Dynamic State Variable Model \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
+The model was used to model location choice, extending the model structure in \cite{Batsleer2015}.
+Each individual vessels had a set of choices, which include the choice to go fishing in a season,  location choice within that season, and the choice to discard one or more species and size classes during a season.
+The model had annual fines for exceeding landings quota as in \cite{Alzorriz2018}).
+In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings quotas at the end of the year:
 \begin{equation}
-\Phi (C_i, Q_i, F_i)= -\sum_i (max( 0, (C_i - Q_i))* F_i),
+\Phi (L_i, Q_i, F_i)= -\sum_i (\textrm{max}( 0, (L_i - Q_i))* F_i),
 \end{equation}
 
-where $C_i$ was the cumulative annual landings or catches for species $i$ for an individual vessel.
+where $L_i$ was the cumulative annual landings for species $i$ for an individual vessel.
 These cumulative landings defined the state of the individual.
-$Q_i$ was the annual individual quota for landings or catches for the quota species.
+$Q_i$ was the annual individual quota for landings for the different species.
 Individual quotas were not transferable.
-In order to reduce computation time, we set quota only for two species.
-$F_i$ was the fine per unit weight for exceeding individual landings or catch quotas was set to $3 x 10^3$ Euro day\textsuperscript{-1}.
-These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
+$F_i$ was the fine per unit weight for exceeding individual landings quota.
 
-The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, s)= \Phi (C_i, Q_i, F_i)$.
-For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$ (A, B and including area 0: “staying in port”), and to keep or discard any combination of the age classes caught of the quota species.
+The maximum expected utility between current season $s$ and the end of the year was $V (L_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, S)= \Phi (C_i, Q_i, F_i)$.
+For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$, or to stay in port.While fishing, any combination of the age classes caught of the quota species can be landed, or discarded.
 This discarding was defined by a matrix $d$.
@@
 \label{sec2.2.1}
-Prices of fish have been assumed to be fixed over time but influenced by the size of the catch.
-Common knowledge and case studies indicate that size and weight of a fish could be important attributes when determining its economic value \cite{Zimmermann2011, Zimmermann2013}.
+Prices of fish have been assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}.
 Therefore, the resulting data of mean weights and average prices were used for further analysis.
@@
 \end{table}
+
+\subsection{Case study parameterization}
+
+ (six seasons) and spatially (two areas) explicit
+The maximum age in the age-structured fish model was 4 years.
+The two stocks are named "stock 1" and "stock 2". and the areas are named "area a" and "area b".
+In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks.
+The difference between the stocks is that all new-borns of stock 1 were born in "area a", while all new-borns of stock 2 were born in "area b".
+
+In our example, the age dependent catchability $q_i (a)$ was assumed independent of age $a$.
+
+ A 20 \% constant migration during any season was set between areas.
+The migration inpopulation numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+
+predict fish weight (W) at length (L). while assuming  ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
+
+ was set to $3 x 10^3$ Euro day\textsuperscript{-1}.
+These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
 

### 2018-05-03 complex model/complexmodel.R
added LW relationship
Used more or less realistic parameters, but as a result had to increase
rec (from 100 to 200)

@@ -128,14 +128,20 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
     SPP1DSCSTEPS  <- 2
     SPP2DSCSTEPS  <- 0
     endy          <- stab.model + NUMRUNS
-    Linf          <- 20
-    K             <- 0.3
-    wts           <- Linf*(1-exp(-K*ages))
+    Linf          <- 50
+    K             <- 0.4
+    alpha         <- 0.00005
+    beta          <- 3
+    sages         <- array(seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)), dim=c(length(season),1,length(ages),1), dimnames=list(season=as.character(season),   year="all", cat=ages, option ="all"))
+    lens          <- Linf*(1-exp(-K*(sages)))
+    wts           <- alpha * lens ^ beta
+    wts           <- aperm(wts, c(3,2,1,4) ) 
+    
     q             <- 0.0005
     natmortality  <- 0.0001
     migconstant   <- 0.2
     
-    recs1          <- c(100,0) 
+    recs1          <- c(200,0) 
     mig1     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
     mig1[,,,"a","a"] <- -migconstant
     mig1[,,,"b","b"] <- -migconstant
@@ -143,7 +149,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
     mig1[,,,"b","a"] <- migconstant
     aperm( mig1,c(1,3,2,4,5))
     
-    recs2          <- c(0,100) 
+    recs2          <- c(0,200) 
     mig2     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
     mig2[,,,"a","a"] <- -migconstant
     mig2[,,,"b","b"] <- -migconstant

### 2018-05-03 doc/MSE_MAR_18.tex
modification in tex

@@
 
-Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Bunnefeld2011}) seeks to study the implications of management strategies using simulation\cite{Punt2016}.
+Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulation\cite{Punt2016}.
 Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}.
 Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable.
-As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy.The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
+As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy.
+The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
 
@@
 These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual.
-The effect of a choice on the utility depends on the economic environment,   such as the home port of the vessel and the distance to fishing grounds, and the bioological enviroment, such as the spatial distribution of the resources.
+The effect of a choice on the utility depends on the economic environment, such as the home port of the vessel and the distance to fishing grounds, and the biological enviroment, such as the spatial distribution of the resources.
 Using such dynamic variable state models, effects of e.g.
@@
 
-Dynamic state variables have also been used to describe discarding behaviour \cite{Gillis1995, Batsleer2013, Batsleer2015}.
+Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995}.
 These studies may shed light on the potential outcomes of the European fisheries management reform of 2013.
@@
 Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO.
-These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{ Alzorriz2018}.
+These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}.
 This will result in fishig effort reallocation and early closures of fisheries once quota have been reached.
@@
 Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
-Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017, Batsleer2013}.
+Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013,Ulrich2017}.
 Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
@@
 
-Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
+Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.
+To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 The operating model thus includes  individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
-The  management body used the MP to make its decision on how to respond to the state of the resource.
+The  management body used the MP to make its decision on how to respond to the state of the resource  (Figure \ref{fig:stockdyn}).
 This management procedure thus includes the data collection from the fishery, how these are interpreted, and the Harvest Control Rule that dictates the limits that are set to the fishery.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{Figures/MSE.eps} 
-\caption{The conceptual framework and how it was mapped.}
+\includegraphics[width=0.55\textwidth]{Figures/MSE.eps} 
+\caption{The Managament Strategy Evaluation framework including the Operating Model (OM) and the Management Procedure (MP).}
+\label{fig:MSE}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.6\textwidth]{Figures/Areadynamics.eps} 
-\caption{The conceptual framework and how it was mapped.}
-\label{fig:MSE}
+\includegraphics[width=0.55\textwidth]{Figures/Areadynamics.eps} 
+\caption{The Operating Model (OM) including the key processes in the dynamics of the fish populations and how it was mapped.}
+\label{fig:stockdyn}
 \end{figure}
@@
 \end{equation}
-where $\sigma$ was a tuning parameter that measures how important was to be near optimal.
-A high value of sigma resulted in uniform probabilities of choices, and that vessels were distributed uniformly among the different fishing areas.
-In contrast, if sigma was small then all vessels concentrated in the optimal location (but note that $\sigma$ should be $> 0$).
+where $\sigma$ was a tuning parameter that measured how important it was to be near the optimal choice.
+A large $\sigma$ resulted in uniform probabilities of choices, with vessels being distributed uniformly across the different fishing areas.
+In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$).
 For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
@@
 The maximum age in the age-structured fish model was 4 years.
-The two stocks are named "stock 1" and "stock 2". and the areas are named "area a" and "area b".
+The two stocks are named ``stock 1" and ``stock 2". and the areas are named ``area a" and ``area b". 
 In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks.
-The difference between the stocks is that all new-borns of stock 1 were born in "area a", while all new-borns of stock 2 were born in "area b".
+The difference between the stocks is that all new-borns of stock 1 were born in ``area a", while all new-borns of stock 2 were born in ``area b".
 

### 2018-05-03 doc/MSE_MAR_18.tex
continued work on methods

@@
 
-Indivdual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight.
-The stock weights are thus calculated from the population numbers as
+Individual body growth was modelled by a von Bertalanffy growth equation to convert numbers to lengths, and an allometric equation to convert length to weight.
+The weights for individuals in the stock and in the catches are thus calculated as
 
 \begin{equation}
-W_i(a,y,s,p) =  \alpha * (N_i (a, y, S, p) * L_{\infty_i} * (1-\exp^{(-K * a)}))^{\beta},
+w_i(a,s) =  \alpha * ( L_{\infty_i} * (1-\exp^{(-K * (a+(s/S))}))^{\beta}.
 \end{equation}
- and the same is done for the catch weights.
 
 The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model.
-The Dynamic State Variable model inputs consist of the expected individual catch rates, which are a random variable.
-This, random variable is normally distributed, and its mean $\hat c_i (a, y, s, p)$  is a function of population size, age dependent catchability $q_i (a)$, and fishing effort $E (y,s,p)$ in any year, season, and area.
+The Dynamic State Variable model inputs consist of the expected individual catch rates, which are random variables.
+These random variables were normally distributed, with means $\hat c_i (a, y, s, p)$ being a function of population size, age dependent catchability $q_i (a)$ in any year, season, and area,
 
 \begin{equation}
-\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a) * E(y,s,p).
+\hat c_i (a, y, s, p) = N_i (a, y, s, p) * q_i (a)  * w_i(a,s).
 \end{equation}
+The standard deviations $\Sigma_i (a, y, s, p)$ of the catch distributions were constant fraction  their means, using a ratio $\eta$.
 
-\subsection{The dynamic-state variable model}
+\subsection{fleet dynamics}
 \label{sec2.2}
@@
 Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}.
-This was done for 700 individual vessels.
+For each year, these forward Monte Carlo Simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area
 
@@
 \centering
-\caption{Model and fishery conditioning.}
+\caption{Model parameters.}
 \label{t:fishery}
@@
 \hline
-\textbf{model}                              &               \\ \hline
+
+\textbf{Population dynamics}                                    &                \\ \hline
 &  \\%& & \\
@@
 &  \\
-\textbf{Fishery variable costs}             &               \\ \hline   
+
+\textbf{Fleet dynamics}                                              &               \\ \hline   
 &  \\%& & \\
+Number of vessels                                                     & 700 \\ 
 fuel costs (Euro day\textsuperscript{-1})                 & 1240    \\ 
@@
 &                 \\
-\textbf{Fishery }&                 \\ \hline   
+
+\textbf{Fishery }                                                       &          \\ \hline   
 &  \\%& & \\
@@
 effort (p,s)                                                                & 1  \\ 
-CPUE & mean($pop*q*wts(yy-2:yy)$)\\ 
-CPUE dev& $0.08*CPUE$\\ 
+Ratio of standard deviations to means $\eta$         & $0.08$\\ 
 &                 \\  \hline%& & \\\hline
@@
 \label{sec2.2.1}
-Prices of fish have been assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}.
-Therefore, the resulting data of mean weights and average prices were used for further analysis.
-We fitted linear mixed models, using mean price and mean body weight as the response variable and covariate, respectively.
-Thus, following \cite{Zimmermann2011}, the statistical models took the form:
+Prices of fish were assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}.
+Following \cite{Zimmermann2011}, therelationship between fish price and body weight was modelled as:
 \begin{equation}
- pw_{i} \sim p_i \bar{w_i} + \beta_0 \times \frac{w_i-\bar{w_i}}{\bar{w_i}},
+ p_i (a,s) = \bar{ p}_i + \gamma_i \times \frac{w_i (a,s) -\bar{w_i}}{\bar{w_i}},
 \end{equation}
-where $p$ is the mean price of species $i$, $\bar{w_i}$ is the mean weight of the corresponding weight classes, and the fraction corresponded to the standarization of the mass relative to the mean observed individual catch mass over all ages classes.
-$\beta_0$ gave the price increase when individial mass was increased by $\bar{w}$.
-We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
- %Because landings corresponding to each weight class varied dramatically, sometimes by several orders of magnitude, we used log-transformed total annual yield per weight class as the statistical weighting factor in the above models.
-Alternatively, to test the sensitivity of the results, we did not apply any weighting but excluded data corresponding to the lowest 10\% quantile of annual yield, so that observations corresponding to very low yields would not influence the results.
-All models were fitted independently to each stock.
+where $p_i (a,s) $ is the price of species $i$ at age $a$ and season $s$.
+The price was a linear function of the weight of inidividuals of species $i$ at age $a$ and season $s$.
+$\bar{p}_i$ is the price of the species at the mean weight over the age range.
+The mean weight over the age range is $\bar{w_i}$.
+$\gamma_i$ gave the price increase when individial mass was increased by $\bar{w_i}$.
+
 \begin{table}[!h]
@@
 These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
+
+We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
 

### 2018-05-03 complex model/complexmodel_DEC.R
added LW relationship and changes in arrays dimensions

@@ -144,7 +144,7 @@ natmortality  <- 0.0001
 migconstant   <- 0.2
 sp1price      <- sp2price      <- 150
 slope1price <- 150
-slope2price <- 0.50*150
+slope2price <- 150 # 0.50*150
 
 # scenario I: discarding is not allowed, YPR based in C (C=L)
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches

### 2018-05-07 complex model/complexmodel_DEC.R
fixed bug in price calcs
there were some brackets missing that resulted in a unintend shift of
prices

@@ -198,8 +198,8 @@ catchMean(sp3)  <- catchMean(sp4) <- catchMean(sp5) <- array(0.01,dim=c(length(a
 catchSigma(sp3) <- catchSigma(sp4)<- catchSigma(sp5)<- array(0.0000001,dim=c(length(ages),length(season),length(areas)),dimnames=list(cat=ages,season=as.character(season),option =areas))
 
 #SIZE DEPENDENT PRICING, following Zimmermann et al. (2011)
-sp1Price <- array(c(sp1price+ slope1price*((wts-mean(wts)/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
-sp2Price <- array(c(sp2price+ slope2price*((wts-mean(wts)/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
+sp1Price <- array(c(sp1price + slope1price*(((wts-mean(wts))/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
+sp2Price <- array(c(sp2price + slope2price*(((wts-mean(wts))/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 sp3Price <- sp4Price <- sp5Price <- array(c(0), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 #---effort and prices used (note that now c is removed (but that if other runs, then make sure to fix/remove code that removes "c" option)                                                                                         
 control     <- DynState.control(spp1LndQuota= 200,  spp2LndQuota=200, spp1LndQuotaFine= 3e6, spp2LndQuotaFine= 3e6, fuelUse = 1, fuelPrice = 150.0, landingCosts= 0,gearMaintenance= 0, addNoFishing= TRUE, increments= 25, spp1DiscardSteps= SPP1DSCSTEPS, spp2DiscardSteps= SPP2DSCSTEPS, sigma= SIGMA, simNumber= SIMNUMBER, numThreads= 20)

### 2018-05-07 complex model/complexmodel_DEC.R
changed slope of price function (now less steep)

@@ -143,8 +143,8 @@ natmortality  <- 0.0001
 
 migconstant   <- 0.2
 sp1price      <- sp2price      <- 150
-slope1price <- 150
-slope2price <- 150 # 0.50*150
+slope1price <- 100
+slope2price <- 100 # 0.50*150
 
 # scenario I: discarding is not allowed, YPR based in C (C=L)
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches

### 2018-05-07 complex model/complexmodel_DEC.R
doubled recruitment
doubled recruitment to account for lower prices that result from fixing
bug in previous commits

@@ -150,7 +150,7 @@ slope2price <- 100 # 0.50*150
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
 # scenario III: discarding ocurred but not perceived, YPR based in L, hr wanted based in landings
 
-recs1          <- c(200,0) 
+recs1          <- c(400,0) 
 mig1     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
 mig1[,,,"a","a"] <- -migconstant
 mig1[,,,"b","b"] <- -migconstant
@@ -158,7 +158,7 @@ mig1[,,,"a","b"] <- migconstant
 mig1[,,,"b","a"] <- migconstant
 aperm( mig1,c(1,3,2,4,5))
 
-recs2          <- c(0,200) 
+recs2          <- c(0,400) 
 mig2     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
 mig2[,,,"a","a"] <- -migconstant
 mig2[,,,"b","b"] <- -migconstant

### 2018-05-07 doc/MSE_MAR_18.tex
continued working on table parms

@@
 
-Mortality in the model resulted only from catches in the fishery.
-Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible in comparison.
-Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
-The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$).
-
-The mortality caused by the catches of the fishery reduce the population numbers.
+Mortality in the model resulted from fishery catches and natural causes such as predation, diseases, and scenesens.
+The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$) and a natural mortality constant $M_i$ that described natural mortality as a fixed fraction of the population.
+These mortalities reduced the population numbers among a cohort of fish.
 Because the model was seasonally structured, the population numbers for seasons 2 to $S$ were dependent on the previous season,
@@
 \begin{equation}
-N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p). 
+N_i (a, y, s+1, p) = N_i (a, y, s, p) - C_i (a, y, s, p) - M_i N_i (a, y, s, p) . 
 \end{equation}
 
-the population numbers in season 1 depended on the numbers in season $S$ of the previous year,
+Likewise, the population numbers in season 1 depended on the numbers in season $S$ of the previous year,
 
 \begin{equation}
-N_i (a+1, y, 1, p) = N_i (a, y, S, p) - C_i (a, y, S, p). 
+N_i (a+1, y, 1, p) = N_i (a, y, S, p) - C_i (a, y, S, p) - M_i N_i (a, y, s, p). 
 \end{equation}
 
-Migration for each species was defined by an arrayx $M_i (a, s, em, im)$ that defined  immigration and emigration on a given stock relative to the stock sizes.
+Migration for each species was defined by an array $D_i (a, s, em, im)$ that defined  immigration and emigration on a given stock relative to the stock sizes.
 The size of that array was defined by the number of age classes, seasons and number of areas.
@@
 Prices of fish were assumed to be dependent over fish age (subsection \ref{sec2.2.1}).
-$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $a$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
+$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
 The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$.
@@
 \resizebox{0.6\textwidth}{!}{%
-\begin{tabular}{@{}lc@{}} 
+\begin{tabular}{@{}llr@{}} 
 \hline
@@
 &  \\%& & \\
-number of vessels                                                      &700            \\ 
-$\sigma $                                                                    & 40            \\
-number of areas                                                         & 2             \\
-number of seasons                                                     & 6             \\
-intial quota $Q_{sp1}$ and $Q_{sp2}$                     & 200 \\%& & \\
+
+maximum age                                                            & $A$                    & 4               \\
+number of areas                                                       &                            & 2               \\
+number of seasons                                                   & $S$                     & 6               \\
+natural mortality                                                       & $M_i$                 & 0.0001     \\
+Asymptotic length                                                    &  $L_{\infty_i}$   & 50            \\
+Growth rate                                                              & $K_i$                 &  0.4          \\
+                                                                                 & $ \alpha_i$        &  0.00005   \\
+                                                                                 & $\beta_i $         &   3              \\
 &               \\
-
 \textbf{Fleet dynamics}                                           &               \\ \hline   
 &  \\%& & \\
-Number of vessels                                                     & 700 \\ 
-fuel costs (Euro day\textsuperscript{-1})                 & 1240    \\ 
-gear maintenance (Euro day\textsuperscript{-1})    & 0          \\ 
-crew share                                                                 & 35\%    \\ 
-landing costs (Euro t\textsuperscript{-1})                & 0           \\ 
+
+Number of vessels                                                    &                       & 700           \\ 
+Fuel costs (Euro day\textsuperscript{-1})               &                       & 1240         \\ 
+Gear maintenance (Euro day\textsuperscript{-1})  &                       & 0               \\ 
+Crew share                                                               &                       & 35\%        \\ 
+Landing costs (Euro t\textsuperscript{-1})              &                       & 0               \\
+Optimal choice error                                                 & $\sigma $       & 40             \\
+
 &                 \\  
-
 \textbf{Fishery }                                                      &             \\ \hline   
 &  \\%& & \\
-q                                                                                & 0.0005 \\ 
-effort (p,s)                                                                & 1  \\ 
-Ratio of standard deviations to means $\eta$         & $0.08$\\ 
+
+Intial quota                                                               &                       & 200 \\%& & \\
+catchability                                                               & $q_i (a)$        & 0.0005        \\ 
+effort (p,s)                                                               &                       & 1                 \\ 
+Price of species at mean weight                               & $\bar{p}_i$   & 150             \\
+Slope of species price                                              & $\gamma_i$   &  100            \\
+Fine for overshooting quota                                    &  $F_i$            & 3e6              \\
+Ratio of standard deviations to catch means          & $\eta$           & 0.08           \\ 
 &                 \\  \hline%& & \\\hline
@@
 We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
+
+Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible in comparison.
+Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
 

### 2018-05-08 doc/MSE_MAR_18.tex
continued working on description case study
Replaced table with prices by figure with price and length. The code for
this was
setwd("D://Nekane_MSE/doc/Figures")
setEPS()
postscript("prices.eps")
#png(filename=paste("MIXEDMP2_CATCH_", paste0("SIGMA ", SIGMA, ";
INCREMENTS ",control@increments, "; SIMNUMBER ",SIMNUMBER, ";
DISCARDSTEPS ",SPP1DSCSTEPS, "; MIGRATION ",migconstant, "; REC1
",paste(recs1, collapse = " "), "; REC2 ",paste(recs2, collapse = " "),
"; SP1PRICE ",paste0(round(sp1Price[,1]), sep = ', ', collapse =
''),";SP2PRICE ",paste0(round(sp2Price[,1]), sep = ', ', collapse = ''),
"

@@
 
-We have to implement alpha and beta, and make sure that age in VBLG is seasonal
+We have to make sure we are clear on the ages of the species: do we start at 0 or at 1?
 
@@
 
-\begin{table}[!h]
+\begin{figure}[!ht]
 \centering
-\caption{Linear mass-prices estimated based on mean prices per mass age-classes. The price slopes $\beta_0$ were assumed to be similar to mean prices.}
-\label{t:pricing}
-\resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}lcccc@{}} 
-\hline
-\textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
-                         & 1     & 2     & 3    & 4   \\ \hline
-                      \\                      
-species 1 and 2            &  778  & 1354   & 1780  & 2096  \\
-\\
-\hline 
-\end{tabular}
-}
-\end{table}
+\includegraphics[width=.69\textwidth]{Figures/prices.eps} 
+\caption{Lenghts and prices of the species}
+\end{figure}
 
@@
 
- (six seasons) and spatially (two areas) explicit
-The maximum age in the age-structured fish model was 4 years.
-The two stocks are named ``stock 1" and ``stock 2". and the areas are named ``area a" and ``area b". 
-In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks.
-The difference between the stocks is that all new-borns of stock 1 were born in ``area a", while all new-borns of stock 2 were born in ``area b".
+To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and six fishing seasons per year.
+There were two fish species in the model, both of which are caught by the fishery.
+The annual number of newborns is equal for the two species, and arbitrarily set to 400 per year.
+The species differ with respect to their nursery grounds: all individuals of species 1 are born in the southern area, while all individuals of species 2 are born in the northern area.
+The maximum age that any individual can reach was 4 years old
 

### 2018-05-14 doc/MSE_MAR_18.tex
updated tex file with spell check

@@
 These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}.
-This will result in fishig effort reallocation and early closures of fisheries once quota have been reached.
+This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.
 However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
@@
 Growth rate                                                              & $K_i$                 &  0.4          \\
-                                                                                 & $ \alpha_i$        &  0.00005   \\
-                                                                                 & $\beta_i $         &   3              \\
+Length-weight conversion factor                             & $ \alpha_i$        &  0.00005   \\
+Length-weight isomorphy factor                              & $\beta_i $         &   3              \\
 &               \\
@@
 where $p_i (a,s) $ is the price of species $i$ at age $a$ and season $s$.
-The price was a linear function of the weight of inidividuals of species $i$ at age $a$ and season $s$.
+The price was a linear function of the weight of individuals of species $i$ at age $a$ and season $s$.
 $\bar{p}_i$ is the price of the species at the mean weight over the age range.
 The mean weight over the age range is $\bar{w_i}$.
-$\gamma_i$ gave the price increase when individial mass was increased by $\bar{w_i}$.
+$\gamma_i$ gave the price increase when individual mass was increased by $\bar{w_i}$.
 
@@
 \includegraphics[width=.69\textwidth]{Figures/prices.eps} 
-\caption{Lenghts and prices of the species}
+\caption{Lengths and prices of the species}
 \end{figure}
@@
 
-To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and six fishing seasons per year.
+To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and six fishing seasons per year (Table \ref{t:fishery}).
 There were two fish species in the model, both of which are caught by the fishery.
@@
 The species differ with respect to their nursery grounds: all individuals of species 1 are born in the southern area, while all individuals of species 2 are born in the northern area.
-The maximum age that any individual can reach was 4 years old
+The maximum age that any individual can reach was 4 years old.
+During their life, individuals grew in length towards the asymptotic length, which was 50 cm.
+for both species.
+Conversion from length to weight was also equal for the two species (Table \ref{table:kysymys}).
+Migration was parameterized so that a gradual diffusion occured between the two areas, euqal to 20 percent of the difference in abundance. 
+ A 20 \% constant migration during any season was set between areas.
+The migration in population numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
 
-In our example, the age dependent catchability $q_i (a)$ was assumed independent of age $a$.
+In the case study, age-dependent catchability links the population biomassed to the catches in the fishery, and it is thus one of the crucial parameters determing the interaction between the two.
+In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
 
  A 20 \% constant migration during any season was set between areas.
-The migration inpopulation numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
-
-predict fish weight (W) at length (L). while assuming  ($L\infty= 20$, $k= 0.093$), with the age-weight relationship ($W= L\infty*(1-e^{-k*a})$); weight per fish was higher for the older fish.
+The migration in population numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
 

### 2018-05-14 doc/MSE_MAR_18.tex
Added figure with distributions
We need to check if legend is correct (i.e. high and low abundances are correctly coloured for the two species.
Code (not included in order not to change model code):
setEPS()
postscript("distributions.eps")
par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
split.screen(c(2,1), screen=1)-> ind
screen( ind[1])

image(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
box()
axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
axis(1,at= seq(0.,0.75,0.25),labels=

@@
 \centering
+\label{f:prices}
 \includegraphics[width=.69\textwidth]{Figures/prices.eps} 
@@
 The maximum age that any individual can reach was 4 years old.
-During their life, individuals grew in length towards the asymptotic length, which was 50 cm.
+During their life, individuals grew in length towards the asymptotic length, which was 50 cm (Figure \ref{f:prices}).
 for both species.
-Conversion from length to weight was also equal for the two species (Table \ref{table:kysymys}).
+Conversion from length to weight was also equal for the two species (Table \ref{t:fishery}).
 Migration was parameterized so that a gradual diffusion occured between the two areas, euqal to 20 percent of the difference in abundance.
- A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
+For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas  (Figure \ref{f:distributions}).
+Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible, and $M_i$ set to 0.0001.
+Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
 
-In the case study, age-dependent catchability links the population biomassed to the catches in the fishery, and it is thus one of the crucial parameters determing the interaction between the two.
+Age-dependent catchability links the population biomasses to the catches in the fishery, and is thus one of the crucial parameters determing the interaction between the two.
 In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
 
- A 20 \% constant migration during any season was set between areas.
-The migration in population numbers ($M_i (a, s, im, em)$) was thus the proportion ($P (a, y, s, i, em, im)$) of the population available or vulnerable to fishing after migration occurs:
-
- was set to $3 x 10^3$ Euro day\textsuperscript{-1}.
+The mean prices for the two species were set to 150 euro per ton, ranging between 63 euro per ton for the youngest age and and 238 ton for the oldest age.
+The fines for overshooting the quota was set  to 500 euro per ton.
 These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
 
-We used mean prices of 150 Euro kg\textsuperscript{-1} for both species.
+\subsection{Management procedure}
+\label{sec2.3}
+The fisheries model 
+The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
+These manegement strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
+For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.
 
-Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible in comparison.
-Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
-
-\subsection{Endogenously determined quotas}
-\label{sec2.3}
-
-The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
-These manegemet strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
-For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.
+\begin{figure}[!ht]
+\centering
+\label{f:distributions}
+\includegraphics[width=.69\textwidth]{Figures/distributions.eps} 
+\caption{Distribution of the two species over the areas as a function of age in unfished situation}
+\end{figure}
 

### 2018-05-14 doc/MSE_MAR_18.tex
pushed changes in tex and code  (for Nekane)
Nekane will split model and plots

@@
 Number of vessels                                                    &                       & 700           \\ 
-Fuel costs (Euro day\textsuperscript{-1})               &                       & 1240         \\ 
-Gear maintenance (Euro day\textsuperscript{-1})  &                       & 0               \\ 
+Fuel costs (Euro fishing season\textsuperscript{-1})               &                       & 100        \\ 
+Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 50               \\ 
 Crew share                                                               &                       & 35\%        \\ 
@@
 The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
-These manegement strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
+These management strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
 For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.

### 2018-05-14 complex model/complexmodel_DEC.R

@@ -5,7 +5,7 @@ library(reshape2)
 library(plyr)
 library(scales)
 
-setwd("~/Dropbox/BoB/MSE/Git/Nekane/complex model/")
+setwd("~/Nekane_MSE/complex model/")
 source("functions.R")
 
 
@@ -121,9 +121,9 @@ ages          <- 1:4
 season        <- 1:6
 areas         <- c("a", "b")
 stab.model    <- 10
-NUMRUNS       <- 80
-MPstart       <- 40
-MPstartLO     <- 65
+NUMRUNS       <- 40
+MPstart       <- 20
+MPstartLO     <- 30
 SIMNUMBER     <- 700 #pos
 SIGMA         <- 40 #sig 
 SPP1DSCSTEPS  <- 1
@@ -184,6 +184,31 @@ pos_catches1<- pos_catches2 <- pop1
 pop1 <- population_dynamics(pop=pop1, startyear=2, endyear=stab.model, season=season, natmortality=natmortality, catches=catches.n.dsvm1[,1,,, drop=F], recruitment=recs1, migration=mig1)
 pop2 <- population_dynamics(pop=pop2, startyear=2, endyear=stab.model, season=season, natmortality=natmortality, catches=catches.n.dsvm1[,1,,, drop=F], recruitment=recs2, migration=mig2)
 
+setEPS()
+postscript("distributions.eps")
+par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
+split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
+split.screen(c(2,1), screen=1)-> ind
+screen( ind[1])
+
+image(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+
+screen( ind[2])
+
+image(matrix(round(aperm(pop2[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+
[…]

### 2018-05-15 doc/previous versions/MSE_JAN_18.tex
splitted code and plots

--- 2017-12-20 doc/previous versions/MSE_DEC.tex
@@
 \resizebox{0.5\textwidth}{!}{%
-\begin{tabular}{@{}lcccc@{}} 
+\begin{tabular}{@{}llcccc@{}} 
 \hline
 \textbf{Market value (Euro kg\textsuperscript{-1})}& \multicolumn{4}{c}{age} \\ 
-                         & 1     & 2     & 3    & 4   \\ \hline
+&                         & 1     & 2     & 3    & 4   \\ \hline
                       \\
-species 1                &  778  & 1354   & 1780  & 2096  \\
-species 2                &  778  & 1354   & 1780  & 2096  \\
+Scenario I &              &       &        &       &     \\                       
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  778  & 1354   & 1780  & 2096  \\
+                      \\
+Scenario II &              &       &        &       &     \\ 
+&species 1                &  778  & 1354   & 1780  & 2096  \\
+&species 2                &  778  & 1354   & 1780  & 2096  \\
 \\
@@
 \label{scenarios}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{l c c  c }
+\resizebox{0.8\textwidth}{!}{%
+\begin{tabular}{l c c}
 \hline
-                                & \textbf{Scenario I}            & \textbf{Scenario II}                 & \textbf{Scenario III}           \\ 
-                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)} & \textit{(LO not fully enforced)} \\ \hline
-& & &  \\
-Discarding                      & Allowed                        & Not allowed                          & Ocurred but not perceived \\
-Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)                      &  $C = L$ ($D \neq 0$) \\
-Selectivity ($l_{ratio}$)         &  $ \neq 1$                     & $=1$                                 & $=1$\\
-Yield per recruitment model     & based in $F_L$                 &       based in $F_C$                 &  based in $F_L$  \\
+                                & \textbf{Scenario I}            & \textbf{Scenario II}                  \\ 
+                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
+& &   \\
+Discarding                      & Allowed                        & Not allowed                 \\
+Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
+Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
+Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
 
-&  & &  \\\hline
+&  &   \\\hline
 \end{tabular}
@@
 %----------------------------------------------------------------------------------------------------------------------------------------
-\begin{figure}[!h]
-\centering
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO.eps} 
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO.eps} 
-\end{minipage}
-\caption{}
-\end{figure}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
-
-Scenario III:
-
-\begin{figure}[!h]
-\centering
-\begin{minipage}{.45\textwidth}
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_LO_nocompliance.eps}  
-\end{minipage}
-\hfill
-\begin{minipage}{.45\textwidth}
-\centering
-\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_LO_nocompliance.eps} 
-\end{minipage}
-\caption{}
-\label{fig:effort_lo}
-\end{figure}
-
-%----------------------------------------------------------------------------------------------------------------------------------------
+%\begin{figure}[!h]
+%\centering
+%\begin{minipage}{.45\textwidth}
+%\includegraphics[width=0.9\textheight, angle=90]{Figures/Catch_scenarioII.eps} 
+%\end{minipage}
+%\hfill
+%\begin{minipage}{.45\textwidth}
+%\includegraphics[width=0.9\textheight, angle=90]{Figures/Effort_scenarioII.eps} 
+%\end{minipage}
+%\caption{}
+%\end{figure}
 
@@
 \label{sec3.1}
+When discarding was allowed, but TACs changed by 15\% based on previous year but trying approximate to Fmsy, TACS showed a reduction and discards increased.
+
+The implementation of LO caused an increase in TACs and revenues were higher than when discarding was allowed.
 

### 2018-05-15 complex model/complexmodel_MAY.R

@@ -1,0 +1,428 @@
+options(width=200)
+library(RDynState5NAsigmaseason6Age)
+library(ggplot2)
+library(reshape2)
+library(plyr)
+library(scales)
+
+setwd("~/Nekane_MSE/complex model/")
+source("functions.R")
+
+
+##############################################################################
+# POPULATION DYNAMICS
+##############################################################################
+
+population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
+  #pop[age,year, season, area]                                                                                                        
+  MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
+  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
+    for (ss in (1:length(season))){
+      # move time ---------------------                                                             
+      if (ss ==1){
+        for(age in 2:(dim(pop)[1])){
+          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
+        }
+      } else {
+        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
+      }
+      
+      # birth/recruitment ---------------------                                                     
+      if (ss ==1){
+        pop[1,as.character(y),as.character(ss),] <- recruitment;
+      }
+      # natural mortality  ---------------------                                                             
+      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),]*(1-natmortality)
+      pop[pop < 1e-20 ] <- 1e-20
+      
+      # remove catches (dims of catches here is ages,season, area, just like in main )
+      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),] - catches[,,as.character(ss),]
+      pop[pop < 1e-20 ] <- 1e-20 
+      
+      #migration
[…]

### 2018-05-15 complex model/complexmodel_plots.R

@@ -1,0 +1,342 @@
+###############################################################################
+# Figures from results
+##############################################################################
+
+#We need to check if legend is correct (i.e. high and low abundances are correctly coloured for the two species.
+setwd("D://Nekane_MSE/doc/Figures")
+setwd("~/Dropbox/BoB/MSE/Git/Nekane/doc/Figures")
+
+
+setEPS()
+postscript("distributions.eps")
+par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
+split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
+split.screen(c(2,1), screen=1)-> ind
+screen( ind[1])
+
+image(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+
+screen( ind[2])
+
+image(matrix(round(aperm(pop2[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+
+#image.plot(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64),  main= "Species 1" )
+screen(2)
+image.plot(legend.only=T,zlim=c(0,400),  col=gray((0:64)/64),  smallplot=c(.2,.4, .3,.7))
+close.screen( all=TRUE)
+dev.off()
+
+
+setEPS()
+postscript("prices.eps")
+
+par(mar = c(5,5,2,5))
+plot(x=seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)),y=
[…]

### 2018-05-15 complex model/complexmodel_MAY.R
splitted code and plots2

@@ -125,7 +125,7 @@ NUMRUNS       <- 40
 MPstart       <- 20
 MPstartLO     <- 30
 SIMNUMBER     <- 700 #pos
-SIGMA         <- 40 #sig 
+SIGMA         <- 20 #sig 
 SPP1DSCSTEPS  <- 1
 SPP2DSCSTEPS  <- 1
 endy          <- stab.model + NUMRUNS
@@ -176,38 +176,13 @@ catches.wt.dsvm1      <- catches.wt.dsvm2      <- array(0, dim=c(length(ages),en
 landings.wt.dsvm1     <- landings.wt.dsvm2     <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
 catches.wt.dsvm.tot1  <- catches.wt.dsvm.tot2  <- array(0, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
 landings.wt.dsvm.tot1 <- landings.wt.dsvm.tot2 <- array(0, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
-quota1                <- quota2                <- array(1.2, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
+quota1                <- quota2                <- array(1.6, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
 
 pos_catches1<- pos_catches2 <- pop1
 
 #run population for 15 year
 pop1 <- population_dynamics(pop=pop1, startyear=2, endyear=stab.model, season=season, natmortality=natmortality, catches=catches.n.dsvm1[,1,,, drop=F], recruitment=recs1, migration=mig1)
 pop2 <- population_dynamics(pop=pop2, startyear=2, endyear=stab.model, season=season, natmortality=natmortality, catches=catches.n.dsvm1[,1,,, drop=F], recruitment=recs2, migration=mig2)
-
-setEPS()
[…]

### 2018-05-15 complex model/complexmodel_MAY.R
splitted code and plots3

@@ -124,11 +124,13 @@ stab.model    <- 10
 NUMRUNS       <- 40
 MPstart       <- 20
 MPstartLO     <- 30
+
 SIMNUMBER     <- 700 #pos
 SIGMA         <- 20 #sig 
 SPP1DSCSTEPS  <- 1
 SPP2DSCSTEPS  <- 1
 endy          <- stab.model + NUMRUNS
+
 Linf          <- 50
 K             <- 0.4
 alpha         <- 0.00005
@@ -145,6 +147,8 @@ migconstant   <- 0.2
 sp1price      <- sp2price      <- 150
 slope1price <- 100
 slope2price <- 100 # 0.50*150
+
+FuelPrice   <- 100
 
 # scenario I: discarding is not allowed, YPR based in C (C=L)
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
@@ -188,7 +192,7 @@ pop2 <- population_dynamics(pop=pop2, startyear=2, endyear=stab.model, season=se
 for (ii in 1:endy){
   for(jj in areas){
     pos_catches1[,as.character(ii),,as.character(jj)] <- pop1[,as.character(ii),,as.character(jj)] *q*wts[,1,,1]
-    pos_catches2[,as.character(ii),,as.character(jj)]<- pop2[,as.character(ii),,as.character(jj)] *q*wts[,1,,1]
+    pos_catches2[,as.character(ii),,as.character(jj)] <- pop2[,as.character(ii),,as.character(jj)] *q*wts[,1,,1]
   }
 }
 
@@ -202,7 +206,7 @@ sp1Price <- array(c(sp1price + slope1price*(((wts-mean(wts))/mean(wts)))), dim=c
 sp2Price <- array(c(sp2price + slope2price*(((wts-mean(wts))/mean(wts)))), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 sp3Price <- sp4Price <- sp5Price <- array(c(0), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 #---effort and prices used (note that now c is removed (but that if other runs, then make sure to fix/remove code that removes "c" option)                                                                                         
[…]

### 2018-05-16 doc/MSE_MAR_18.tex
continued on methods section

@@
 Conversion from length to weight was also equal for the two species (Table \ref{t:fishery}).
-Migration was parameterized so that a gradual diffusion occured between the two areas, euqal to 20 percent of the difference in abundance.
+Migration was parameterized so that a gradual diffusion occurred between the two areas, euqal to 20 percent of the difference in abundance.
 For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas  (Figure \ref{f:distributions}).
-Mortality from natural causes such as predation, diseases, and scenesens was assumed to be negligible, and $M_i$ set to 0.0001.
-Althought such absence of natural mortality impossible in reality, one could argue that the model thus mimiced long-lived species.
+Mortality from natural causes such as predation, diseases, and senescence was assumed to be negligible, and $M_i$ set to 0.0001.
+Although such absence of natural mortality impossible in reality, one could argue that the model thus mimicked long-lived species.
 
-Age-dependent catchability links the population biomasses to the catches in the fishery, and is thus one of the crucial parameters determing the interaction between the two.
+Age-dependent catchability links the population biomasses to the catches in the fishery, and is thus one of the crucial parameters determining the interaction between the two.
 In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
@@
 \label{sec2.3}
-The fisheries model 
+The input to the fisheries model consist of the the expected catch rates and the individual quota that are set for each individual in the fleet.
+These individual quota are set in a management procedure, which mimics the decisions of a management body.
+this management body makes observations on the state of the resources, and the exploitation characteristics of the fishery.
+In the model, the management body is assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages.
+these observations stem from fisheries independent observations, such as surveys with known catchabilities.
+In addition, the management body makes annual observations of the catches, including their age distribution.
+The observed biomass and catches allow for estimation of the annual harvest rates, and the estimation of a yield per recruit curve, that is dependent on fish growth and mortality.
+The yield curve is dome shaped, with a maximum that is called $H_{max}$.
+Fishing at a harvest rate that is equal to $H_{max}$  should lead to maximum sustainable yields.
+The yield per recruit curve omits the potential effect of the feedback between adult biomass and recruitment, which anyway is absent in the population dynamics, that assume a constant recruitment.
+The Harvest Control Rule used by the management body resuslts in annual quotas such that the harvest rate in a year corresponds to $H_{max}$.
+These quotas are then divided equally over the individual vessels in the simulations.
+
+\subsection{Scenarios}
+\label{sec2.4}
+The model is set up in three consecutive time windows.
+In the first window of 10 years, newborns enter the population in the absence of fishing.
+This results in a virgin stock status.
+Then, the fleet of 700 vessels starts fishing in the absence of any fisheries regulations.
+This window lasts xxx years and gradually results in a situation where the stocks are overfished, i.e.
+the harvest rates are larger than $H_{max}$.
+Finally, the management procedure is started where the HCR controls the annual quotas.
+
 The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
@@
 \includegraphics[width=.69\textwidth]{Figures/distributions.eps} 
-\caption{Distribution of the two species over the areas as a function of age in unfished situation}
+\caption{Distribution of the two species over the areas as a function of age when stocks are in virgin stock status }
 \end{figure}
@@
 
-HCRs in the form of quotas were determined endogenously by simple and relatively linear reduction in catch if fishing mortality exceeded Fmsy.
-This required that fishing mortality for each stock must not exceed Fmsy, which was a limit reference point.
-An Optimum Yield (OY) was determined based on a MSY that must prevented overfishing, and that could be reduced from the MSY limit to take into account social, economic and precautionary considerations.
-Therefore, quota was determined by a set of harvest control rule parameters consisting of the exploitaion rate at MSY.
-
-The resulted biomass harvested rate in a year $y$, depended in the manager catch perception (based in Table~\ref{t:fishery}) and the standing biomass:
-\begin{equation}
- H_{a, y, s} =\frac
-                % Nominator
-                {C_{i (numbers)}(a, y, s, p)}
-                % Denominator
-                {C_{i (numbers)}(a, y, s, p) \times B_{i (numbers)}(a, y, s, p)},
-\end{equation}
-
 \begin{equation}
@@
 Through a yield per recruitment model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.
-
-\begin{equation}
- H_{MSY} = YIELD CURVE..., 
-\end{equation}
 

### 2018-06-23 complex model/complexmodel_MAY.R
looking for stable parameters

@@ -122,11 +122,11 @@ season        <- 1:6
 areas         <- c("a", "b")
 stab.model    <- 10
 NUMRUNS       <- 25
-MPstart       <- 30
+MPstart       <- 20
 #MPstartLO     <- 26
 
-SIMNUMBER     <- 400 #pos
-SIGMA         <- 5000 #sig 
+SIMNUMBER     <- 8000 #pos 5000 WORKS
+SIGMA         <- 6500 #sig 
 SPP1DSCSTEPS  <- 0
 SPP2DSCSTEPS  <- 0
 endy          <- stab.model + NUMRUNS
@@ -140,11 +140,11 @@ lens          <- Linf*(1-exp(-K*(sages)))
 wts           <- alpha * lens ^ beta
 wts           <- aperm(wts, c(3,2,1,4))
 
-q             <- 0.001
+q             <- 0.00005
 natmortality  <- 0.0001
 
 migconstant   <- 0.05
-sp1price      <- sp2price      <- 1500
+sp1price      <- sp2price      <- 30000
 slope1price <- 1000
 slope2price <- 1000 # 0.50*150
 
@@ -154,7 +154,7 @@ FuelPrice   <- 0
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
 # scenario III: discarding ocurred but not perceived, YPR based in L, hr wanted based in landings
 
-recs1          <- c(200,0) 
+recs1          <- c(500,0) 
 mig1     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
 mig1[,,,"a","a"] <- -migconstant
 mig1[,,,"b","b"] <- -migconstant
@@ -162,7 +162,7 @@ mig1[,,,"a","b"] <- migconstant
 mig1[,,,"b","a"] <- migconstant
 aperm( mig1,c(1,3,2,4,5))
 
-recs2          <- c(0,200) 
+recs2          <- c(0,500) 
 mig2     <- array(0, dim=c(length(ages),1,length(season),length(areas), length(areas)), dimnames=list(cat=ages,year="all",season=as.character(season), from =areas, to=areas)) 
 mig2[,,,"a","a"] <- -migconstant
 mig2[,,,"b","b"] <- -migconstant
@@ -181,7 +181,7 @@ catches.wt.dsvm1      <- catches.wt.dsvm2      <- array(0, dim=c(length(ages),en
[…]

### 2018-06-23 complex model/complexmodel_MAY.R
working model for 6 ages, 6 seasons
Current setting is for quota on both species

Need to check:
1) difference between yield in YPR curve and model outcomes
2) is the plotted selectivity correct

@@ -205,7 +205,7 @@ sp2Price <- array(c(sp2price + slope2price*(((wts-mean(wts))/mean(wts)))), dim=c
 sp3Price <- sp4Price <- sp5Price <- array(c(0), dim=c(length(ages),length(season)), dimnames=list(cat=ages,season=as.character(season)))
 #---effort and prices used (note that now c is removed (but that if other runs, then make sure to fix/remove code that removes "c" option)                                                                                         
 control     <- DynState.control(spp1LndQuota= quota1[,1,,],  spp2LndQuota=quota2[,1,,], spp1LndQuotaFine= 3e6, spp2LndQuotaFine= 3e6, fuelUse = 1, fuelPrice = FuelPrice, landingCosts= 0,gearMaintenance= 0, 
-                                addNoFishing= TRUE, increments= 25, spp1DiscardSteps= SPP1DSCSTEPS, spp2DiscardSteps= SPP2DSCSTEPS, sigma= SIGMA, simNumber= SIMNUMBER, numThreads= 68, verbose=1)
+                                addNoFishing= TRUE, increments= 25, spp1DiscardSteps= SPP1DSCSTEPS, spp2DiscardSteps= SPP2DSCSTEPS, sigma= SIGMA, simNumber= SIMNUMBER * 0.5 , numThreads= 68, verbose=1)
 
 #this is where our loop starts, after we set up stable population
 for(yy in (stab.model):(stab.model+NUMRUNS)){
@@ -244,7 +244,12 @@ for(yy in (stab.model):(stab.model+NUMRUNS)){
     control@spp1LndQuota <-  quota1[,yy,,]
     control@spp2LndQuota <-  quota2[,yy,,]
   }
-
+  
+  # because we wanted gradual introduction of vessels, we only put all vessels after stabmodel
+  if (yy > stab.model){
+    control@simNumber <-  as.integer(SIMNUMBER)
+  }
+  
   # if we are in MPLO period, then set discardsteps =0 
   #if (yy > MPstartLO){
    # control@spp1DiscardSteps <-  0

### 2018-06-23 complex model/complexmodel_MAY.R
final model with 6 seasons 6 ages
More gradual introduction of fleet

@@ -245,10 +245,11 @@ for(yy in (stab.model):(stab.model+NUMRUNS)){
     control@spp2LndQuota <-  quota2[,yy,,]
   }
   
-  # because we wanted gradual introduction of vessels, we only put all vessels after stabmodel
-  if (yy > stab.model){
-    control@simNumber <-  as.integer(SIMNUMBER)
-  }
+  # because we wanted gradual introduction of vessels, we only put all vessels after stabmodel. First line is not strictly necessary becuase already defined when constructing control 
+  if (yy == (stab.model)  ) control@simNumber <-  as.integer(SIMNUMBER*0.5)
+  if (yy == (stab.model+1)) control@simNumber <-  as.integer(SIMNUMBER*0.75)
+  if (yy >  (stab.model+1)) control@simNumber <-  as.integer(SIMNUMBER)
+          
   
   # if we are in MPLO period, then set discardsteps =0 
   #if (yy > MPstartLO){

### 2018-06-23 complex model/complexmodel_plots.R
changed plots

@@ -3,11 +3,11 @@
 ##############################################################################
 library(fields)
 #We need to check if legend is correct (i.e. high and low abundances are correctly coloured for the two species.
-#setwd("D://Nekane_MSE/doc/Figures")
 #setwd("~/Dropbox/BoB/MSE/Git/Nekane/doc/Figures")
 setwd("~/Nekane_MSE/doc/Figures")
 
-#setEPS()
+#DISTRIBUTIONS
+##############################################################################
 postscript("distributions.eps")
 par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
 split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
@@ -17,14 +17,14 @@ screen( ind[1])
 image(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
 box()
 axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
-axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
 
 screen( ind[2])
 
 image(matrix(round(aperm(pop2[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
 box()
 axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
-axis(1,at= seq(0.,0.75,0.25),labels= seq(1:4))
+axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
 
 #image.plot(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64),  main= "Species 1" )
 screen(2)
@@ -33,7 +33,8 @@ close.screen( all=TRUE)
 dev.off()
 
 
-#setEPS()
+#PRICES
+##############################################################################
 postscript("prices.eps")
 par(mar = c(5,5,2,5))
[…]

### 2018-06-23 complex model/complexmodel_plots.R
changed plots

@@ -133,7 +133,7 @@ points(apply(hr1,c(1,2),mean)[6,], col="red", pch=23)
 lines(hr1wanted[1,,1,1], col="red")    #wanted hr
 lines(apply(hr1,c(2),mean), col="black") #observed hr
 
-legend("topright", inset=.05, legend=c("Age 1","Age 2","Age 3","Age 4",,"Age 5","Age 6"), pch=c(19,19,21,21,23,23), col=c("black","red", "black","red", "black","red"), bty='n', cex=0.8)
+legend("topright", inset=.05, legend=c("Age 1","Age 2","Age 3","Age 4","Age 5","Age 6"), pch=c(19,19,21,21,23,23), col=c("black","red", "black","red", "black","red"), bty='n', cex=0.8)
 
 #YPR plots
 plot(x=yc1noMP$hr, y=yc1noMP$landings, type="l", xlim=xlimYPR, ylim=ylim,xaxs='i', yaxs='i',  xlab="Harvest rate", ylab = "Yield per recruit", panel.first=grid(col = "ivory2"))

### 2018-06-23 RDynState5NAsigmaseason12Age/src/RDynState.cpp
started 12 season model

@@ -1,0 +1,958 @@
+#include <Rdefines.h>
+#include <Rinternals.h>
+#include <R.h>
+
+#include <vector>
+#include <math.h>
+#include <stdlib.h>
+#include <time.h>   
+#include <algorithm>
+#include <numeric>
+#include "matalloc.h"
+#include <omp.h>
+
+using namespace std;
+
+int             outofbounds_int;
+bool            outofbounds_bool;
+double          outofbounds_double;
+
+// 160 by 98  works fine
+#define SPP1CAPACITY     1800    // max number of first choke sp (larger than MAXNOINC * MAXHORIZON *NOSIZES= 25 * 6* 12 = 1800) 
+#define SPP2CAPACITY     1800    // max number of second choke sp 
+#define NOSPEC              5    // number of species used in the analysis
+#define NOSIZES             6    // number of size classes used in the analysis
+#define MAXNOINC           25    // max number of increments
+#define MAXHORIZON         12    // number of seasons to fish (= time)
+
+typedef unsigned UFINT;
+
+typedef float  (*FTYPE)[SPP2CAPACITY]; //matalloc(sizeof(float), (void *)0, 2, kSpp1Cap, kSpp2Cap); /* make pointer called FTYPE for FF0Star and FF1 arrays (that only has states and stores V for best choices) */
+typedef float  (*FCTYPE)[SPP2CAPACITY][3]; // (FCTYPE) matalloc(sizeof(float), (void*) 0,3, kSpp1Capacity, kSpp2Capacity,kNPatch) ;  /* make pointer called FCTYPE for FF0 array (that has states and patch, storing V for all patches/choices)_  */
+typedef float  (*ITYPE)[SPP1CAPACITY][SPP2CAPACITY][3]; //  matalloc(sizeof(float), (void *)0, 4, MAXHORIZON, kSpp1Capacity,kSpp2Capacity,kNPatch);  /* make pointer called ITYPE optimal choice array*/
+typedef double (*PTYPE)[MAXHORIZON][NOSPEC][NOSIZES][MAXNOINC]; /* make pointer called PTYPE for information on statistical distribution prey*/
[…]

### 2018-06-23 complex model/complexmodel_MAY.R

@@ -1,5 +1,5 @@
 options(width=240)
-library(RDynState5NAsigmaseason6Age)
+library(RDynState5NAsigmaseason12Age)
 library(ggplot2)
 library(reshape2)
 library(plyr)
@@ -118,7 +118,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
 #  for (sig in SIGMAS){
 
 ages          <- 1:6
-season        <- 1:6
+season        <- 1:12
 areas         <- c("a", "b")
 stab.model    <- 10
 NUMRUNS       <- 25
@@ -140,7 +140,7 @@ lens          <- Linf*(1-exp(-K*(sages)))
 wts           <- alpha * lens ^ beta
 wts           <- aperm(wts, c(3,2,1,4))
 
-q             <- 0.00005
+q             <- 0.000025
 natmortality  <- 0.0001
 
 migconstant   <- 0.05

### 2018-06-23 complex model/complexmodel_MAY.R
fixed bug where integer division was user rather than division

@@ -295,11 +295,11 @@ for(yy in (stab.model):(stab.model+NUMRUNS)){
   
   #calculate numbers caught from weight caught 
   for(jj in areas){
-    catches.n.dsvm1 [,yy,,as.character(jj)]  <- catches.wt.dsvm1[,yy,,as.character(jj)]%/%wts[,1,,1]
-    catches.n.dsvm2 [,yy,,as.character(jj)]  <- catches.wt.dsvm2[,yy,,as.character(jj)]%/%wts[,1,,1]
+    catches.n.dsvm1 [,yy,,as.character(jj)]  <- catches.wt.dsvm1[,yy,,as.character(jj)] / wts[,1,,1]
+    catches.n.dsvm2 [,yy,,as.character(jj)]  <- catches.wt.dsvm2[,yy,,as.character(jj)] / wts[,1,,1]
     
-    landings.n.dsvm1 [,yy,,as.character(jj)] <- landings.wt.dsvm1[,yy,,as.character(jj)]%/%wts[,1,,1]
-    landings.n.dsvm2 [,yy,,as.character(jj)] <- landings.wt.dsvm2[,yy,,as.character(jj)]%/%wts[,1,,1]
+    landings.n.dsvm1 [,yy,,as.character(jj)] <- landings.wt.dsvm1[,yy,,as.character(jj)] / wts[,1,,1]
+    landings.n.dsvm2 [,yy,,as.character(jj)] <- landings.wt.dsvm2[,yy,,as.character(jj)] / wts[,1,,1]
   }
 
   # calculatae what happens to population based on catches

### 2018-06-23 complex model/complexmodel_MAY.R
working model and plots for 12 season model
sigma may be reduced further

@@ -121,12 +121,12 @@ ages          <- 1:6
 season        <- 1:12
 areas         <- c("a", "b")
 stab.model    <- 10
-NUMRUNS       <- 25
-MPstart       <- 20
+NUMRUNS       <- 30
+MPstart       <- 25
 #MPstartLO     <- 26
 
-SIMNUMBER     <- 8000 #pos 5000 WORKS
-SIGMA         <- 6500 #sig 
+SIMNUMBER     <- 8000 #pos 8000 WORKS
+SIGMA         <- 5000 #sig 6500 WORKS
 SPP1DSCSTEPS  <- 0
 SPP2DSCSTEPS  <- 0
 endy          <- stab.model + NUMRUNS

### 2018-06-23 complex model/complexmodel_plots.R

@@ -102,7 +102,7 @@ postscript(file=paste("MIXEDMP2_CATCH_", paste0("SIGMA ", SIGMA, "; INCREMENTS "
 par(mfrow=c(2,5),oma = c(3,0,0,0) + 0.1, mar = c(4,4,1,1) + 0.1)
 
 ylim <- c(0, ceiling(max(sort(catches.wt.dsvm.tot1,partial=length(catches.wt.dsvm.tot1)-1)[length(catches.wt.dsvm.tot1)-1])/100)*100)
-xlimYPR <- c(0,0.3)
+xlimYPR <- c(0,0.15)
 
 plot(catches.wt.dsvm.tot1, type="l",  xlim=c(stab.model,endy), ylim=ylim, xaxs='i', yaxs='i', xlab= "Year", ylab = "Total catches (weight)", xaxt = "n", panel.first=grid(NA, NULL,col = "ivory2"))
 polygon(x=c(pyrnoMP-2,pyrnoMP+2,pyrnoMP+2,pyrnoMP-2), border=NA, y=c(rep(ylim,each=2)), col="ivory4")

### 2018-06-26 complex model/Figuresforchapter.R
changed plots

@@ -1,0 +1,424 @@
+###############################################################################
+# Figures from results
+##############################################################################
+library(fields)
+require(graphics)
+
+setwd("~/Dropbox/BoB/MSE/Git/Nekane_MSE/doc/Figures")
+
+#DISTRIBUTIONS
+##############################################################################
+postscript(file="Distributions.eps", onefile=FALSE, horizontal=FALSE,width=5,height=4.25)
+load("~/Dropbox/BoB/MSE/Git/Nekane_MSE/complex model/1quota12seasons.RData")
+par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
+split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
+
+split.screen(c(2,1), screen=1)-> ind
+screen( ind[1])
+image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
+
+screen( ind[2])
+image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
+box()
+axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
+axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
+
+#image.plot(matrix(round(aperm(pop1[,5,,],c(2,1,3))), ncol=2), col=gray((0:64)/64),  main= "Species 1" )
+screen(2)
+image.plot(legend.only=T,zlim=c(0,400),  col=gray((0:64)/64),  smallplot=c(.2,.4, .3,.7))
+close.screen( all=TRUE)
+dev.off()
+
+#PRICES
[…]

### 2018-06-26 complex model/complexmodel.R
new script to plot final figures, deleted previous and unused versions of figures and docs

@@ -1,652 +1,0 @@
-options(width=200)
-library(RDynState5NAsigmaseason6Age)
-library(ggplot2)
-library(reshape2)
-library(plyr)
-library(scales)
-
-setwd("~/Dropbox/BoB/MSE/Git/Nekane/complex model/")
-source("functions.R")
-
-
-##############################################################################
-# POPULATION DYNAMICS
-##############################################################################
-
-population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
-  #pop[age,year, season, area]                                                                                                        
-  MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
-  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
-    for (ss in (1:length(season))){
-      # move time ---------------------                                                             
-      if (ss ==1){
-        for(age in 2:(dim(pop)[1])){
-          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
-        }
-      } else {
-        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
-      }
-      
-      # birth/recruitment ---------------------                                                     
-      if (ss ==1){
-        pop[1,as.character(y),as.character(ss),] <- recruitment;
-      }
-      # natural mortality  ---------------------                                                             
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),]*(1-natmortality)
-      pop[pop < 1e-20 ] <- 1e-20
-      
-      # remove catches (dims of catches here is ages,season, area, just like in main )
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),] - catches[,,as.character(ss),]
-      pop[pop < 1e-20 ] <- 1e-20 
-      
-      #migration
-      for (age in (1:dim(pop)[1])){
[…]

### 2018-06-26 complex model/complexmodel_DEC.R

@@ -1,700 +1,0 @@
-options(width=200)
-library(RDynState5NAsigmaseason6Age)
-library(ggplot2)
-library(reshape2)
-library(plyr)
-library(scales)
-
-setwd("~/Nekane_MSE/complex model/")
-source("functions.R")
-
-
-##############################################################################
-# POPULATION DYNAMICS
-##############################################################################
-
-population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
-  #pop[age,year, season, area]                                                                                                        
-  MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
-  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
-    for (ss in (1:length(season))){
-      # move time ---------------------                                                             
-      if (ss ==1){
-        for(age in 2:(dim(pop)[1])){
-          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
-        }
-      } else {
-        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
-      }
-      
-      # birth/recruitment ---------------------                                                     
-      if (ss ==1){
-        pop[1,as.character(y),as.character(ss),] <- recruitment;
-      }
-      # natural mortality  ---------------------                                                             
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),]*(1-natmortality)
-      pop[pop < 1e-20 ] <- 1e-20
-      
-      # remove catches (dims of catches here is ages,season, area, just like in main )
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),] - catches[,,as.character(ss),]
-      pop[pop < 1e-20 ] <- 1e-20 
-      
-      #migration
[…]

### 2018-06-26 complex model/complexmodel_MAY.R

@@ -5,7 +5,7 @@ library(reshape2)
 library(plyr)
 library(scales)
 
-setwd("~/Nekane_MSE/complex model/")
+#setwd("~/Nekane_MSE/complex model/")
 source("functions.R")
 
 
@@ -143,7 +143,7 @@ wts           <- aperm(wts, c(3,2,1,4))
 q             <- 0.000025
 natmortality  <- 0.0001
 
-migconstant   <- 0.05
+migconstant   <- 0.025 # 
 sp1price      <- sp2price      <- 30000
 slope1price <- 1000
 slope2price <- 1000 # 0.50*150
@@ -348,7 +348,8 @@ for(yy in (stab.model):(stab.model+NUMRUNS)){
     # In the transition we assume landing ratio equal 1
     #   if (yy == MPstart){ #MPstartLO){ # landing.ratio are equal to 1 for all years before estimating them
       quota1[,yy+1,,] <- prel.quota1
-      quota2[,yy+1,,] <- prel.quota2
+      #quota2[,yy+1,,] <- prel.quota2
+      
     #}else{
     #  tac.constrained1 <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
     #  tac.constrained2 <- c(quota2[,yy,,]*0.85, quota2[,yy,,]*1.15)
@@ -428,11 +429,34 @@ days[is.na(days)]<- 0
 rownames(days)  <- days$option
 days            <- days[,-1]
 
+# Estimate mean effort days by season in the two study periods. Three year mean before the MP and after the MP
+daysmen        <- subset(dsvm_res_allyrs,(spp %in% "sp1"))
+daysmen        <- subset(daysmen ,(cat %in% 1))
+daysmen $year   <- factor(daysmen $year, levels= 1:endy)
+daysmen $season <- factor(daysmen $season, levels= 1:max(season))
+daysmen         <- aggregate(effort~ spp+cat+option+year+season, FUN=sum, data=daysmen )
+daysmen         <- with(daysmen ,data.frame(year, effort, option, season))
+daysmen <- daysmen [!daysmen $option=="Stay in port",]
+#days      <- dcast(days ,option~year, value.var="effort")
+daysmen[is.na(daysmen)]<- 0
+
+nompdays <- aggregate(effort~ option+season, FUN=mean, data=subset(daysmen,(year %in% c((MPstart-3):MPstart))))
[…]

### 2018-06-26 complex model/complexmodel_OCT.R

@@ -1,623 +1,0 @@
-options(width=200)
-library(RDynState5NAsigmaseason6Age)
-library(ggplot2)
-library(reshape2)
-library(plyr)
-library(scales)
-
-setwd("~/Dropbox/BoB/MSE/Git/Nekane/complex model/")
-source("functions.R")
-
-
-##############################################################################
-# POPULATION DYNAMICS
-##############################################################################
-
-population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
-  #pop[age,year, season, area]                                                                                                        
-  MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
-  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
-    for (ss in (1:length(season))){
-      # move time ---------------------                                                             
-      if (ss ==1){
-        for(age in 2:(dim(pop)[1])){
-          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
-        }
-      } else {
-        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
-      }
-      
-      # birth/recruitment ---------------------                                                     
-      if (ss ==1){
-        pop[1,as.character(y),as.character(ss),] <- recruitment;
-      }
-      # natural mortality  ---------------------                                                             
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),]*(1-natmortality)
-      pop[pop < 1e-20 ] <- 1e-20
-      
-      # remove catches (dims of catches here is ages,season, area, just like in main )
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),] - catches[,,as.character(ss),]
-      pop[pop < 1e-20 ] <- 1e-20 
-      
-      #migration
[…]

### 2018-06-26 complex model/complexmodel_plots.R

@@ -8,7 +8,7 @@ setwd("~/Nekane_MSE/doc/Figures")
 
 #DISTRIBUTIONS
 ##############################################################################
-postscript("distributions.eps")
+postscript("distributions.eps", horiz=TRUE,onefile=FALSE,width=8.5,height=11,paper=letter)
 par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
 split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
 split.screen(c(2,1), screen=1)-> ind

### 2018-06-26 doc/MSE_MAR_18.tex
added new script for ploting, deleted old and unused files and added last figures and doc

@@
 \centering
-\includegraphics[width=0.55\textwidth]{Figures/Areadynamics.eps} 
+\includegraphics[width=0.4\textwidth]{Figures/Areadynamics.eps} 
 \caption{The Operating Model (OM) including the key processes in the dynamics of the fish populations and how it was mapped.}
@@
 
-maximum age                                                            & $A$                    & 4               \\
+maximum age                                                            & $A$                    & 6               \\
 number of areas                                                       &                            & 2               \\
@@
 Length-weight isomorphy factor                              & $\beta_i $         &   3              \\
+Migration                                                   & $D_i $         &   5\%              \\
 &               \\
@@
 \label{f:prices}
-\includegraphics[width=.69\textwidth]{Figures/prices.eps} 
+\includegraphics[width=.69\textwidth]{Figures/Prices.eps} 
 \caption{Lengths and prices of the species}
@@
 \label{f:distributions}
-\includegraphics[width=.69\textwidth]{Figures/distributions.eps} 
+\includegraphics[width=.69\textwidth]{Figures/Distributions.eps} 
 \caption{Distribution of the two species over the areas as a function of age when stocks are in virgin stock status }
@@
 
-\begin{table}[!h]
-\centering
-\caption{}
-\label{scenarios}
-\resizebox{0.8\textwidth}{!}{%
-\begin{tabular}{l c c}
-\hline
-                                & \textbf{MP}            & \textbf{LO}                  \\ 
-                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
-& &   \\
-Discarding                      & Allowed                        & Not allowed                 \\
-Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
-Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
-Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
+%\begin{table}[!h]
+%\centering
+%\caption{}
+%\label{scenarios}
+%\resizebox{0.8\textwidth}{!}{%
+%\begin{tabular}{l c c}
+%\hline
+%                                & \textbf{MP}            & \textbf{LO}                  \\ 
+%                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
+%& &   \\
+%Discarding                      & Allowed                        & Not allowed                 \\
+%Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
+%Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
+%Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
 
-&  &   \\\hline
-\end{tabular}
-}
-\end{table}
+%&  &   \\\hline
+%\end{tabular}
+%}
+%\end{table}
+% 
+%\begin{enumerate}
 
-\begin{enumerate}
-
-\item
-\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
+%\item
+%\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
 Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
 
-\item
-\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
+%\item
+%\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
 
-\end{enumerate}
+%\end{enumerate}
 
@@
 \centering
-\includegraphics[width=.69\textwidth]{Figures/figurecatches.eps} 
-\caption{Caches and harvest rates- Prices of the two species are equal (a truly mixed fishery)}
+\includegraphics[width=\textwidth]{Figures/Catches.eps} 
+\caption{Caches- Prices of the two species are equal (a truly mixed fishery)}
 \end{figure}
@@
 \centering
-\includegraphics[width=.69\textwidth]{Figures/figurecatches2.eps} 
-\caption{Change in selectivity- Prices of the two species are equal (a truly mixed fishery)}
+\includegraphics[width=\textwidth]{Figures/Yields.eps} 
+\caption{Yields- Prices of the two species are equal (a truly mixed fishery)}
 \end{figure}
@@
 \centering
-\includegraphics[width=.82\textwidth]{Figures/figureffort.eps} 
+\includegraphics[width=\textwidth]{Figures/Selectivity.eps} 
+\caption{Changes in selectivity- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Effort.eps} 
 \caption{Effort- Prices of the two species are equal (a truly mixed fishery)}
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/figurerevenue.eps} 
+\includegraphics[width=\textwidth]{Figures/Catchesbyarea.eps} 
+\caption{Caches by area- Prices of the two species are equal (a truly mixed fishery)}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Economics.eps} 
 \caption{Revenues- Prices of the two species are equal (a truly mixed fishery)}

### 2018-06-26 simple model/functions.R

@@ -1,86 +1,0 @@
-
-###############################################################################
-# Extract the results from DSVM
-##############################################################################
-
-extract_dsvm_res <- function(z, control, ages, season){
-  
-  simNumber <-control@simNumber 
-  sp        <- c("sp1","sp2","sp3","sp4","sp5")
-  
-  dsvm_res             <- as.data.frame(rbind(as.matrix(spp1Landings(sim(z))),
-                                              as.matrix(spp2Landings(sim(z))),
-                                              as.matrix(spp3Landings(sim(z))),
-                                              as.matrix(spp4Landings(sim(z))),
-                                              as.matrix(spp5Landings(sim(z)))))      
-  names(dsvm_res)      <- "landings.wt"
-  dsvm_res$discards.wt <- c(rbind(as.matrix(spp1Discards(sim(z))), 
-                                  as.matrix(spp2Discards(sim(z))), 
-                                  as.matrix(spp3Discards(sim(z))), 
-                                  as.matrix(spp4Discards(sim(z))), 
-                                  as.matrix(spp5Discards(sim(z)))))
-  
-  dsvm_res$catch.wt    <- dsvm_res$ landings + dsvm_res$discards
-  dsvm_res$effort      <- rep(rep(as.matrix(effort(sim(z))),each=length(ages)),length(sp))
-  dsvm_res$option      <- rep(rep(as.matrix(choice(sim(z))),each=length(ages)),length(sp))
-  dsvm_res$spp         <- as.factor(c(rep(sp, each=(simNumber*length(ages)*length(season)))))
-  dsvm_res$cat         <- ages
-  dsvm_res$season      <- c(rep(season, each=simNumber*length(ages)))
-  dsvm_res$vessel      <- rep(1:simNumber,each=length(ages))  
-  
-  dsvm_res$option[is.na(dsvm_res$option)] <- "Stay in port"
-  
-  dsvm_res[c(1:4)]     <- lapply(dsvm_res[c(1:4)], function(x) as.numeric(as.character(x)))
-  dsvm_res[c(5:9)]     <- lapply(dsvm_res[c(5:9)], function(x) as.factor(x))
-  
-  is.num               <- sapply(dsvm_res, is.numeric)
-  dsvm_res[is.num]     <- lapply(dsvm_res[is.num], round, 6)
-  
-  # Just focus on sp1 and sp2
[…]

### 2018-06-26 simple model/simplemodel.R

@@ -1,335 +1,0 @@
-options(width=200)
-library(RDynState5NAsigmaseason6Age)
-library(ggplot2)
-library(reshape2)
-library(plyr)
-library(scales)
-
-setwd("~/wur/N/Projecten/MARIFLEET/Nekane/simple\ model/")
-source("functions.R")
-
-
-##############################################################################
-# POPULATION DYNAMICS
-##############################################################################
-
-population_dynamics <- function(pop, startyear, endyear, season, natmortality, catches, recruitment, migration){
-  #pop[age,year, season, area]                                                                                                        
- MigToArea <- array(0, dim=dim(migration), dimnames= dimnames(migration))
-  for (y in (startyear:endyear)){     # need to think this, but maybe is better change last year (y in (endy:(endy +10)))           
-    for (ss in (1:length(season))){
-      # move time ---------------------                                                             
-      if (ss ==1){
-        for(age in 2:(dim(pop)[1])){
-          pop[age,as.character(y),as.character(ss),] <- pop[age-1,as.character(y-1),as.character(length(season)),];
-        }
-      } else {
-        pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss-1),];
-      }
-      
-      # birth/recruitment ---------------------                                                     
-      if (ss ==1){
-        pop[1,as.character(y),as.character(ss),] <- recruitment;
-      }
-      # natural mortality  ---------------------                                                             
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),]*(1-natmortality)
-      pop[pop < 1e-20 ] <- 1e-20
-      
-      # remove catches (dims of catches here is ages,season, area, just like in main )
-      pop[,as.character(y),as.character(ss),] <- pop[,as.character(y),as.character(ss),] - catches[,,as.character(ss),]
[…]

### 2018-06-26 doc/MSE_MAR_18.tex
updated modifications in doc and figures

@@
 \includegraphics[width=\textwidth]{Figures/Catches.eps} 
-\caption{Caches- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled total annual catches (thousand tonnes) for all vessels for both (a,c) species 1 and (b,d) species 2 in relation to the available individual quota (red line).
+In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
+MP year reflects the year where the management plan was introduced}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Yields.eps} 
-\caption{Yields- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled yield per recruitment curves (thousand tonnes) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period (black line: period before the MP; grey line: period after the introduction of the MP).
+(a,c,d) After the introduction of the MP all species are underharvested to reach Fmsy, while in (b) species 2 is overharvested due to the abscence of any quota restriction.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Selectivity.eps} 
-\caption{Changes in selectivity- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled changes in selectivity for both (a,c) species 1 and (b,d) species 2 in relation to the catch decision options made based in the availble individual quota.
+In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
+Black lines: selectivity curve before the MP period; grey line: selectivity curve after the introduction of the MP.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Effort.eps} 
-\caption{Effort- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled spatial allocation of effort per year (\%) when (a) only species 1 has quota limitations and (d) both species are quota limited.
+MP years reflect the year where the management plan was introduced.
+Panels (b-c) and (e-f) are the average monthly effort (days/month).
+Panels (b,e) are based on the 3 years monthly effort average (days) before the introduction of the MP, while (b,e) are the 3 years monthly effort average based on the period after the introduction of the MP.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Catchesbyarea.eps} 
-\caption{Caches by area- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled catches (thousand tonnes/year) by area (blue line: Northern area, red line: Southern area) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period.
+In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
+MP year reflects the year where the management plan was introduced.}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Economics.eps} 
-\caption{Revenues- Prices of the two species are equal (a truly mixed fishery)}
+\caption{Modelled main economics (thousand euro/ year) for the fleet under the management scenarios: (a-c) when the fishery is quota contrained only by species 1, while (d-f) both species constrained the fishery.
+Panels (a,d) and (b,e) shown the average annual net and grooss revenues, respectively, by vessel.
+MP year reflects the year where the management plan was introduced.
+Trade-offs between net revenue (black lline), gross revenue (grey line), fuel cost (blue line) and annual fines (red line) for the fleet are shown in panels (c,f).}
 \end{figure}
@@
 
-\newpage
+\clearpage
+
 \section{Discussion}

### 2018-06-27 doc/MSE_MAR_18.tex
fixed changes in figures

@@
 \section{Introduction}
-\label{sec1}
 
@@
 \section{Methods}
-\label{sec2}
 
@@
 \includegraphics[width=0.55\textwidth]{Figures/MSE.eps} 
-\caption{The Managament Strategy Evaluation framework including the Operating Model (OM) and the Management Procedure (MP).}
+\caption{Conceptual overview of the Managament Strategy Evaluation approach, including the Operating Model (OM) and the Management Procedure (MP) components of the framework.}
 \label{fig:MSE}
@@
 \includegraphics[width=0.4\textwidth]{Figures/Areadynamics.eps} 
-\caption{The Operating Model (OM) including the key processes in the dynamics of the fish populations and how it was mapped.}
+\caption{The Operating Model (OM) component, including the key processes in the dynamics of the fish populations.}
 \label{fig:stockdyn}
@@
 \subsection{Population dynamics}
-\label{sec2.1}
 
-In order to allow the fishery to make spatial and temporal choices, the model needed to be seasonally and spatially explicit.
+To allow the fishery to make spatial and temporal choices, the model needed to be seasonally and spatially explicit.
 The dynamics of the fish stocks were modelled using an age-structured model that was spatially explicit, with seasonal time steps.
@@
 The ages in the model range between age 0 and age $A$, the maximum age in the model.
-the seasons range between season 1 and season $S$, the last season within each year.
+The seasons range between season 1 and season $S$, the last season within each year.
 Individuals were born at age 0, at the start of each year $y$, in season 1.
@@
 
-\subsection{fleet dynamics}
-\label{sec2.2}
+\subsection{Fleet dynamics}
 
@@
 The maximum expected utility between current season $s$ and the end of the year was $V (L_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, S)= \Phi (C_i, Q_i, F_i)$.
-For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$, or to stay in port.While fishing, any combination of the age classes caught of the quota species can be landed, or discarded.
+For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$, or to stay in port.
+While fishing, any combination of the age classes caught of the quota species can be landed, or discarded.
 This discarding was defined by a matrix $d$.
@@
 where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
-Prices of fish were assumed to be dependent over fish age (subsection \ref{sec2.2.1}).
+Prices of fish were assumed to be dependent over fish age (subsection X).
 $G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
@@
 \begin{equation}
-U (C_i, Q_i, F_i, s) = R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]).
+U (C_i, Q_i, F_i, s) = R(p, d, s)- G(p) - C(p) + E_{plaa, d}[V (C_i^\prime, Q_i, F_i, s+1]).
 \end{equation}
@@
 Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}.
-For each year, these forward Monte Carlo Simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area
+For each year, these forward Monte Carlo Simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area.
+The effort allocation component of the operating model provides the link between the management decisions and the biological component of the operating model.
 
@@
 
+new-born individuals                                        & $R_i (p)$           & 500               \\
 maximum age                                                 & $A$                 & 6               \\
-number of areas                                                       &                            & 2               \\
-number of seasons                                                   & $S$                     & 6               \\
+number of areas                                             & $p$                 & 2               \\
+number of seasons                                           & $S$                 & 12               \\
 natural mortality                                           & $M_i$               & 0.0001     \\
 Asymptotic length                                           & $L_{\infty_i}$      & 50            \\
-Growth rate                                                              & $K_i$                 &  0.4          \\
-Length-weight conversion factor                             & $ \alpha_i$        &  0.00005   \\
+Growth rate                                                 & $K_i$               &  0.6         \\
+Length-weight conversion factor                             & $ \alpha_i$         &  0.0002   \\
 Length-weight isomorphy factor                              & $\beta_i $          &   3              \\
-Migration                                                   & $D_i $         &   5\%              \\
+Migration                                                   & $D_i $              &   2.5\%              \\
 &               \\
@@
 
-Number of vessels                                                    &                       & 700           \\ 
-Fuel costs (Euro fishing season\textsuperscript{-1})               &                       & 100        \\ 
-Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 50               \\ 
-Crew share                                                               &                       & 35\%        \\ 
+Number of vessels                                           &                       & 8000           \\ 
+Fuel costs (Euro fishing season\textsuperscript{-1})        &                       & 0        \\ 
+Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
+Crew share                                                  &                       & 0\%        \\ 
 Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
-Optimal choice error                                                 & $\sigma $       & 40             \\
+Optimal choice error                                        & $\sigma $             & 5000            \\
 
@@
 Intial quota                                                &                       & 200 \\%& & \\
-catchability                                                               & $q_i (a)$        & 0.0005        \\ 
+catchability                                                & $q_i (a)$             & 0.000025        \\ 
 effort (p,s)                                                &                       & 1                 \\ 
-Price of species at mean weight                               & $\bar{p}_i$   & 150             \\
-Slope of species price                                              & $\gamma_i$   &  100            \\
+Price of species at mean weight                             & $\bar{p}_i$           & 30000             \\
+Slope of species price                                      & $\gamma_i$            &  1000            \\
 Fine for overshooting quota                                 &  $F_i$                & 3e6              \\
@@
 \subsubsection{Size dependent pricing}
-\label{sec2.2.1}
+
 Prices of fish were assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}.
@@
 
-\begin{figure}[!ht]
-\centering
-\label{f:prices}
-\includegraphics[width=.69\textwidth]{Figures/Prices.eps} 
-\caption{Lengths and prices of the species}
-\end{figure}
-
 \subsection{Case study parameterization}
 
-To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and six fishing seasons per year (Table \ref{t:fishery}).
+To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery}).
 There were two fish species in the model, both of which are caught by the fishery.
-The annual number of newborns is equal for the two species, and arbitrarily set to 400 per year.
-The species differ with respect to their nursery grounds: all individuals of species 1 are born in the southern area, while all individuals of species 2 are born in the northern area.
-The maximum age that any individual can reach was 4 years old.
-During their life, individuals grew in length towards the asymptotic length, which was 50 cm (Figure \ref{f:prices}).
-for both species.
+The annual number of newborns was equal for the two species, and arbitrarily set to 500 per year.
+The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Figure \ref{f:distributions}).
+The maximum age that any individual can reach was 6 years old.
+During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Figure \ref{f:prices}).
 Conversion from length to weight was also equal for the two species (Table \ref{t:fishery}).
-Migration was parameterized so that a gradual diffusion occurred between the two areas, euqal to 20 percent of the difference in abundance.
+Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance.
 For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Figure \ref{f:distributions}).
@@
 Although such absence of natural mortality impossible in reality, one could argue that the model thus mimicked long-lived species.
-
-Age-dependent catchability links the population biomasses to the catches in the fishery, and is thus one of the crucial parameters determining the interaction between the two.
-In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
-
-The mean prices for the two species were set to 150 euro per ton, ranging between 63 euro per ton for the youngest age and and 238 ton for the oldest age.
-The fines for overshooting the quota was set  to 500 euro per ton.
-These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
-
-\subsection{Management procedure}
-\label{sec2.3}
-The input to the fisheries model consist of the the expected catch rates and the individual quota that are set for each individual in the fleet.
-These individual quota are set in a management procedure, which mimics the decisions of a management body.
-this management body makes observations on the state of the resources, and the exploitation characteristics of the fishery.
-In the model, the management body is assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages.
-these observations stem from fisheries independent observations, such as surveys with known catchabilities.
-In addition, the management body makes annual observations of the catches, including their age distribution.
-The observed biomass and catches allow for estimation of the annual harvest rates, and the estimation of a yield per recruit curve, that is dependent on fish growth and mortality.
-The yield curve is dome shaped, with a maximum that is called $H_{max}$.
-Fishing at a harvest rate that is equal to $H_{max}$  should lead to maximum sustainable yields.
-The yield per recruit curve omits the potential effect of the feedback between adult biomass and recruitment, which anyway is absent in the population dynamics, that assume a constant recruitment.
-The Harvest Control Rule used by the management body resuslts in annual quotas such that the harvest rate in a year corresponds to $H_{max}$.
-These quotas are then divided equally over the individual vessels in the simulations.
-
-\subsection{Scenarios}
-\label{sec2.4}
-The model is set up in three consecutive time windows.
-In the first window of 10 years, newborns enter the population in the absence of fishing.
-This results in a virgin stock status.
-Then, the fleet of 700 vessels starts fishing in the absence of any fisheries regulations.
-This window lasts xxx years and gradually results in a situation where the stocks are overfished, i.e.
-the harvest rates are larger than $H_{max}$.
-Finally, the management procedure is started where the HCR controls the annual quotas.
-
-The two management strategies we considered differ by the formulae used to determine exploitation rates (HCR).
-These management strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
-For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.
 
@@
 
-%\begin{table}[!h]
-%\centering
-%\caption{}
-%\label{scenarios}
-%\resizebox{0.8\textwidth}{!}{%
-%\begin{tabular}{l c c}
-%\hline
-%                                & \textbf{MP}            & \textbf{LO}                  \\ 
-%                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
-%& &   \\
-%Discarding                      & Allowed                        & Not allowed                 \\
-%Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
-%Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
-%Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
+Age-dependent catchability linked the population biomasses to the catches in the fishery, and was thus one of the crucial parameters determining the interaction between the two.
+In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
 
-%&  &   \\\hline
-%\end{tabular}
-%}
-%\end{table}
-% 
-%\begin{enumerate}
+The mean prices for the two species were set to 30 thousand euro per ton, ranging between 29 thousand euro per ton for the youngest age and 30.5 thousand ton for the oldest age (Figure \ref{f:prices}).
+The fines for overshooting the quota was set  to 3000 thousand euro per ton (Table \ref{t:fishery}).
+These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
 
-%\item
-%\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
-Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
+\begin{figure}[!ht]
+\centering
+\label{f:prices}
+\includegraphics[width=.69\textwidth]{Figures/Prices.eps} 
+\caption{Lengths and prices of the species}
+\end{figure}
 
-%\item
-%\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
+\subsection{Management procedure}
 
-%\end{enumerate}
+The input to the fisheries model consisted of the the expected catch rates and the individual quota that were set for each individual in the fleet.
+These individual quota were set in a management procedure, which mimicked the decisions of a management body.
+This management body made observations on the state of the resources, and the exploitation characteristics of the fishery.
+In the model, the management body was assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages.
+These observations stem from fisheries independent observations, such as surveys with known catchabilities.
+In addition, the management body made annual observations of the catches, including their age distribution.
+The observed biomass and catches allowed for estimation of the annual harvest rates, and the estimation of a yield per recruit curve, that was dependent on fish growth and mortality.
+The yield curve was dome shaped, with a maximum that is called $H_{max}$.
+Fishing at a harvest rate that is equal to $H_{max}$  should lead to maximum sustainable yields.
+The yield per recruit curve omited the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment.
+The Harvest Control Rule used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{max}$.
+These quotas were then divided equally over the individual vessels in the simulations.
+
+The management model encompassed the harvest control rules (HCR).
+The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures.
+
+% Management actions are rarely implemented without error.
+This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
+when and where harvest is taken) are not accounted for in the HCR.
+By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
 
@@
 
-The management model encompassed the harvest control rules (HCR).
-The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
-Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures (section \ref{sec2.3}).
+\subsection{Scenarios}
 
-% Management actions are rarely implemented without error.
-This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
-when and where harvest is taken) are not accounted for in the HCR.
-By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
+The model was set up in three consecutive time windows.
+In the first window of 10 years, newborns entered the population in the absence of fishing.
+This resulted in a virgin stock status.
+Then, the fleet of 8000 vessels started fishing in the absence of any fisheries regulations.
+This window lasted 15 years and gradually resulted in a situation where the stocks were overfished, i.e.
+the harvest rates were larger than $H_{max}$.
+Finally, the management procedure started where the HCR controls the annual quotas.
+
+The two management strategies we considered differ simply by the
+
+formulae used to determine exploitation rates (HCR).
+These management strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
+For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.
+
+The comparison among these scenarios resulted in the 
+However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
+Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
+Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
+Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013,Ulrich2017}.
+Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
+
+%\begin{table}[!h]
+%\centering
+%\caption{}
+%\label{scenarios}
+%\resizebox{0.8\textwidth}{!}{%
+%\begin{tabular}{l c c}
+%\hline
+%                                & \textbf{MP}            & \textbf{LO}                  \\ 
+%                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
+%& &   \\
+%Discarding                      & Allowed                        & Not allowed                 \\
+%Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
+%Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
+%Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
+
+%&  &   \\\hline
+%\end{tabular}
+%}
+%\end{table}
+% 
+%\begin{enumerate}
+
+%\item
+%\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
+Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
+
+%\item
+%\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
+
+%\end{enumerate}
 
 \section{Results}
-\label{sec3}
 
@@
 \subsection{Catch decisions and location choice}
-\label{sec3.1}
+
 When discarding was allowed, but TACs changed by 15\% based on previous year but trying approximate to Fmsy, TACS showed a reduction and discards increased.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Yields.eps} 
-\caption{Modelled yield per recruitment curves (thousand tonnes) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period (black line: period before the MP; grey line: period after the introduction of the MP).
-(a,c,d) After the introduction of the MP all species are underharvested to reach Fmsy, while in (b) species 2 is overharvested due to the abscence of any quota restriction.}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
 \includegraphics[width=\textwidth]{Figures/Selectivity.eps} 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Effort.eps} 
+\includegraphics[width=\textwidth]{Figures/Yields.eps} 
+\caption{Modelled yield per recruitment curves (thousand tonnes) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period (black line: period before the MP; grey line: period after the introduction of the MP).
+(a,c,d) After the introduction of the MP all species are underharvested to reach Fmsy, while in (b) species 2 is overharvested due to the abscence of any quota restriction.}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Efforteconomics.eps} 
 \caption{Modelled spatial allocation of effort per year (\%) when (a) only species 1 has quota limitations and (d) both species are quota limited.
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Economics.eps} 
+\includegraphics[width=\textwidth]{Figures/Meaneffort.eps} 
+\caption{Modelled spatial allocation of effort per year (\%) when (a) only species 1 has quota limitations and (d) both species are quota limited.
+MP years reflect the year where the management plan was introduced.
+Panels (b-c) and (e-f) are the average monthly effort (days/month).
+Panels (b,e) are based on the 3 years monthly effort average (days) before the introduction of the MP, while (b,e) are the 3 years monthly effort average based on the period after the introduction of the MP.}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=\textwidth]{Figures/Mean_Economics.eps} 
 \caption{Modelled main economics (thousand euro/ year) for the fleet under the management scenarios: (a-c) when the fishery is quota contrained only by species 1, while (d-f) both species constrained the fishery.
@@
 \section{Discussion}
-\label{sec4}
+
 tbd

### 2018-06-27 complex model/Figuresforchapter.R
fixed changes in figures

@@ -15,13 +15,13 @@ split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
 
 split.screen(c(2,1), screen=1)-> ind
 screen( ind[1])
-image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
+image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
 box()
 axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
 axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
 
 screen( ind[2])
-image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2), col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
+image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
 box()
 axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
 axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
@@ -42,10 +42,8 @@ par(new = T)
 plot(x=seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)),
      y=c(lens),ylim=c(0,max(lens)), axes=F, xlab=NA,ylab=NA, type="l",lty=2)
 axis(side = 4)
-mtext(side = 4, line = 3, 'Fish length (cm)')
-legend("topleft",
-       legend=c("Price", "Length"),
-       lty=c(1,2), pch=c(NA, NA), col=c("black", "black"))
+mtext(side = 4, line = 3, 'Fish length (cm)', font=2)
+legend("bottomright", inset=.05, legend=c("Price", "Length"), pch=c(NA, NA), lty=c(1,2), lwd=c(2.5,2.5),col=c("black","black"), bty='n',text.font =2, cex=0.8)
 dev.off()
 
 #CATCHES
@@ -53,8 +51,9 @@ dev.off()
[…]

### 2018-07-03 complex model/Figuresforchapter.R
fixed bug in data frame to array function. Sort function added to solve order problem in seasons

@@ -8,41 +8,47 @@ setwd("~/Dropbox/BoB/MSE/Git/Nekane_MSE/doc/Figures")
 
 #DISTRIBUTIONS
 ##############################################################################
-postscript(file="Distributions.eps", onefile=FALSE, horizontal=FALSE,width=5,height=4.25)
+postscript(file="Distributions.eps", onefile=FALSE, horizontal=FALSE,width=8.5,height=3.25)
 load("~/Dropbox/BoB/MSE/Git/Nekane_MSE/complex model/1quota12seasons.RData")
-par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 3.1, 1.1))
-split.screen( rbind(c(0, .8,0,1), c(.8,1,0,1)))
+par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 1.1, 1.1))
+split.screen( rbind(c(0, .9,0,1), c(.9,1,0,1)))
 
-split.screen(c(2,1), screen=1)-> ind
+split.screen(c(1,2), screen=1)-> ind
 screen( ind[1])
-image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main= "Species 1", ylab="Area" )
+image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main= "", ylab="" )
 box()
+mtext("Species 1", side=1, line = -1.3, adj = 0.99, font=2, cex = 1)
+mtext(text="Area"  ,side=2,line=2     , adj = 0.50, font=2)
+
 axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
 axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
 
 screen( ind[2])
-image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main="Species 2", xlab="Age (years)", ylab="Area")
+image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main="", xlab="", ylab="")
 box()
+mtext("Species 2", side=1, line = -1.3, adj = 0.99, font=2, cex = 1)
[…]

### 2018-07-03 complex model/functions.R

@@ -80,7 +80,7 @@ catch_dataframe_to_array <- function(dsvm_result, ages, season, areas, stock, ca
   names(catch)   <- c("cat", "season","option","data")
   
   catch          <- tapply(catch[,"data"], list(factor(x = catch[,"cat"], levels = unique(catch[,"cat"])), 
-                                                factor(x = catch[,"season"], levels = unique(catch[,"season"])), 
+                                                factor(x = catch[,"season"], levels = sort(unique(catch[,"season"]))), 
                                                 factor(x = catch[,"option"], levels = unique(areas))), sum)
   return(catch)
 }

### 2018-07-03 doc/MSE_MAR_18.tex
changes in figures and added some text in scenarios methodology

@@
 We have to make sure we are clear on the ages of the species: do we start at 0 or at 1?
+Final run: introduce fuel costs, prices with steeper slope, maybe reduce sigma, remove 15\% TAC change in MSE diagram ...what else?
 
@@
 \begin{equation}
-w_i(a,s) =  \alpha * ( L_{\infty_i} * (1-\exp^{(-K * (a+(s/S))}))^{\beta}.
+w_i(a,s) =  \alpha * ( L_{\infty_i} * (1-\exp^{(-K * (a+(s/S)))}))^{\beta}.
 \end{equation}
@@
 To simulate a fleet of individual fishing vessels,we used a Dynamic State Variable Model \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
-The model was used to model location choice, extending the model structure in \cite{Batsleer2015}.
-Each individual vessels had a set of choices, which include the choice to go fishing in a season,  location choice within that season, and the choice to discard one or more species and size classes during a season.
-The model had annual fines for exceeding landings quota as in \cite{Alzorriz2018}).
+The model was used to model location choice, same model structure as in \cite{Alzorriz2018}, but without the choice to discard one or more species and size classes while fishing.
+Each individual vessels had a set of choices, which include the choice to go fishing in a season, location choice within that season.
+The model had annual fines for exceeding landings quota as in \cite{Alzorriz2018}.
 In order to calculate state dependent choices during the year, we started by defining the annual fines for exceeding landings quotas at the end of the year:
 \begin{equation}
-\Phi (L_i, Q_i, F_i)= -\sum_i (\textrm{max}( 0, (L_i - Q_i))* F_i),
+\Phi (C_i, Q_i, F_i)= -\sum_i (\textrm{max}( 0, (C_i - Q_i))* F_i),
 \end{equation}
 
-where $L_i$ was the cumulative annual landings for species $i$ for an individual vessel.
-These cumulative landings defined the state of the individual.
-$Q_i$ was the annual individual quota for landings for the different species.
+where $C_i$ was the cumulative annual catches for species $i$ for an individual vessel.
+These cumulative catches defined the state of the individual.
+$Q_i$ was the annual individual quota for catches for the different species.
 Individual quotas were not transferable.
-$F_i$ was the fine per unit weight for exceeding individual landings quota.
+$F_i$ was the fine per unit weight for exceeding individual catches quota.
 
-The maximum expected utility between current season $s$ and the end of the year was $V (L_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, S)= \Phi (C_i, Q_i, F_i)$.
+The maximum expected utility between current season $s$ and the end of the year was $V (C_i, Q_i, F_i, s)$, and the model started by setting $V (C_i, Q_i, F_i, S)= \Phi (C_i, Q_i, F_i)$.
 For preceding seasons, the expected utility depended on individual choices, and each time step individuals chose to visit fishing area $p$, or to stay in port.
-While fishing, any combination of the age classes caught of the quota species can be landed, or discarded.
-This discarding was defined by a matrix $d$.
-The size of that matrix was defined by the number of species under quota constraints and the number of age classes.
-Each element could take the value 0 (discard) or 1 (keep on board and land).
+While fishing, any combination of the age classes caught of the quota species had to be landed.
 The expected utility for each state and each time step $s$ was calculated backward using stochastic dynamic programming \cite{ClarkandMangel2000}:
 \begin{equation}
-V (C_i, Q_i, F_i, s) = max_{p,d}( R(p, d, s)- G(p) - C(p) + E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1])),
+V (C_i, Q_i, F_i, s) = max_{p}( R(p, s)- G(p) - C(p) + E_{p}[V (C_i^\prime, Q_i, F_i, s+1)]),
 \end{equation}
 
-where $R(p, d, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$, and $d$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
-Prices of fish were assumed to be dependent over fish age (subsection X).
-$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative landings and fish prices.
-The term $E_{p, d}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$, and $d$.
+where $R(p, s)$ was the expected immediate contribution of the gross revenue from the sales of fish in a season resulting from choices $p$ (gross revenues resulted from multiplying catches different age classes of species 1, species 2 and prices).
+Prices of fish were assumed to be dependent over fish age (subsection size dependent pricing).
+$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices.
+The term $E_{p}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$.
 The transition of these states were based on normal distributions of catch rates, using the means and variances for the species, as explained in the model conditioning section, following \cite{Poos2010}.
@@
 \begin{equation}
-U (C_i, Q_i, F_i, s) = R(p, d, s)- G(p) - C(p) + E_{plaa, d}[V (C_i^\prime, Q_i, F_i, s+1]).
+U (C_i, Q_i, F_i, s) = R(p, s)- G(p) - C(p) + E_{p}[V (C_i^\prime, Q_i, F_i, s+1)].
 \end{equation}
@@
 \begin{equation}
-\Delta_{p, d}(C_i, Q_i, F_i, s) =  U^* (C_i, Q_i, F_i, s) - U (C_i, Q_i, F_i, s),
+\Delta_{p}(C_i, Q_i, F_i, s) =  U^* (C_i, Q_i, F_i, s) - U (C_i, Q_i, F_i, s),
 \end{equation}
@@
 \begin{equation} \label{eq:solve}
-P_{p, d}(C_i, Q_i, F_i, s) = \frac
+P_{p}(C_i, Q_i, F_i, s) = \frac
                 % Nominator
-                {e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}}
+                {e^{ -\Delta_{p}(C_i, Q_i, F_i, s)/\sigma}}
                 % Denominator
-                {\sum_p \sum_d e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}},
+                {\sum_p e^{ -\Delta_{p}(C_i, Q_i, F_i, s)/\sigma}},
 \end{equation}
@@
 In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$).
-For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
+For computations, we used $\sigma = 5000$.
+%(noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
 
-The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location and discarding behaviour for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota.
+The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota.
 Further details for this procedure can be found in \cite{Alzorriz2018, Batsleer2016} and \cite{Dowling2011}.
@@
 \begin{equation}
- p_i (a,s) = \bar{ p}_i + \gamma_i \times \frac{w_i (a,s) -\bar{w_i}}{\bar{w_i}},
+ p_i (a,s) = \bar{p}_i + \gamma_i \times \frac{w_i (a,s) -\bar{w_i}}{\bar{w_i}},
 \end{equation}
@@
 The annual number of newborns was equal for the two species, and arbitrarily set to 500 per year.
-The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Figure \ref{f:distributions}).
+The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Figure \ref{f:distribution}).
 The maximum age that any individual can reach was 6 years old.
@@
 Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance.
-For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Figure \ref{f:distributions}).
+For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Figure \ref{f:distribution}).
 Mortality from natural causes such as predation, diseases, and senescence was assumed to be negligible, and $M_i$ set to 0.0001.
@@
 \centering
-\label{f:distributions}
-\includegraphics[width=.69\textwidth]{Figures/Distributions.eps} 
-\caption{Distribution of the two species over the areas as a function of age when stocks are in virgin stock status }
+\includegraphics[width=\textwidth]{Figures/Distributions.eps} 
+\caption{Distribution of the two species over the areas as a function of age when stocks are in virgin stock status.}
+\label{f:distribution}
 \end{figure}
@@
 \centering
+\includegraphics[width=.5\textwidth]{Figures/Prices.eps} 
+\caption{Lengths and prices of the species.}
 \label{f:prices}
-\includegraphics[width=.69\textwidth]{Figures/Prices.eps} 
-\caption{Lengths and prices of the species}
 \end{figure}
@@
 
-The input to the fisheries model consisted of the the expected catch rates and the individual quota that were set for each individual in the fleet.
+The management model encompassed the harvest control rules (HCR).
+The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+The input to the fisheries model consisted of the expected catch rates and the individual quota that were set for each individual in the fleet.
 These individual quota were set in a management procedure, which mimicked the decisions of a management body.
@@
 In addition, the management body made annual observations of the catches, including their age distribution.
-The observed biomass and catches allowed for estimation of the annual harvest rates, and the estimation of a yield per recruit curve, that was dependent on fish growth and mortality.
-The yield curve was dome shaped, with a maximum that is called $H_{max}$.
-Fishing at a harvest rate that is equal to $H_{max}$  should lead to maximum sustainable yields.
+
+The observed biomass and catches allowed for estimation of the annual harvest rates ($H_i$), and the estimation of a yield per recruit curve, that was dependent on fish growth and mortality.
+The yield curve was dome shaped, with a maximum that is called $H_{i,max}$.
+Fishing at a harvest rate that is equal to $H_{i,max}$  should lead to maximum sustainable yields.
 The yield per recruit curve omited the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment.
-The Harvest Control Rule used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{max}$.
+The Harvest Control Rule used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$.
 These quotas were then divided equally over the individual vessels in the simulations.
 
-The management model encompassed the harvest control rules (HCR).
-The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
-Costs and benefits, both economic and in terms of risks to both stock and livelihoods, could be computed and compared across different scenarios and management procedures.
-
-% Management actions are rarely implemented without error.
-This error can come from two main sourrces: (i) resource users do not comply with the regulations, and (ii) the individual dynamics of resource users (e.g.
-when and where harvest is taken) are not accounted for in the HCR.
-By evaluating a range of HCRs against a set of plausible operating models using multiple performance metrics (section \ref{sec2.3}), MSE enables to study the different trade-offs involved with each management procedure.
-
 \begin{equation}
- l_{ratio} (a, y, s) = \frac
+ Q_{i,y+1} = \frac
                 % Nominator
-                {L_{i (numbers)}(a, y, s, p)}
+                {\sum( [\frac{H_{i,max}(a,y,s)}{\bar{H_i(a,s)}} \times H_i(a,s) \times sum( N_i (a, y, s, p))] \times w_i(a,s)) }
                 % Denominator
-                {C_{i (numbers)}(a, y, s, p)},
+                {number\  of\  vessels}.
 \end{equation}
-
-Through a yield per recruitment model, knowing growth, harvest rates and recruitment we can estimate how much to reduce the quota for achieving the MSY.
-
-\begin{equation}
- Q_{MSY} = \frac
-                % Nominator
-                {\sum( [\frac{H_{MSY}}{\bar{H}} \times H (a, s) \times l_{ratio} (a, s) \times B_{(numbers)}(a, s)] \times wts) }
-                % Denominator
-                {number\  of\  vessels}, 
-\end{equation}
-\begin{equation}
- Q (y+1) = \max( \min(Q_{MSY}, Q(y) \times 1.15), Q(y) \times 0.85), 
-\end{equation}
-
-%Given the landing parameters associated
-
-%hr<- catches.n/ (catches.n+ pop)
-%landings.ratio <- landings.n/ cathces.n
-%OY <- 
-%hr_MSY <- OY[OY_landings==max(OY_landingslandings),]_hr
-
-%prel.quota <-  sum(sweep((hr1wanted[1]/mean(hr1[,yy,]))* hr1[,yy,]*landings.ratio1[,yy,]*apply(pop1[,yy+1,,],c(1,2), sum) ,1,wts,"*"))/SIMNUMBER
-%        tac.constrained <- c(quota1[,yy,,]*0.85, quota1[,yy,,]*1.15)
-%        quota1[,yy+1,,] <- max(min(prel.quota, tac.constrained[2]), tac.constrained[1])
-
-%From an economic perspective, a linear increase towards maximum TAC is the optimal strategy under plausible assumptions on market demand for fish and harvesting cost, as shown in Froese et al., 2010.
-The economic rationale is that, beyond Blim, the stock is within safe biological limits, and as it becomes more and more productive, it is able to supply the market to an increasing extent.
-
-%A linear reduction in fishing mortality has been implemented in Australia (DAFF 2007) a nd is common in the USA.
-The differences between a linear decline in fishing mortality and a linear decline in TAC are small, see Figure A2 in Appendix S4.
-
-%estimating the hr Fisheries are managed by a total allowable catch (TAC). A maximum TAC is set for each stock so that the respective target biomass is maintained
-%on average. This maximum TAC may be taken as long as biomass fluctuations remain above Bmsy.
 
@@
 the harvest rates were larger than $H_{max}$.
-Finally, the management procedure started where the HCR controls the annual quotas.
+During this period the fishery was unmanaged, by meaned with unconstrained quota for both species.
+Finally, the management procedure started where the HCR, fishing at harvest rates equal to $H_{i,max}$, controls the annual quotas.
+This last period lasted 15 years.
 
-The two management strategies we considered differ simply by the
+The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
+For this purpose, species 1 was the one to do so in first scenario, while in second scenario both species were controlled by HCR.
+The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objective of the management plan.
 
-formulae used to determine exploitation rates (HCR).
-These management strategies were chosen such that, if  the  management  model  was  the  correct  model  of  reality,  the corresponding  equilibrium  state  of  the  system  would  meet  the  objective  of  the  plan.
-For this purpose, each strategy was implemented after a period where the fishery was unmanaged, by meaned with unconstrained quota for both species.
-
-The comparison among these scenarios resulted in the 
-However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
-Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
-Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
-Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013,Ulrich2017}.
-Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
-
-%\begin{table}[!h]
-%\centering
-%\caption{}
-%\label{scenarios}
-%\resizebox{0.8\textwidth}{!}{%
-%\begin{tabular}{l c c}
-%\hline
-%                                & \textbf{MP}            & \textbf{LO}                  \\ 
-%                                & \textit{(landings selectivity)}& \textit{(fulll avoidance of discards)}\\ \hline
-%& &   \\
-%Discarding                      & Allowed                        & Not allowed                 \\
-%Manager catch perception        & $C = L + D$                    &        $C = L$ ($D = 0$)    \\
-%Selectivity ($l_{ratio}$)       &  $ \neq 1$                     & $=1$                      \\
-%Yield per recruitment model     & based in $F_L$                 &       based in $F_C$          \\
-
-%&  &   \\\hline
-%\end{tabular}
-%}
-%\end{table}
-% 
-%\begin{enumerate}
-
-%\item
-%\textbf{MP Scenario: Landings selectivity},  under this scenario the fishery is under landings selectivity, only landings contribute to yield.
-Fishing mortality accounts for all catches, regardless wether these catches are landed or discarded.
-
-%\item
-%\textbf{LO Scenario: Fulll avoidance of discards}, this scenario contemplates the LO with landings selectivity only and fulll avoidance of discards.
-
-%\end{enumerate}
+%changes in selectivity.  
+%These management scenarios were chosen such that, if the management model was the correct model of reality, the corresponding equilibrium state of the system would meet the objective of the plan.
 
@@
 In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
-MP year reflects the year where the management plan was introduced}
+MP year reflects the year where the management plan was introduced.}
+\label{f:catches}
 \end{figure}
@@
 Black lines: selectivity curve before the MP period; grey line: selectivity curve after the introduction of the MP.}
+\label{f:selectivity}
 \end{figure}
@@
 (a,c,d) After the introduction of the MP all species are underharvested to reach Fmsy, while in (b) species 2 is overharvested due to the abscence of any quota restriction.}
+\label{f:yields}
 \end{figure}
@@
 MP year reflects the year where the management plan was introduced.}
+\label{f:effortecon}
 \end{figure}
@@
 Trade-offs between net revenue (black lline), gross revenue (grey line), fuel cost (blue line) and annual fines (red line) for the fleet are shown in panels (c,f).}
+\label{f:meanecon}
 \end{figure}

### 2018-07-18 doc/MSE_MAR_18.tex
modified slope of prices, reduced sigma and introduced fuel costs

@@
 
-To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas and discarding patterns, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery}).
+To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery}).
 There were two fish species in the model, both of which are caught by the fishery.
@@
 During this period the fishery was unmanaged, by meaned with unconstrained quota for both species.
-Finally, the management procedure started where the HCR, fishing at harvest rates equal to $H_{i,max}$, controls the annual quotas.
+Finally, to assess the effect of the landing obligation (where discarding is no longer permitted) on the objectives of the MSY, the management procedure started where the HCR, fishing at harvest rates equal to $H_{i,max}$, controls the annual quotas.
 This last period lasted 15 years.
@@
 The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
-For this purpose, species 1 was the one to do so in first scenario, while in second scenario both species were controlled by HCR.
-The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objective of the management plan.
+For this purpose, in a single-stock context: species 1 was the one to do so in first scenario, while in a mixed-fishery context: both species were controlled by HCR in second scenario.
+The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objective of the landing obligation regarding the objectives of the CFP, specially the MSY.
 

### 2018-07-18 complex model/Figuresforchapter.R
modified slope of prices, reduced sigma and introduced fuel costs

@@ -284,7 +284,7 @@ abline(v=MPstart-9.5, lty=2)
 box()
 mtext("(a)", side=1, line = -1.3, adj = 0.03, font=2, cex = 1) 
 
-ylim= c(0, ceiling(max(aggregate(value ~ year, FUN=sum, data=netrev)[-c(1:5),])))
+ylim= c(0, ceiling(max(aggregate(value ~ year, FUN=sum, data=grossrev)[-c(1:5),])))
 plot(aggregate(value ~ year, FUN=sum, data=netrev), type="l", ylim=ylim, xlab = "Year", ylab = "Total net revenue", xaxt = "n", panel.first=grid(NA, NULL,col = "ivory2"))
 lines(aggregate(value ~ year, FUN=sum, data=netrev), type="l")
 lines(aggregate(value ~ year, FUN=sum, data=grossrev), type="l", col="gray")
@@ -309,7 +309,7 @@ abline(v=MPstart-9.5, lty=2)
 box()
 mtext("(c)", side=1, line = -1.3, adj = 0.03, font=2, cex = 1)
 
-ylim= c(0, ceiling(max(aggregate(value ~ year, FUN=sum, data=netrev)[-c(1:5),])))
+#ylim= c(0, ceiling(max(aggregate(value ~ year, FUN=sum, data=netrev)[-c(1:5),])))
 plot(aggregate(value ~ year, FUN=sum, data=netrev), type="l", ylim=ylim, xlab = "Year", ylab = "Total net revenue", xaxt = "n", panel.first=grid(NA, NULL,col = "ivory2"))
 lines(aggregate(value ~ year, FUN=sum, data=netrev), type="l")
 lines(aggregate(value ~ year, FUN=sum, data=grossrev), type="l", col="gray")
@@ -323,7 +323,6 @@ mtext("(d)", side=1, line = -1.3, adj = 0.03, font=2, cex = 1)
 legend("bottomright", inset=.05, legend=c("Net revenue","Gross revenue", "Fuel cost", "Annual fine"), pch=c(1,1,1,1), lwd=c(2,2,2,2),col=c("black","gray","blue","red"), bty='n',text.font =2, cex=0.8)
 
 mtext(text="Year",side=1,line=0,font=2,outer=TRUE)
-
 
 dev.off()
 

### 2018-07-18 complex model/complexmodel_MAY.R

@@ -126,7 +126,7 @@ MPstart       <- 25
 #MPstartLO     <- 26
 
 SIMNUMBER     <- 8000 #pos 8000 WORKS
-SIGMA         <- 5000 #sig 6500 WORKS
+SIGMA         <- 4000#5000 #sig 6500 WORKS
 SPP1DSCSTEPS  <- 0
 SPP2DSCSTEPS  <- 0
 endy          <- stab.model + NUMRUNS
@@ -145,10 +145,10 @@ natmortality  <- 0.0001
 
 migconstant   <- 0.025 # 
 sp1price      <- sp2price      <- 30000
-slope1price <- 1000
-slope2price <- 1000 # 0.50*150
+slope1price <- 0.50*sp1price
+slope2price <- 0.50*sp2price#1000 # 0.50*150
 
-FuelPrice   <- 0
+FuelPrice   <- 2000
 
 # scenario I: discarding is not allowed, YPR based in C (C=L)
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
@@ -242,7 +242,7 @@ for(yy in (stab.model):(stab.model+NUMRUNS)){
   # if we are in MP period, then set quota based on last year
   if (yy > MPstart){
     control@spp1LndQuota <-  quota1[,yy,,]
-    control@spp2LndQuota <-  quota2[,yy,,]
+    #control@spp2LndQuota <-  quota2[,yy,,]
   }
   
   # because we wanted gradual introduction of vessels, we only put all vessels after stabmodel. First line is not strictly necessary becuase already defined when constructing control 

### 2018-07-24 doc/MSE_MAR_18.tex
last run results and modified plots

@@
 TBD
+The EU common fisheries policy is designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities.
+the reform includes making use of the MSY reference points as targets for exploiting commercially important fish stocks.
+To incentivise individual business to improve selectivity and avoid unwanted catches, the CFP also introduced a landing obligation.
+and may not necessarily reduce harvest rates but will change the underlying selectivity patterns of harvests.
+Such behaviors Explicitly modelling Using a state-dependent decision-making model combined
 We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses a dynamic state variable model including errors in decision-making in its operating model.}
@@
 \section{Introduction}
-
-We have to make sure we are clear on the ages of the species: do we start at 0 or at 1?
-Final run: introduce fuel costs, prices with steeper slope, maybe reduce sigma, remove 15\% TAC change in MSE diagram ...what else?
+%We have to make sure we are clear on the ages of the species: do we start at 0 or at 1?
 
@@
 These studies may shed light on the potential outcomes of the European fisheries management reform of 2013.
-That reform included using the use of the Maximum Sustainable Yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
+That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
 Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO.
@@
 
-To enable longer term prediction of the effects of changes in fisheries management, and thus complete the Management Strategy Evaluation, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
+To enable longer term prediction of the effects of changes in fisheries management, and thus complete the MSE, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
 We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e.
@@
 
-The Managament Strategy Evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
-The framework consists of an Operating Model (OM) and a Management Procedure (MP).
-In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality.
-(ii) A management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by FMSY reference points.
-(iii) a fleet of individual fishers who aim to make the best use of their annual quota.
-Within this mixed fishery, individual vessels make adaptive choices on fishing location and discarding that depend on the distribution of the resources, and the quota that they are allocated.
+The managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
+The framework consists of an operating model (OM) and a management procedure (MP).
+In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by Hmax reference points; and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.
+Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold.
 Each of these elements is discussed in more detail below.
@@
 
-The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
-The operating model thus includes  individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
-The  management body used the MP to make its decision on how to respond to the state of the resource (Figure \ref{fig:stockdyn}).
-This management procedure thus includes the data collection from the fishery, how these are interpreted, and the Harvest Control Rule that dictates the limits that are set to the fishery.
+The operating model captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
+The OM thus includes individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
+The management body used the MP to make its decision on how to respond to the state of the resource (Fig.\ref{fig:stockdyn}).
+This management procedure thus includes the data collection from the fishery, how these are interpreted, and the harvest control rule (HCR) that dictates the limits that are set to the fishery.
 
@@
 \includegraphics[width=0.55\textwidth]{Figures/MSE.eps} 
-\caption{Conceptual overview of the Managament Strategy Evaluation approach, including the Operating Model (OM) and the Management Procedure (MP) components of the framework.}
+\caption{Conceptual overview of the MSE approach, including the OM and the MP components of the framework.}
 \label{fig:MSE}
@@
 \includegraphics[width=0.4\textwidth]{Figures/Areadynamics.eps} 
-\caption{The Operating Model (OM) component, including the key processes in the dynamics of the fish populations.}
+\caption{The OM component, including the key processes in the dynamics of the fish populations.}
 \label{fig:stockdyn}
@@
 Once the backward calculations were finished, the forward part is a Monte Carlo simulation where the probabilities of choices were sampled randomly using the probabilities in~\ref{eq:solve}.
-For each year, these forward Monte Carlo Simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area.
+For each year, these forward Monte Carlo simulations determine the fishing effort in each season and area $E(y,s,p)$, and the catches $C_i (a, y, s, p)$ for each age in each season and area.
 The effort allocation component of the operating model provides the link between the management decisions and the biological component of the operating model.
@@
 Number of vessels                                           &                       & 8000           \\ 
-Fuel costs (Euro fishing season\textsuperscript{-1})        &                       & 0        \\ 
+Fuel costs (Euro fishing season\textsuperscript{-1})        &                       & 1200        \\ 
 Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
@@
 Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
-Optimal choice error                                        & $\sigma $             & 5000            \\
+Optimal choice error                                        & $\sigma $             & 3000            \\
 
@@
 
-Intial quota                                                &                       & 200 \\%& & \\
+Intial quota  (tonnes)                                      &                       & 200 \\%& & \\
 catchability                                                & $q_i (a)$             & 0.000025        \\ 
 effort (p,s)                                                &                       & 1                 \\ 
-Price of species at mean weight                             & $\bar{p}_i$           & 30000             \\
-Slope of species price                                      & $\gamma_i$            & 1000            \\
-Fine for overshooting quota                                 &  $F_i$                & 3e6              \\
+Price of species at mean weight (Euro tn\textsuperscript{-1})& $\bar{p}_i$          & 30             \\
+Slope of species price                                      & $\gamma_i$            & 15            \\
+Fine for overshooting quota  (Euro tn\textsuperscript{-1})  &  $F_i$                & 3e3              \\
 Ratio of standard deviations to catch means                 & $\eta$                & 0.08           \\ 
@@
 The annual number of newborns was equal for the two species, and arbitrarily set to 500 per year.
-The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Figure \ref{f:distribution}).
+The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Fig.
+\ref{f:distribution}).
 The maximum age that any individual can reach was 6 years old.
-During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Figure \ref{f:prices}).
+During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Fig.
+\ref{f:prices}).
 Conversion from length to weight was also equal for the two species (Table \ref{t:fishery}).
 Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance.
-For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Figure \ref{f:distribution}).
+For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Fig.
+\ref{f:distribution}).
 Mortality from natural causes such as predation, diseases, and senescence was assumed to be negligible, and $M_i$ set to 0.0001.
@@
 \includegraphics[width=\textwidth]{Figures/Distributions.eps} 
-\caption{Distribution of the two species over the areas as a function of age when stocks are in virgin stock status.}
+\caption{Biomass distribution in numbers over the areas as a function of age when stocks are in virgin stock status for species 1 (a) and species 2 (b).
+For species 1 nursery ground occurs in the northern area, while species 2 in the southern area (high biomass in white).
+Over the two areas, initial biomass distribution shows a clear segregation of the younger ages (high and low fish biomass, in white and black respectively), while the older ages show similar distributions (grey colours).}
 \label{f:distribution}
@@
 
-The management model encompassed the harvest control rules (HCR).
-The HCR referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+The management model encompassed the harvest control rules.
+The HCRs referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
 The input to the fisheries model consisted of the expected catch rates and the individual quota that were set for each individual in the fleet.
 These individual quota were set in a management procedure, which mimicked the decisions of a management body.
-This management body made observations on the state of the resources, and the exploitation characteristics of the fishery.
+This management body made observations on the state of thedf resources, and the exploitation characteristics of the fishery.
 In the model, the management body was assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages.
@@
 The yield per recruit curve omited the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment.
-The Harvest Control Rule used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$.
+The HCR used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$.
 These quotas were then divided equally over the individual vessels in the simulations.
@@
 The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
-For this purpose, in a single-stock context: species 1 was the one to do so in first scenario, while in a mixed-fishery context: both species were controlled by HCR in second scenario.
+For this purpose, in a single-species catch quota context: species 1 was the one to do so in first scenario, while in a mixed-fishery catch quota context: both species were controlled by HCR in second scenario.
 The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objective of the landing obligation regarding the objectives of the CFP, specially the MSY.
+To be able to analyse the results of this complex actions of implementing the MP and the LO, we simulated the effects of implementing them without any exceptions or flexibility
 
@@
 \section{Results}
-
-We simulated the effects of implementing a Management Plan (MP) and the Landing Obligation (LO) without any exceptions or flexibility.
-To be able to analyse the results of this complex actions of implementing the MP and the LO, the implementations should start simultaneously for all stocks and fleets.
-Under the MP, the Harvest Control Rule for the TACs is to advise $F_{MSY}$ unless the biomass falls below a trigger biomass.
-If the biomass falls below this value, the advised TAC is based on the biomass trends.
-To check for trends, the latest biomass levels are compared to previous ones.
-If there is a decrease of a 20\% or more, the TAC is reduced by 15\%.
-If the comparision shows an increase of more then 20 \% the TAC is increased by 15\%.
 
@@
 \includegraphics[width=\textwidth]{Figures/Catches.eps} 
-\caption{Modelled total annual catches (thousand tonnes) for all vessels for both (a,c) species 1 and (b,d) species 2 in relation to the available individual quota (red line).
-In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
-MP year reflects the year where the management plan was introduced.}
+\caption{Modelled total annual catches (thousand tonnes) for all vessels and both species: species 1 (a,c) and species 2 (b,d) in relation to the available individual quota (red line).
+In the top panels only species 1 quota constrained the fishery (a,b), while in the bottom panels both species quota constrained the fishery (c,d).
+MP year reflects the year where the management plan was introduced.
+Grey and green areas reflect the pre- and post-manage 5-year period, respectively.}
 \label{f:catches}
@@
 \includegraphics[width=\textwidth]{Figures/Selectivity.eps} 
-\caption{Modelled changes in selectivity for both (a,c) species 1 and (b,d) species 2 in relation to the catch decision options made based in the availble individual quota.
-In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
-Black lines: selectivity curve before the MP period; grey line: selectivity curve after the introduction of the MP.}
+\caption{Modelled changes in harvest rates for both species in relation to the catch decision options made based in the availble individual quota: species 1 (a,c) and species 2 (b,d).
+In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).
+Black lines: mean harvest rates during the pre-manage period (unconstrained fishery); grey line: mean harvest rates during the post-manage period (constrained fishery).
+Periods are relative to the introduction of the management plan (grey and green areas in Fig.
+\ref{f:catches}).}
 \label{f:selectivity}
@@
 \includegraphics[width=\textwidth]{Figures/Yields.eps} 
-\caption{Modelled yield per recruitment curves (thousand tonnes) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period (black line: period before the MP; grey line: period after the introduction of the MP).
-(a,c,d) After the introduction of the MP all species are underharvested to reach Fmsy, while in (b) species 2 is overharvested due to the abscence of any quota restriction.}
+\caption{Modelled yield per recruitment curves (thousand tonnes) for both species in relation to the introduction of the management plan: species 1 (a,c) and species 2 (b,d).
+In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).
+Black lines: expected yield during the pre-manage period (unconstrained fishery); grey line: expected yield during the post-manage period (constrained fishery).
+Periods are relative to the introduction of the management plan (grey and green areas in Fig.
+\ref{f:catches}).
+Blue dots represent the observed catches at the mean harvest rate for each year during the 5-year period (filled dots: pre-manage period and empty dots: post-manage period), while red dots are the expected yields per recruit.
+During the post-manage period all species are underharvested to reach Fmsy (a,c,d), except species 2 in the single-stock case scenario which is overharvested due to the abscence of any quota restriction (b).}
 \label{f:yields}
@@
 \includegraphics[width=\textwidth]{Figures/Efforteconomics.eps} 
-\caption{Modelled spatial allocation of effort per year (\%) when (a) only species 1 has quota limitations and (d) both species are quota limited.
+\caption{Modelled spatial allocation of effort per year (\%) and the respective economic performance when only species 1 has quota limitations (a, b) and both species are quota limited (c,d).
+Trade-offs between net revenue (black line), gross revenue (grey line), fuel cost (blue line) and annual fines (red line) for the fleet are shown in panels (b,d).
 MP years reflect the year where the management plan was introduced.
-Panels (b-c) and (e-f) are the average monthly effort (days/month).
-Panels (b,e) are based on the 3 years monthly effort average (days) before the introduction of the MP, while (b,e) are the 3 years monthly effort average based on the period after the introduction of the MP.}
+In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d)}
 \end{figure}
@@
 In panels (a,b) only species 1 quota constrained the fishery, while in (c,d) both species quota constrained the fishery.
-MP year reflects the year where the management plan was introduced.}
-\label{f:effortecon}
+MP year reflects the year where the management plan was introduced.
+Grey and green areas reflect the pre- and post-manage 5-year period, respectively.}
+\label{f:catchesbyarea}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Meaneffort.eps} 
-\caption{Modelled spatial allocation of effort per year (\%) when (a) only species 1 has quota limitations and (d) both species are quota limited.
-MP years reflect the year where the management plan was introduced.
-Panels (b-c) and (e-f) are the average monthly effort (days/month).
-Panels (b,e) are based on the 3 years monthly effort average (days) before the introduction of the MP, while (b,e) are the 3 years monthly effort average based on the period after the introduction of the MP.}
+\caption{Modelled average spatial allocation of effort (days/month) during the pre-manage (a,c) and post-manage periods (b,d).
+Periods are relative to the introduction of the management plan (grey and green areas in Fig.
+\ref{f:catchesbyarea}).}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{Figures/Mean_Economics.eps} 
-\caption{Modelled main economics (thousand euro/ year) for the fleet under the management scenarios: (a-c) when the fishery is quota contrained only by species 1, while (d-f) both species constrained the fishery.
-Panels (a,d) and (b,e) shown the average annual net and grooss revenues, respectively, by vessel.
-MP year reflects the year where the management plan was introduced.
-Trade-offs between net revenue (black lline), gross revenue (grey line), fuel cost (blue line) and annual fines (red line) for the fleet are shown in panels (c,f).}
+\caption{Modelled main economics by vessel (thousand euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
+MP year reflects the year where the management plan was introduced.}
 \label{f:meanecon}
@@
 
+Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity.
+
+Catch quota management is a more important mechanism than the actual obligation to land. An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.
+
+The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics
 tbd

### 2018-08-01 doc/MSE_MAR_18.tex
results written and main ideas for discussion

@@
 
-\abstract{
-TBD
-The EU common fisheries policy is designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities.
-the reform includes making use of the MSY reference points as targets for exploiting commercially important fish stocks.
-To incentivise individual business to improve selectivity and avoid unwanted catches, the CFP also introduced a landing obligation.
-and may not necessarily reduce harvest rates but will change the underlying selectivity patterns of harvests.
-Such behaviors Explicitly modelling Using a state-dependent decision-making model combined
-We contrast three types of management plans to achieve maximum sustainable yields (MSY) from multiple stocks and compare their effectiveness based on a management strategy evaluation (MSE) that uses a dynamic state variable model including errors in decision-making in its operating model.}
+\abstract{Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY).
+The EU common fisheries policy, designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the MSY reference points as targets for exploiting commercially important fish stocks.
+To incentivise individual business to improve selectivity and avoid unwanted catches, it also introduced a landing obligation.
+Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests.
+The impact of exploiting at MSY and no discarding allowance on the optimal effort allocation between fleets is analyzed.
+We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model.
+When accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
+Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation.
+Optimal effort at MSY generates effort reductions of 37\% to 62\% in the mid-term, thus there would be vessels that could benefit up to 1.5 times larger profits than before the CFP implementation.}
 
@@
 
-\keywords{management stategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield}
+\keywords{management strategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield}
 
@@
 Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.
-To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences.
+To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 Fuel costs (Euro fishing season\textsuperscript{-1})        &                       & 1200        \\ 
-Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
-Crew share                                                  &                       & 0\%        \\ 
-Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
+%Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
+%Crew share                                                  &                       & 0\%        \\ 
+%Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
 Optimal choice error                                        & $\sigma $             & 3000            \\
@@
 effort (p,s)                                                &                       & 1                 \\ 
-Price of species at mean weight (Euro tn\textsuperscript{-1})& $\bar{p}_i$          & 30             \\
+Price of species at mean weight (Euro kg\textsuperscript{-1})& $\bar{p}_i$          & 30             \\
 Slope of species price                                      & $\gamma_i$            & 15            \\
-Fine for overshooting quota  (Euro tn\textsuperscript{-1})  &  $F_i$                & 3e3              \\
+Fine for overshooting quota  (Euro tn\textsuperscript{-1})  &  $F_i$                & 3000              \\
 Ratio of standard deviations to catch means                 & $\eta$                & 0.08           \\ 
@@
 
-\subsection{Case study parameterization}
+\subsection{Case study parametrization}
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Distributions.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Distributions.eps} 
 \caption{Biomass distribution in numbers over the areas as a function of age when stocks are in virgin stock status for species 1 (a) and species 2 (b).
@@
 The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
-For this purpose, in a single-species catch quota context: species 1 was the one to do so in first scenario, while in a mixed-fishery catch quota context: both species were controlled by HCR in second scenario.
-The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objective of the landing obligation regarding the objectives of the CFP, specially the MSY.
-To be able to analyse the results of this complex actions of implementing the MP and the LO, we simulated the effects of implementing them without any exceptions or flexibility
+For this purpose, in a single-species catch quota context: quota on species 1 was was controllef by HCR in first scenario, while in both-species catch quota context: quotas on both species were controlled by HCRs in second scenario.
+The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
+To be able to analyse the results of this complex actions of implementing the MP and the LO, we simulated the effects of implementing them without any exceptions or flexibility.
 
@@
 
-\subsection{Catch decisions and location choice}
+% CATCHES
+Stochastic simulations were carried out for the management scenarios contemplated: moving from an unmanaged period to a situation where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY, while it was not possible to discard.
 
-When discarding was allowed, but TACs changed by 15\% based on previous year but trying approximate to Fmsy, TACS showed a reduction and discards increased.
-
-The implementation of LO caused an increase in TACs and revenues were higher than when discarding was allowed.
-Some fishing opportunities are always lost under the landing obligation; the final fishing mortalities are lower than the target fishing mortalities (at least for one species).
-However, this does not necessarily imply that the catches will be lower with landing obligation than without it.
-Catches will be reduced in the initial years of landing obligation implementation.
-However, at Spawning stock biomass increases with LO, the catches will increase too.
-Although the fishing opportunities are lost due to the choke effect, the biomass of some stocks might increase; after a few years, this extra abundance could be converted into more catches.
-This implies that are fleets that, in the mid-term, could benefit from the landing obligation.
+Figure \ref{f:catches} shows the expected catch of the two target species and the impact of implementing the management scenarios on simulated catch.
+When the fishery was unmanaged, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner (Hmax), thus both species were overexploited, reaching total catches of more than 4 thousand tons per species.
+In both scenarios, the introduction of PM caused an immediate reduction in catches of target species in the short-term.
+However, after such a reduction in the initial years, the stock biomass increases in the mid-term (after a transition of $<$5 years), and therefore mid-term TACs and catches were higher than before the implementation of the MP (Fig.
+\ref{f:catches}, c and d).
+The most notable differences in catch between the different scenarios lied in the absence of a regulatory quota for species 2.
+In the case where the fishery was limited only by HCR for species 1; higher catches for species 1 were observed in the mid-term.
+While species 2 maintained similar catches to those observed in the absence of management.
+As there was no quota restriction on species 2, catches of species 1 were adjusted to the estimated TAC in the HCR (Fig.
+\ref{f:catches}, a and b).
+However, for the fishery that was governed by quotas on both species, differences between catches and quotas were more noticeable due to the choke effect.
+As there was a similar spatial overlap of mature species and an antagonic spatial distribution of young species between the two fishing areas (Fig.
+\ref{f:distribution}), some fishing opportunities would always be lost for keeping the final fishing mortalities of both species below each target fishing mortality.
+The choke effect reduced the effort and made the fleet incapable of fishing their quota share.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Catches.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Catches.eps} 
 \caption{Modelled total annual catches (thousand tonnes) for all vessels and both species: species 1 (a,c) and species 2 (b,d) in relation to the available individual quota (red line).
@@
 
+% HARVEST RATES
+Modelling results indicated that the age composition of the harvest was different during the post-manage than it was during the pre-manage period (Fig.
+\ref{f:selectivity}).
+In the pre-manage period harvesting above the Hmax (with values around 0.03, Fig.
+\ref{f:catches}) caused instability in both populations.
+Overharvesting was caused by more significant harvesting in juvenile specimens than in adults, with a downward sloping harvest rate curves over fish age (rather than selectively targeting older fish with higher price; Fig.4), with average harvest rates of 0.06 (Fig.
+\ref{f:selectivity} and \ref{f:yields}, black lines).
+These harvests and levels of biomass remained stable over time due to the constant recruitment introduced into the population dynamics.
+On the other hand, after the MP implementation, there was a immediate decrease in harvest rates with values equal to or below Hmax.
+After a few years, in the mid-term, 2 to 6-year old fish were harvested at the same rate during the post-manage period, while yearling fish (age class 1) was harvested at lower rates (Fig.
+\ref{f:selectivity} and \ref{f:yields}, grey lines).
+
 \begin{figure}[!ht]
 \centering
-\includegraphics[width=\textwidth]{Figures/Selectivity.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Selectivity.eps} 
 \caption{Modelled changes in harvest rates for both species in relation to the catch decision options made based in the availble individual quota: species 1 (a,c) and species 2 (b,d).
@@
 
+% YIELDS 
+Expected yields as a function of harvest show that during the post-manage period both species were underharvested to reach a sustainable harvesting (Fig.
+\ref{f:yields}; grey lines in a,c,d), except species 2 in the single-stock quota case scenario which was overharvested due to the abscence of any quota restriction (Fig.
+\ref{f:yields}; panel b).
+Indeed, modelling results suggested that higher sustainable yields could potentially be achieved with lower values of fishing mortality.
+When accounting for quotas on both species, MSY yields were 25\% higher than in the unmanaged situation; while accounting just on one specie quota, MSY yield for specie 1 was 50\% higher than the unconstrined fishery value.
+These differences in quota scenarios could be explained due to the already commented choke effect.
+In general, the expected average yields shown similar values than the observed catches during the pre- and post-manage period selected; specially displaying little variability the MSY yields when accounting for quotas on both species scenario (Fig.
+\ref{f:yields}).
+
 \begin{figure}[!ht]
 \centering
-\includegraphics[width=\textwidth]{Figures/Yields.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Yields.eps} 
 \caption{Modelled yield per recruitment curves (thousand tonnes) for both species in relation to the introduction of the management plan: species 1 (a,c) and species 2 (b,d).
@@
 \ref{f:catches}).
-Blue dots represent the observed catches at the mean harvest rate for each year during the 5-year period (filled dots: pre-manage period and empty dots: post-manage period), while red dots are the expected yields per recruit.
-During the post-manage period all species are underharvested to reach Fmsy (a,c,d), except species 2 in the single-stock case scenario which is overharvested due to the abscence of any quota restriction (b).}
+Blue dots represent the observed catches at the mean harvest rate for each year during the 5-year period (filled dots: pre-manage period and empty dots: post-manage period), while red dots are the expected yields at sustainable harvest rates.}
 \label{f:yields}
@@
 
+% EFFORT
+The fishing effort that would lead the fishery toward the biological management reference points can be seen in Figure \ref{f:effort}.
+In an unmanaged fishing situation, given the spatial and price distribution of the species' parameters (Fig.
+\ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable.
+Therefore, 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 million Euros/ year (Fig.
+\ref{f:effort} b and d).
+Fuel costs related to fishing choices were 40\% of the incurred gross revenues (Fig.
+\ref{f:effort} b an d).
+Fishing in the northern (southern) area, during the pre-manage period, showed annual average catches of 3,100 tonnes/ year and 1,200 tonnes/year of species 1 and 2, respectively (1,200 tonnes/year and 3,100 tonnes/ year), which in terms of activity would be fishing around 3100 days per area and season (Fig.
+\ref{f:catchesbyarea} and  \ref{f:meaneffort} a and c).
+Some variability was appreciated between patch choices in the two scenarios (Fig.
+\ref{f:effort} a and c, \ref{f:catchesbyarea} and  \ref{f:meaneffort}); however, gross and net revenues at the end of each year resulted in similar values (Fig\ref{f:effort} b and d).
+
 \begin{figure}[!ht]
 \centering
-\includegraphics[width=\textwidth]{Figures/Efforteconomics.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Efforteconomics.eps} 
 \caption{Modelled spatial allocation of effort per year (\%) and the respective economic performance when only species 1 has quota limitations (a, b) and both species are quota limited (c,d).
@@
 MP years reflect the year where the management plan was introduced.
-In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d)}
+In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).}
+\label{f:effort}
 \end{figure}
+
+When accounting in one species MSY quota, in the short-term, the fleet generated 42.31\% less revenue and a 37.50\% reduction in fishing effort than when compared to the unmanaged situation (Fig.
+\ref{f:effort}).
+At a low TAC of specie 1 in the short-term, the "effort-by-stock" required to catch species 1 TAC became smaller than the effort needed to harvest an unmanaged stock (species 2); thus, depending upon the different exploitation patterns of the fishing grounds.
+Therefore, vessels allocated 80\% (20\%) of the fishing effort in the southern (northern) fishing ground, while maintaining (reducing) both species catches (Fig.
+\ref{f:catchesbyarea}).
+On the other hand, such reductions in effort and changes in exploitation patterns led to an increase in sotck biommass that generated a completely different picture in the mid-term.
+Maintaining the distribution of effort observed in the short-term, in the mid-term, the fleet would generate an average increase revenue of 76.92\% with respect to the pre-managed period, while average fuel cost were reduced to 20\% of the incurred gross revenues (Fig.
+\ref{f:effort} b).
+After the initial 57 \% to 32 \% decline in cathes of species 1 and 2 due to the MP implementation, fishing in the southern (northern) area, during the post-manage period, showed in average 181.67\% and 12.87\% higher catches of species 1 and 2 than the pre-manage period, respectively (16.91\% and 38.12\% lower catches of species 1 and 2)(Fig.
+\ref{f:catchesbyarea} a and b).
+In terms of activity the average number of fishing days per season droped to 3650 days (Fig.
+\ref{f:meaneffort} b).
+Vessels were incentivized to allocate the highest amount of effort early in the year (with around 1075 and 3700 fishing days in the northern and southern fishing grounds respectively), progressively decreasing to lower levels over the course of the year (with around 121 and 1634 fishing days per area; Fig.
+\ref{f:meaneffort} d).
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Catchesbyarea.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Catchesbyarea.eps} 
 \caption{Modelled catches (thousand tonnes/year) by area (blue line: Northern area, red line: Southern area) for both (a,c) species 1 and (b,d) species 2 in relation to the management plan period.
@@
 
+When accounting in MSY quotas on both species, the inclusion of one more MSY quota revealed further reductions in fishing opportunities (Fig.
+\ref{f:effort}).
+In the short-term, the fleet generated 61.54\% less revenue and a 62.5\% reduction in fishing effort than when compared to the unmanaged situation.
+Vessels indistinctly allocated the fishing effort in both fishing grounds (such in the pre-manage period) due to same model conditioning for both species.
+Therefore, after the initial 62 \% to 65 \% decline in cathes due to the MP implementation, fishing in the northern (southern) area, during the post-manage period, somulations showed in average 2.57\% and 54.54\% higher catches of species 1 and 2 than the pre-manage period, respectively (55.46\% and 4.20\%)(Fig.
+\ref{f:catchesbyarea} c and d).
+The limiting species quota incentivized vessels to choose to allocate the highest amount of effort early in the year (with around 2500 fishing days per area), progressively decreasing to lower levels over the course of the year (with around 50 fishing days per area; Fig.
+\ref{f:meaneffort} d).
+In the mid-term, the fleet would generate an average increase revenue of 107.69\% with respect to the pre-managed period.
+
 \begin{figure}[!ht]
 \centering
-\includegraphics[width=\textwidth]{Figures/Meaneffort.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Meaneffort.eps} 
 \caption{Modelled average spatial allocation of effort (days/month) during the pre-manage (a,c) and post-manage periods (b,d).
@@
 \ref{f:catchesbyarea}).}
+\label{f:meaneffort}
 \end{figure}
+
+Modelling results sugested that moving from an unmanaged period to a sustainable stock situation, where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY while it was not possible to discard, moderate and high reductions of fishing effort are expected (Fig.
+\ref{f:effort}, b and d).
+Thus, moderate to high reductions in the number of vessels modelled are expected too (effort cost per vessel 1 day; Fig.
+\ref{f:meaneffort}, b and d).
+The relative annual contribution by vessel to the gross revenue was arouond 28 thousand Euros, reduced to 17 thousand Euros of net revenues due to the costs associated to fuel during the unmanage situation (Fig.
+\ref{f:meanecon}).
+This implies that, in the short-term, revenues under the two scenarios are predicted to result in a loss of profitability (52.89\% when accounting in one species MSY quota and 34.43\% when accounting in b oth MSY quota; Fig.
+\ref{f:meanecon}).
+However, in the mid-term, when compared to the pre-manage period, accounting in one species MSY quota resulted in a 179.16\% increase in revenues per vessel; while accounting in two was 209.38\% increase (Fig.
+\ref{f:meanecon}).
 
@@
 \centering
-\includegraphics[width=\textwidth]{Figures/Mean_Economics.eps} 
+\includegraphics[width=0.8\textwidth]{Figures/Mean_Economics.eps} 
 \caption{Modelled main economics by vessel (thousand euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
@@
 
-\clearpage
-
 \section{Discussion}
 
-Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity.
+\begin{itemize}
+\item{Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity, at least, in the short-term.
+Effort is constrained by some stock, as predicted, in general terms, by Ulrich et al.
+(2011).}
+\item{Drastic reduction of effort would increase profits if MSY and LO were contemplated.
+However, no flexibilities or year transfer swaps were considered.
+Be carefull if applying flexibilities, because as Prellezo 2016 (the results obtained from applying any kind of exemption or flexibility are, simply, negative) said, any increased in mortality beyond the MSY would be negative.}
+\item{Catch quota management is a more important mechanism than the actual obligation to land.
+An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.}
+\item{The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics.}
 
-Catch quota management is a more important mechanism than the actual obligation to land. An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.
+\end{itemize}
 
-The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics
-tbd
-prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative.
+%Although this may seem implausible too, a drastic reduction of effort would increase profits if MSY and LO were contemplated.
+However, no flexibilities or year transfer swaps were consider.
+Be carefull if applying flexibilities, because as Prellezo said, any increased in mortality beyond the MSY would be negative...
+%prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative.
 Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future.
@@
 
-The largest yield (or catch) that can be taken from a species' stock over an indefinite period.
+%The largest yield (or catch) that can be taken from a species' stock over an indefinite period.
 Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely.
@@
 
-There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983).
+%There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983).
 This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource.
-For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis.
+%For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis.
 Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.
 
-As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing.
+%As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing.
 It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere.
@@
 
-The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced.
+%The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced.
 This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g.
@@
 
-The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition.
+%The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition.
 Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points.
@@
 
-In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points.
+%In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points.
 In evaluating investment projects, including resource management, economists discount the future value of any commodity.

### 2018-08-30 doc/MSE_MAR_18.tex
some discussion

@@
 \renewcommand{\theequation}{Eq. \thesubsection.\arabic{equation}} % This line ads "Eq." in front of your equation numbering.
+
+\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
 
@@
 We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model.
-When accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
+The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
 Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation.
-Optimal effort at MSY generates effort reductions of 37\% to 62\% in the mid-term, thus there would be vessels that could benefit up to 1.5 times larger profits than before the CFP implementation.}
+Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.4 times larger profits than before the CFP implementation.
+The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}
 
@@
 Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
-Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013,Ulrich2017}.
+Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013, Ulrich2017}.
 Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
-
+This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.
+However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
+Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
+Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
 To enable longer term prediction of the effects of changes in fisheries management, and thus complete the MSE, one needs to couple the fleet dynamic models that forecast the fleet response to management to  biological dynamic models that forecast the fish population responses to the changing fleet response.
@@
 
-The managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
-The framework consists of an operating model (OM) and a management procedure (MP).
-In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by Hmax reference points; and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.
+Managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
+In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.
 Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold.
 Each of these elements is discussed in more detail below.
-The essential elements of the framework are summarized in Figure \ref{fig:MSE}.
+The essential elements of the framework are summarized in Figure \ref{fig:MSE} and \ref{fig:stockdyn}.
 
-The operating model captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
-The OM thus includes individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance.
-The management body used the MP to make its decision on how to respond to the state of the resource (Fig.\ref{fig:stockdyn}).
-This management procedure thus includes the data collection from the fishery, how these are interpreted, and the harvest control rule (HCR) that dictates the limits that are set to the fishery.
+The framework consists of an operating model (OM) and a management procedure (MP).
+The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
+The OM thus includes individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance (Fig.
+\ref{fig:stockdyn}).
+The management body used the MP to make its decision on how to respond to the state of the resource.
+This MP thus includes the data collection from the fishery, how these are interpreted, and the harvest control rule (HCR) that dictates the limits that are set to the fishery.
 
@@
 Individuals were born at age 0, at the start of each year $y$, in season 1.
-The number of new-born individuals $R_i (p)$ in the model was a function of the area $p$ and independent of the size of the adult population.
+The number of recruits $R_i (p)$ in the model was a function of the area $p$ and independent of the size of the adult population.
 The population numbers at age 0 are thus
@@
 In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$).
-For computations, we used $\sigma = 5000$.
+For computations, we used $\sigma = 3000$ (Table \ref{t:fishery}).
 %(noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
@@
 
-new-born individuals                                        & $R_i (p)$           & 500               \\
+number of recruits                                          & $R_i (p)$           & 500               \\
 maximum age                                                 & $A$                 & 6               \\
 number of areas                                             & $p$                 & 2               \\
-number of seasons                                           & $S$                 & 12               \\
-natural mortality                                           & $M_i$               & 0.0001     \\
+number of months                                            & $S$                 & 12               \\
+natural mortality                                           & $M_i$               & 1\e{-4} \\ %0.0001     \\
 Asymptotic length                                           & $L_{\infty_i}$      & 50            \\
 Growth rate                                                 & $K_i$               & 0.6         \\
-Length-weight conversion factor                             & $ \alpha_i$         &  0.0002   \\
+Length-weight conversion factor                             & $ \alpha_i$         & 2\e{-4} \\% 0.0002   \\
 Length-weight isomorphy factor                              & $\beta_i $          &   3              \\
-Migration                                                   & $D_i $              &   2.5\%              \\
+Migration                                                   & $D_i $              & 2.5\e{-2} \\%  2.5\%              \\
 &               \\
@@
 Number of vessels                                           &                       & 8000           \\ 
-Fuel costs (Euro fishing season\textsuperscript{-1})        &                       & 1200        \\ 
-%Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
-%Crew share                                                  &                       & 0\%        \\ 
-%Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
+Fuel costs (Euro fishing month\textsuperscript{-1})         &                       & 2000        \\ 
+Gear maintenance (Euro fishing season\textsuperscript{-1})  &                       & 0               \\ 
+Crew share                                                  &                       & 0\%        \\ 
+Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
 Optimal choice error                                        & $\sigma $             & 3000            \\
@@
 
-Intial quota  (tonnes)                                      &                       & 200 \\%& & \\
-catchability                                                & $q_i (a)$             & 0.000025        \\ 
+Intial quota  (kg vessel\textsuperscript{-1})               &                       & 200 \\%& & \\
+catchability                                                & $q_i (a)$             & 2.5\e{-5} \\%0.000025        \\ 
 effort (p,s)                                                &                       & 1                 \\ 
@@
 Slope of species price                                      & $\gamma_i$            & 15            \\
-Fine for overshooting quota  (Euro tn\textsuperscript{-1})  &  $F_i$                & 3000              \\
+Fine for overshooting quota  (Euro kg\textsuperscript{-1})  &  $F_i$                & 3000              \\
 Ratio of standard deviations to catch means                 & $\eta$                & 0.08           \\ 
@@
 
-To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery}).
+To mimic spatially heterogeneous fish populations in a mixed fishery where fishers make sequential choices on fishing areas, the model was divided into a 'north' and 'south' area, and twelve fishing seasons per year (Table \ref{t:fishery} and Fig.
+\ref{f:distribution}).
 There were two fish species in the model, both of which are caught by the fishery.
-The annual number of newborns was equal for the two species, and arbitrarily set to 500 per year.
+The annual number of recruits was equal for the two species, and arbitrarily set to 500 per year (Table \ref{t:fishery}).
 The species differed with respect to their nursery grounds: all individuals of species 1 were born in the northern area, while all individuals of species 2 were born in the southern area (Fig.
@@
 The maximum age that any individual can reach was 6 years old.
-During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Fig.
+During their life, individuals grew in length towards the asymptotic length, which was 50 cm for both species (Table \ref{t:fishery}).
+Conversion from length to weight was also equal for the two species (Fig.
 \ref{f:prices}).
-Conversion from length to weight was also equal for the two species (Table \ref{t:fishery}).
-Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance.
+Migration was parameterized so that a gradual diffusion occurred between the two areas, equal to 2.5 percent of the difference in abundance (Table \ref{t:fishery}).
 For the unfished situation this led to a clear segregation of the younger ages over the two areas, while the older ages were equally distributed over the two areas (Fig.
@@
 Age-dependent catchability linked the population biomasses to the catches in the fishery, and was thus one of the crucial parameters determining the interaction between the two.
-In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species.
+In the case study, this age-dependent catchability $q_i (a)$ was assumed independent of age $a$, and equal for the two species (Table \ref{t:fishery}).
 
-The mean prices for the two species were set to 30 thousand euro per ton, ranging between 29 thousand euro per ton for the youngest age and 30.5 thousand ton for the oldest age (Figure \ref{f:prices}).
-The fines for overshooting the quota was set  to 3000 thousand euro per ton (Table \ref{t:fishery}).
+The mean prices for the two species were set to 30 euro per kg, ranging between 17.3 euro per kg for the youngest age and 36.5  euro per kg for the oldest age (Table \ref{t:fishery} and Fig.
+\ref{f:prices}).
+The fines for overshooting the quota was set  to 3,000 euro per kg (Table \ref{t:fishery}).
 These high fines combined with an assumed 100\% detection of exceeding quotas resulted in model results in which fishers comply with quota regulations.
@@
 \includegraphics[width=.5\textwidth]{Figures/Prices.eps} 
-\caption{Lengths and prices of the species.}
+\caption{Lengths and prices of the two species.}
 \label{f:prices}
@@
 The management model encompassed the harvest control rules.
-The HCRs referenced to biological reference points (Fmsy) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
+The HCRs referenced to biological reference points ($H_{max}$) to produce management actions in the form of a harvest or effort level: changes in selectivity or spatial and temporal reallocations or restrictions of fishing effort.
 The input to the fisheries model consisted of the expected catch rates and the individual quota that were set for each individual in the fleet.
 These individual quota were set in a management procedure, which mimicked the decisions of a management body.
-This management body made observations on the state of thedf resources, and the exploitation characteristics of the fishery.
+This management body made observations on the state of the resources, and the exploitation characteristics of the fishery.
 In the model, the management body was assumed to collect annual observations of the biomass of the stocks, including the distribution of the biomass over the different ages.
@@
 The model was set up in three consecutive time windows.
-In the first window of 10 years, newborns entered the population in the absence of fishing.
+In the absence of a plusgroup and 6 ages, the first window of 10 years was enough to get to stable populations.
+During this first period, new young individuals entered the population in the absence of fishing.
 This resulted in a virgin stock status.
-Then, the fleet of 8000 vessels started fishing in the absence of any fisheries regulations.
-This window lasted 15 years and gradually resulted in a situation where the stocks were overfished, i.e.
-the harvest rates were larger than $H_{max}$.
-During this period the fishery was unmanaged, by meaned with unconstrained quota for both species.
-Finally, to assess the effect of the landing obligation (where discarding is no longer permitted) on the objectives of the MSY, the management procedure started where the HCR, fishing at harvest rates equal to $H_{i,max}$, controls the annual quotas.
-This last period lasted 15 years.
+The consecutive 15 years, the fleet of 8000 vessels started fishing in the absence of any fisheries regulations.
+During this time window, the fishery was unmanaged, by meaned with unconstrained quota for both species (pre-manage period).
+During the last 15 years of projections, the management procedure was introduced, the fishery was constrained by setting MSY targets while discarding was not allowed (post-manage period).
 
+Two scenarios based on two management scenarios were examined in the model.
 The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
-For this purpose, in a single-species catch quota context: quota on species 1 was was controllef by HCR in first scenario, while in both-species catch quota context: quotas on both species were controlled by HCRs in second scenario.
+For this purpose, in a single-species catch quota context: quota on species 1 was was controlled by HCR in first scenario, while in both-species catch quota context: quotas on both species were controlled by HCRs in second scenario.
 The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
-To be able to analyse the results of this complex actions of implementing the MP and the LO, we simulated the effects of implementing them without any exceptions or flexibility.
+To be able to analyse the results of this complex actions of implementing the MSY and the LO, we simulated the effects of implementing the LO without any exceptions or flexibility.
+When necessary, for example, in the case of estimating the modelled monthly mean of harvest rates, 5 years windows were selected to characterise the fleet in that pre- and post-manage period for both species (grey and green areas in Fig.
+\ref{f:catches}, respectively).
 
@@
 Figure \ref{f:catches} shows the expected catch of the two target species and the impact of implementing the management scenarios on simulated catch.
-When the fishery was unmanaged, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner (Hmax), thus both species were overexploited, reaching total catches of more than 4 thousand tons per species.
+When the fishery was unmanaged, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$), thus resulted in a situation where the stocks were overfished, reaching total catches of more than 4 tons per species.
 In both scenarios, the introduction of PM caused an immediate reduction in catches of target species in the short-term.
@@
 \includegraphics[width=0.8\textwidth]{Figures/Catches.eps} 
-\caption{Modelled total annual catches (thousand tonnes) for all vessels and both species: species 1 (a,c) and species 2 (b,d) in relation to the available individual quota (red line).
+\caption{Modelled total annual catches (tonnes) for all vessels and both species: species 1 (a,c) and species 2 (b,d) in relation to the available individual quota (red line).
 In the top panels only species 1 quota constrained the fishery (a,b), while in the bottom panels both species quota constrained the fishery (c,d).
@@
 \ref{f:selectivity}).
-In the pre-manage period harvesting above the Hmax (with values around 0.03, Fig.
+In the pre-manage period harvesting above the $H_{max}$ (with values around 0.03, Fig.
 \ref{f:catches}) caused instability in both populations.
@@
 These harvests and levels of biomass remained stable over time due to the constant recruitment introduced into the population dynamics.
-On the other hand, after the MP implementation, there was a immediate decrease in harvest rates with values equal to or below Hmax.
+On the other hand, after the MP implementation, there was a immediate decrease in harvest rates with values equal to or below $H_{max}$.
 After a few years, in the mid-term, 2 to 6-year old fish were harvested at the same rate during the post-manage period, while yearling fish (age class 1) was harvested at lower rates (Fig.
@@
 \ref{f:yields}; panel b).
-Indeed, modelling results suggested that higher sustainable yields could potentially be achieved with lower values of fishing mortality.
-When accounting for quotas on both species, MSY yields were 25\% higher than in the unmanaged situation; while accounting just on one specie quota, MSY yield for specie 1 was 50\% higher than the unconstrined fishery value.
+Indeed, higher sustainable yields achieved with lower values of fishing mortality.
+When quotas on both species, yields were 25\% higher than in the unmanaged situation; just on one specie quota, yield for specie 1 was 50\% higher than the unconstrined fishery value.
 These differences in quota scenarios could be explained due to the already commented choke effect.
@@
 \includegraphics[width=0.8\textwidth]{Figures/Yields.eps} 
-\caption{Modelled yield per recruitment curves (thousand tonnes) for both species in relation to the introduction of the management plan: species 1 (a,c) and species 2 (b,d).
+\caption{Modelled yield per recruitment curves (tonnes) for both species in relation to the introduction of the management plan: species 1 (a,c) and species 2 (b,d).
 In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).
@@
 \ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable.
-Therefore, 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 million Euros/ year (Fig.
+Therefore, 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 thousand Euros/ year (Fig.
 \ref{f:effort} b and d).
@@
 \ref{f:effort} b an d).
-Fishing in the northern (southern) area, during the pre-manage period, showed annual average catches of 3,100 tonnes/ year and 1,200 tonnes/year of species 1 and 2, respectively (1,200 tonnes/year and 3,100 tonnes/ year), which in terms of activity would be fishing around 3100 days per area and season (Fig.
+Fishing in the northern (southern) area, during the pre-manage period, showed annual average catches of 3.1 tonnes/ year and 1.2 tonnes/year of species 1 and 2, respectively (1.2 tonnes/year and 3.1 tonnes/ year), which in terms of activity would be fishing around 3100 days per area and season (Fig.
 \ref{f:catchesbyarea} and  \ref{f:meaneffort} a and c).
@@
 \ref{f:catchesbyarea}).
-On the other hand, such reductions in effort and changes in exploitation patterns led to an increase in sotck biommass that generated a completely different picture in the mid-term.
+On the other hand, such reductions in effort and changes in exploitation patterns led to an increase in stock biommass that generated a completely different picture in the mid-term.
 Maintaining the distribution of effort observed in the short-term, in the mid-term, the fleet would generate an average increase revenue of 76.92\% with respect to the pre-managed period, while average fuel cost were reduced to 20\% of the incurred gross revenues (Fig.
@@
 \ref{f:meaneffort}, b and d).
-The relative annual contribution by vessel to the gross revenue was arouond 28 thousand Euros, reduced to 17 thousand Euros of net revenues due to the costs associated to fuel during the unmanage situation (Fig.
+The relative annual contribution by vessel to the gross revenue was arouond 28 Euros, reduced to 17 Euros of net revenues due to the costs associated to fuel during the unmanage situation (Fig.
 \ref{f:meanecon}).
-This implies that, in the short-term, revenues under the two scenarios are predicted to result in a loss of profitability (52.89\% when accounting in one species MSY quota and 34.43\% when accounting in b oth MSY quota; Fig.
+This implies that, in the short-term, revenues under the two scenarios are predicted to result in a loss of profitability (52.89\% when accounting in one species MSY quota and 34.43\% when accounting in both MSY quota; Fig.
 \ref{f:meanecon}).
-However, in the mid-term, when compared to the pre-manage period, accounting in one species MSY quota resulted in a 179.16\% increase in revenues per vessel; while accounting in two was 209.38\% increase (Fig.
+However, in the mid-term, when compared to the pre-manage period, accounting in one species MSY quota resulted in a 1.32 times larger gross revenues per vessel; while accounting in two species MSY quotas was 1.37 times larger benefits (Fig.
 \ref{f:meanecon}).
@@
 \includegraphics[width=0.8\textwidth]{Figures/Mean_Economics.eps} 
-\caption{Modelled main economics by vessel (thousand euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
+\caption{Modelled main economics by vessel (euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
 MP year reflects the year where the management plan was introduced.}
@@
 \section{Discussion}
+
+Using a managent strategy evaluation framework, the potential consequences of two species quota scenarios were examined, allowing the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
+The MSE framework included temporal and spatial key processes in the dynamics of two fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995, Kell2007}.
+The analysis extended earlier approaches which used short-term state-dependent decision-making models (e.g.
+\cite{Alzorriz2018, Batsleer2016, Poos2010} by including the consequent biomass of fish stocks and the management procedure to make its decision on how to respond to the state of the resource.
+If the landing obligation changes selectivity patterns as predicted in \cite{Alzorriz2018, Batsleer2016}, the use of the present MSE allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling the feedback between fleet and stock dynamics.
+Nevertheless, the framework used made several assumptions: that there was perfect knowledge of the stochastic nature of catch rates, full compliance of fishers with the discarding policy, lack of transferability of individual quotas, and lack of interference competition among fishers \cite{Alzorriz2018}.
+Regarding fish populations, also important assumptions were established: constant mortality and almost neligible, constant recruitment along the years, equal conversion length to weight for the two species, constant migration that lead to a clear segregation of the youngers, age-dependant catchability assumed independent of age and equal for the two species and prices were age dependant but fixed over time.
+Also, we used monthly time steps in the model to keep computation time low, while the time scale at which fishing trips occur may be much smaller.
+Finally, we did not include the existence of any landing obligation discard flexibilities (CFP Article 15, \cite{CFP2013}) or year transfer swaps.
+We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under the landing obligation.
+Despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery fleet and examine whether potential harvest strategies can achieve the management objectives.
+
+When the fishery was unmanaged, with unconstrained quota for both species (pre-manage period), the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$), thus resulted in a situation where the stocks were overfished, reaching total catches of more than 4 tons per species (Fig.
+\ref{f:catches}), which in terms of activity would be fishing around 3100 days per area and season (Fig.
+\ref{f:catchesbyarea} and \ref{f:meaneffort} a and c).
+Given the spatial and price distribution of the species' parameters (Fig.
+\ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable.
+Our results suggested that 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 thousand Euros/ year (Fig.
+\ref{f:effort} b and d).
+Fuel costs related to fishing choices were 40\% of the incurred gross revenues (Fig.
+\ref{f:effort} b an d).
+
+Moving from an unmanaged period to a situation where one or both species were regulated by quotas and forced to comply simulataneously with management objectives (i.e.
+exploiting at MSY and no discarding allowance) on the optimal effort allocation between fleets was analysed.
+For example, the underlying assumption of the two quota scenarios that fleets would switch targeted stocks to their relative abundance and management constraints was tested.
+Under both scenarios, our model results showed that simultaneously implementation of MSY and LO objectives led to a very restrictive fishing activity.
+Effort is constrained by some quota stock, as predicted in general terms by \cite{Ulrich2017}.
+Short-term reductions in effort would be needed while mid-term interactions between fleets and stocks that occur in mixed fisheries such as t
+
+%Consequently, no comparisons could be made to verify the hypothesis stating that an effort regulation is more biologically robust as there are no incentives for discards and black landings.
+This hypothesis would be interesting to investigate in the model as, according to Faroese scientists, it seems to be one of the main advantages of this management system %(Jákupsstovu & Reinert, personal communication).
+
+Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests.
+We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model.
+The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
+Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation.
+Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation.
+The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.
 
@@
 Be carefull if applying flexibilities, because as Prellezo 2016 (the results obtained from applying any kind of exemption or flexibility are, simply, negative) said, any increased in mortality beyond the MSY would be negative.}
+\item{once the status of the stock has been assessed relative to its MSY level, the calculation of a TAC comes automatically from the rules of the Management Procedure.
+Shifting areas toward complying TAC limitations produced changes in selectivity ant therefore, constant changes in MSY levels.
+(e.g.
+shifting selectivity toward smaller/younger fish will generally reduce MSY).
+if it is required that the TAC should correspond to the level where MSY can be obtained, it will be necessary to monitor continuously the stock abundance and to revise TACs on a regular basis.}
+\item{The  MSY under two quota species may be below what was achieved in one quota species because overall selectivity has shifted to smaller fish and low harvest of older fish; the impact of this change in selectivity on estimates of MSY is clearly seen in the results from Fig.
+\ref{f:selectivity} and \ref{f:yields}.
+While  the  potential  MSY  has decreased over time the stock biomass required to produce this MSY has increased.}
 \item{Catch quota management is a more important mechanism than the actual obligation to land.
 An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.}
-\item{The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics.}
+\item{The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics (Catchpole and Riveiro Santos, 2015) .
+Data risks should not be overlooked (Hentati-Sundberg et al., 2014).
+Fish stock assessment provides the foundation for scientific advice to decision-makers on sustainable catch levels.
+A critical feature of stock assessment methodology is its dependence on accurate commercial data, which makes it susceptible to biased reporting (Walters and Maguire, 1996; Cotter et al., 2004; Bousquet et al., 2010).
+The reasons for misreporting and underreporting include e.g.
+IUU fishing (Agnew et al., 2009), discard obligation (Kelleher, 2005; Wielgus et al., 2010) and catches from recreational fisheries (Zeller et al., 2011).
+Such misreporting can be driven by fishing fleet overcapacity, resulting from restrictive TACs and/or effort regulations (Patterson, 1998; Beare et al., 2005; Dobby et al., 2008) that create economic incentives for non-compliance (Sumaila et al., 2006; Gallic and Cox, 2006; Osterblom et al., 2011).}
 
 \end{itemize}
+
+%When discarding was allowed, fishers did not avoid areas where catching undersized specimens was likely because there was no incentive to do so.
+Under all three implementations (scenarios) of the landing obligation, however, our model showed that the Basque trawling fleet was likely to move to areas with lower catch rates of quota species to avoid exceeding quota limitations.
+We showed that in the short term the landing obligation would likely result in a loss in profitability for the fleet.
+This loss in profitability was driven by the ``choke'' effect for species such as mackerel (scenario II and III), which was triggered by reductions in effort as a result of rapid quota exhaustion and by greater costs associated with fishing further from home ports (scenario IV).
+
+%Our results suggest that the Basque fleet is limited by hake, horse mackerel and mackerel quota.
+In scenarios II and III, the choke effect of mackerel is so strong that the effect of other species limits (hake under scenario II, horse mackerel under scenario III) is negligible.
+Consequently, under both these scenarios our model predicts that the fleet will avoid using OTB gear and will reduce fishing effort in sub-area 7 and 8 because of their high mackerel catch rates (Fig.
+\ref{fig:effort_modelval}).
+Thus, fishers will likely visit sub-area 6 and 8 (using PTB gear).
+Revenues under these two scenarios are predicted to result in a loss of profitability (60\% less than the status quo; Table ~\ref{t:net_rev}).
+These findings indicate that trading quotas (at rates undertaken in the past) does not alleviate the choking effect of the landing obligation.
+Therefore, new ways of swapping quotas would be needed under the landing obligation to alleviate the choke effect of any of the quota limiting species.
+Fleets with high discards rates of certain species will be likely to have larger incentives to increase their quotas for those species than fleets with low discard rates.
+
+%In scenario IV, the addition of inter-species quota-swapping flexibility to the landing obligation (e.g.
+9\% of the horse mackerel catch quota could go to increase mackerel quota) substantially decreased the mackerel choke effect and increased both fishing opportunities and net revenues, when compared to scenarios II and III.
+Under this scenario, revenue was 30\% lower than when discarding was allowed (scenario I).
+
+%Our results suggest that reductions in fishing opportunities under a landing obligation that are caused by choke species could be alleviated by the use of inter-species quota flexibility, and that there will be strong incentives to use this flexibility.
+Management bodies should ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation.
+Use of inter-species quota flexibility could cause serious impacts to stock abundance when quota from a high volume/low value species is transferred to a low volume/high value stock.
+Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries.
+
+%The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance.
+With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2005, Batsleer2013}.
+In our model, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations.
+However, if detection rates are low and the short–term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.
+
+%The landing obligation was implemented to incentivise individual businesses to improve selectivity and avoid unwanted catches.
+The landing obligation shifts the emphasis from prescribing the fishing gears that are permissible, to a results-based system where the onus is on the fishers to use fishing gear that will avoid unwanted catches.
+Therefore, fishers should be incentivised to develop alternative gear or adapt their fishing behaviour to mitigate the full impact of the landing obligation \cite{Alzorriz2016}.
+Discard ban policies may not necessarily reduce harvest rates but will change the underlying selectivity patterns of harvests \cite{Borges2016}.
+Our results reflect the short-term effects of a discard ban on selectivity patterns and net revenues implemented under three scenarios.
+If the landing obligation changes selectivity patterns as we predict, then we would also expect longer term changes in both fish populations and the economy of the fishery.
+Such analyses are the domain of bio-economic management strategy evaluations where the feedback between fleet and stock dynamics are modelled explicitly \cite{Andersen2010, Prellezo2016a, Punt2016}.
+In such models, effects of the changing selectivity of the fishing fleets on the reference points for sustainable exploitation can be evaluated.
 

### 2018-09-04 complex model/Figuresforchapter.R
changes in plots

@@ -10,30 +10,31 @@ setwd("~/Dropbox/BoB/MSE/Git/Nekane_MSE/doc/Figures")
 ##############################################################################
 postscript(file="Distributions.eps", onefile=FALSE, horizontal=FALSE,width=8.5,height=3.25)
 load("~/Dropbox/BoB/MSE/Git/Nekane_MSE/complex model/1quota12seasons.RData")
-par(oma=c(0,0,0,0), mar=c(4.1, 4.1, 1.1, 1.1))
+par(oma=c(0,0,0,0), mar=c(4.1, 4.5, 1.1, 1.1))
 split.screen( rbind(c(0, .9,0,1), c(.9,1,0,1)))
 
 split.screen(c(1,2), screen=1)-> ind
 screen( ind[1])
-image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main= "", ylab="" )
+image(matrix(round(aperm(pop1[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:72)/72), axes=F, main= "", ylab="" )
 box()
-mtext("Species 1", side=1, line = -1.3, adj = 0.99, font=2, cex = 1)
-mtext(text="Area"  ,side=2,line=2     , adj = 0.50, font=2)
+mtext("(a)", side=1, line = -1.3, adj = 0.96, font=2, cex = 1)
+mtext(text="Area"  ,side=2,line=3.5    , adj = 0.50, font=2)
 
-axis(2,at= seq(0.,1,1),labels= c("S","N"), las=1)
-axis(1,at= seq(0.,0.833,0.166),labels= seq(1:max(ages)))
+axis(2,at= seq(0.,1,1),labels= c("South","North"), las=1)
+axis(1,at= seq(0.,1,0.1666),labels= seq(1:max(ages+1)))
 
 screen( ind[2])
-image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:64)/64), axes=F, main="", xlab="", ylab="")
+image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:72)/72), axes=F, main="", xlab="", ylab="")
 box()
[…]

### 2018-09-04 complex model/Figuresforchapter.R
new sigma search
increased rec, made weights more realistic and reduced prices.
With these values we need new sigmas

@@ -27,7 +27,7 @@ screen( ind[2])
 image(matrix(round(aperm(pop2[,(stab.model-1),,],c(2,1,3))), ncol=2)[ , c(2,1)], col=gray((0:72)/72), axes=F, main="", xlab="", ylab="")
 box()
 mtext("(b)", side=1, line = -1.3, adj = 0.96, font=2, cex = 1)
-mtext(text="Age (year)"  ,side=1,line=2,, adj = -0.60, font=2)
+mtext(text="Age (year)"  ,side=1,line=2, adj = -0.60, font=2)
 axis(2,at= seq(0.,1,1),labels= c("South","North"), las=1)
 axis(1,at= seq(0.,1,0.1666),labels= seq(1:max(ages+1)))
 
@@ -38,21 +38,29 @@ image.plot(legend.only=T,zlim=c(0,500),  col=gray((0:72)/72), smallplot=c(.15,.3
 close.screen( all=TRUE)
 dev.off()
 
-#PRICES
+#PRICES and FISH LENGTHS
 ##############################################################################
-postscript(file="Prices.eps", onefile=FALSE, horizontal=FALSE,width=4.25,height=3.25)
+postscript(file="weights.eps", onefile=FALSE, horizontal=FALSE,width=4.25,height=3.25)
 par(mar = c(3,5,2,4))
-plot(x=seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)),y=sort(c(sp1Price/1000)), xlab="", ylab= "", ylim=c(0,50), type="l", lty=1, las=1)
+plot(x=seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)),y=sort(c(sp1Price/1000)), xlab="", ylab= "", ylim=c(0,8), type="l", lty=1, las=1)
 mtext(side = 2, line = 3.5, 'Price (euro per kg)', font=2, cex = 1)
 mtext(side = 1, line = 2, 'Age (year)', font=2, cex = 1)
 par(new = T)
 plot(x=seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)),
      y=c(lens),ylim=c(0,max(lens)), axes=F, xlab=NA,ylab=NA, type="l",lty=2)
-axis(side = 4,, las=1)
+axis(side = 4, las=1)
 mtext(side = 4, line = 2.5, 'Fish length (cm)', font=2, cex = 1)
[…]

### 2018-09-04 complex model/complexmodel_MAY.R

@@ -111,11 +111,9 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
   return(res)
 }
 
-#SIGMAS        <- c(10, 40, 70, 100)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
-#NVESSELS      <- c(600, 700, 800)     #SIMNUMBER     <- 600
+SIGMAS        <- seq(5000,100000,5000)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
 
-#for (pos in NVESSELS){
-#  for (sig in SIGMAS){
+for (SIGMA in SIGMAS){
 
 ages          <- 1:6
 season        <- 1:12
@@ -123,10 +121,8 @@ areas         <- c("a", "b")
 stab.model    <- 10
 NUMRUNS       <- 30
 MPstart       <- 25
-#MPstartLO     <- 26
 
 SIMNUMBER     <- 8000 #pos 8000 WORKS
-SIGMA         <- 4000#5000 #sig 6500 WORKS
 SPP1DSCSTEPS  <- 0
 SPP2DSCSTEPS  <- 0
 endy          <- stab.model + NUMRUNS
@@ -134,7 +130,7 @@ endy          <- stab.model + NUMRUNS
 Linf          <- 50 #50
 K             <- 0.6 #0.4
 alpha         <- 0.0002#0.00005
-beta          <- 3
+beta          <- 2.4
 sages         <- array(seq(min(ages)+((1/max(season))/2), max(ages+1),1/max(season)), dim=c(length(season),1,length(ages),1), dimnames=list(season=as.character(season),   year="all", cat=ages, option ="all"))
 lens          <- Linf*(1-exp(-K*(sages)))
 wts           <- alpha * lens ^ beta
@@ -144,7 +140,7 @@ q             <- 0.000025
 natmortality  <- 0.0001
 
 migconstant   <- 0.025 # 
-sp1price      <- sp2price      <- 30000
+sp1price      <- sp2price      <- 5000
 slope1price <- 0.50*sp1price
 slope2price <- 0.50*sp2price#1000 # 0.50*150
 
@@ -154,7 +150,7 @@ FuelPrice   <- 2000
 # scenario II: discarding is allowed, YPR based in L, hr wanted based in catches
 # scenario III: discarding ocurred but not perceived, YPR based in L, hr wanted based in landings
 
-recs1          <- c(500,0) 
[…]

### 2018-09-04 complex model/complexmodel_MAY.R
minor bugfixes after previous commit
increased initial quota and started search at higher sigma

@@ -111,7 +111,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
   return(res)
 }
 
-SIGMAS        <- seq(5000,100000,5000)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
+SIGMAS        <- seq(10000,100000,5000)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
 
 for (SIGMA in SIGMAS){
 
@@ -175,7 +175,7 @@ catches.wt.dsvm1      <- catches.wt.dsvm2      <- array(0, dim=c(length(ages),en
 landings.wt.dsvm1     <- landings.wt.dsvm2     <- array(0, dim=c(length(ages),endy,length(season),length(areas)), dimnames=list(cat=ages,   year=as.character(1:endy), season=as.character(season), option =areas))
 catches.wt.dsvm.tot1  <- catches.wt.dsvm.tot2  <- array(0, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
 landings.wt.dsvm.tot1 <- landings.wt.dsvm.tot2 <- array(0, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
-quota1                <- quota2                <- array(1000, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
+quota1                <- quota2                <- array(5000, dim=c(1           ,endy,              1,            1), dimnames=list(cat="all", year=as.character(1:endy), season="all",                option ="all"))
 
 pos_catches1 <- pos_catches2 <- pop1
 

### 2018-09-04 complex model/complexmodel_MAY.R
improved range for searching sigma

@@ -111,7 +111,7 @@ yield_curve <- function(hr,lratio, wts, natmortality, R=1, sequence = seq(0.001,
   return(res)
 }
 
-SIGMAS        <- seq(10000,100000,5000)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
+SIGMAS        <- seq(30000,120000,5000)   #SIGMA         <- 100 #comes from 2// chanheg from 300 to 200
 
 for (SIGMA in SIGMAS){
 

### 2018-09-04 doc/MSE_MAR_18.tex
added some text to tex file

@@
 
+Some text about increased use of Management strategy evaluation, where the focus has often been on the populaiton dynamics of a single species (examples, key refs).
+However, the human dimension is equally imoprtant (paper Fulton 2011 in FaF, mayby Hilborn 1985).
+MSE with explicit human dim is emerging (Ono 2017,  Andersen et al.
+2010).
+Here we use an MSE with DSVM, including location choice and population is hetorgeneous.
+This is especiallyt relevant
+
 Using a managent strategy evaluation framework, the potential consequences of two species quota scenarios were examined, allowing the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
@@
 \begin{itemize}
+\item{in mixed fisheries, constraining quota for a single species brings risk of changing exploitation patterns, with younger fish being targeted for, hence eroding production potential of stocks}
 \item{Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity, at least, in the short-term.

### 2018-09-04 complex model/complexmodel_MAY.R
fixed bug in daysmen object
We tried to make duplicate rownames in the daysmen object. Now we made sure that these are unique. Note that means that the plotting code for the daysmen object may need to be updated

@@ -438,10 +438,13 @@ nompdays <- aggregate(effort~ option+season, FUN=mean, data=subset(daysmen,(year
 mpdays <- aggregate(effort~ option+season, FUN=mean, data=subset(daysmen,(year %in% c((yy-3):yy))))
 
 # First two rows are the NOMP area north and south. And 2 last rows, MP area north and south
-daysmen<- dcast(nompdays ,option~season, value.var="effort")
-daysmen<- rbind(daysmen, dcast(mpdays ,option~season, value.var="effort"))
-rownames(daysmen)  <- daysmen$option
-daysmen            <- daysmen[,-1]
+daysnomp <- dcast(nompdays ,option~season, value.var="effort")
+rownames(daysnomp) <- paste("nomp",daysnomp$option, sep="_")
+                            
+daysmp <- dcast(mpdays ,option~season, value.var="effort")
+rownames(daysmp) <- paste("mp",daysmp$option, sep="_")
+daysmen          <- rbind(daysnomp,daysmp)
+daysmen[,-1]
 
 #economics
 netrev      <-  melt(economics_res_allyrs, id.vars = "year", measure.vars = c("NetRev"))

### 2018-09-18 doc/MSE_MAR_18.tex
version that NA e-mailed to JJP
and that NA, ADR, and JJP worked on collectively

--- 2018-08-30 doc/MSE_MAR_18.tex
@@
 
-\abstract{Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY).
-The EU common fisheries policy, designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the MSY reference points as targets for exploiting commercially important fish stocks.
+%Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY). 
+\abstract{The EU common fisheries policy (CFP), designed for the long-term environmental, economic, and social sustainability of fishing and aquaculture activities, includes making use of the  maximum sustainable yield (MSY) reference points as targets for exploiting commercially important fish stocks.
 To incentivise individual business to improve selectivity and avoid unwanted catches, it also introduced a landing obligation.
@@
 
-\keywords{management strategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield}
+\keywords{management strategy evaluation; dynamics state variable model; errors in decision-making; landing obligation; maximum sustainable yield, common fisheries policy}
 
@@
 In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$).
-For computations, we used $\sigma = 3000$ (Table \ref{t:fishery}).
+For computations, we used $\sigma = 65000$ (Table \ref{t:fishery}).
 %(noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).
@@
 Landing costs (Euro t\textsuperscript{-1})                  &                       & 0               \\
-Optimal choice error                                        & $\sigma $             & 3000            \\
+Optimal choice error                                        & $\sigma $             & 65000            \\
 
@@
 \ref{f:catches}) caused instability in both populations.
-Overharvesting was caused by more significant harvesting in juvenile specimens than in adults, with a downward sloping harvest rate curves over fish age (rather than selectively targeting older fish with higher price; Fig.4), with average harvest rates of 0.06 (Fig.
+Overharvesting was caused by more significant harvesting in juvenile specimens than in adults, with a downward sloping harvest rate curves over fish age (rather than selectively targeting older fish with higher price; Fig.
+\ref{f:selectivity}), with average harvest rates of 0.06 (Fig.
 \ref{f:selectivity} and \ref{f:yields}, black lines).
@@
 
-Using a managent strategy evaluation framework, the potential consequences of two species quota scenarios were examined, allowing the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
-The MSE framework included temporal and spatial key processes in the dynamics of two fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995, Kell2007}.
+Our results suggest that in mixed fisheries, constraining quota for a single-species brought risk of eroding the production potential of unconstrained species when species differ only in spatial distribution.
+Both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat.
+The changes in production potential were caused by changing exploitation patterns in the fishery due to quota constraints in one species and the resulting spatial reallocation of fishing effort towards the nursery habitat of the other.
+The resulting change in age selectivity of the fishery decreased the production of the unmanaged stock (Fig.
+\ref{f:selectivity}).
+
+Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged.
+However, the stochastic nature of the catches combined with the individual quota result in individual fishers stopping when one of their quotas is reached.
+At the population level this results in overal cacthes that are lower than the TACs set based on the MSY principle.
+Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere (xxxx,xxxx,xx), but thereit was attributed to a difference in the life-history characteristics of the different species.
+Here it is only because of schochasticiyty in catches that is pervasive in all fisheries.DEPENDS ON QUOTA TRADING, AND IF THAT IS POSSIBLE/ HOW GOOD IS MARKET
+
+WHEN MOVING TO MSY HARVEST RATES THERE ARE FEW YEARS WITH REDUCED CATCHES LIKE CLASSICAL MODELS.
+
+The forecasted effects of fisheries management in mixed fisheries result from individual choices by fishers on management and spatial heterogeneity in the resources rather than in differences in biological characestics of the resources.
+This emphsizes the need for incorporating detailed fishing dynamics.
+Histrocially, focus of management strategy evaluation has been primarily on managing the population (biological) dynamics of single and multispecies fisheries \cite{Butterworth1997, DelaMare1996, Dichmont2006, Punt1992, Punt1999, Punt2002, Smith1999}.
+Few MSE studies have explicitly incorporated the direct effects of human dimension harvesting on the stocks, although is equally important \cite{Fulton2011, Hilborn1985}.
+Human behaviour and decision-making are intrinsic parts of conservation and natural resource management.
+MSE with explicit human (social) behaviour dimension is increasingly being used \cite{Andersen2010, Ono2017} as fisheries science takes other sources of uncertainty to a greater extent to identify management actions that ensure the sustainability of fisheries.
+
+Our modeling MSE approach use a dynamic state variable model, including fishing location choices and heterogeneous populations, to evaluate the costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted in response of adopting a major policy change such as the CFP \cite{CFP2013}.
+Starting from an overfished mixed fishery under open access conditions but with a landing obligation constraint, where discarding is not allowed, the response of individual effort allocation and the interannual dynamics for both fleet dynamics and quota allocation is counted when adopting MSY objectives while landing all catches.
 The analysis extended earlier approaches which used short-term state-dependent decision-making models (e.g.
-\cite{Alzorriz2018, Batsleer2016, Poos2010} by including the consequent biomass of fish stocks and the management procedure to make its decision on how to respond to the state of the resource.
-If the landing obligation changes selectivity patterns as predicted in \cite{Alzorriz2018, Batsleer2016}, the use of the present MSE allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling the feedback between fleet and stock dynamics.
+\cite{Alzorriz2018, Batsleer2016, Poos2010}) by including the consequent biomass of fish stocks and the management procedure to make its decision on how to respond to the state of the resource.
+The MSE framework, was thought of as a minimum realistic model \cite{Kell2007, Punt1995}, and included temporal and spatial key processes in the dynamics of two fish populations, the fisheries and a reevaluation of management targets.
+To keep the model simple, the two fish populations modelled here are exactly the same apart from their spatial distribution difference (Fig.
+\ref{f:distribution}).
+As a result, if the landing obligation changes selectivity patterns in the shor-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields.
+ERROR IN DECISION MAKING, NO DISCARING ALLOWED BY FISHERS, AND PERFECT COMPLIANCE TO THIS
+
+When the studied mixed fishery was unmanaged, with unconstrained quota for both species (pre-manage period) and no discarding allowance, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$).
+Overfishing occurs in both stocks when fishers exceed $H_{max}$, whereas rebuilding requires reducing exploitation below $H_{max}$ (Fig.
+\ref{f:yields}).
+Therefore, during the pre-manage period, a high exploitation rate caused a decline in total biomass and fishers mainly harvest the individuals that would normally be added to the population, with younger fish being targeted for.
+Our results suggested that given the spatial and price species distribution parameters (Fig.
+\ref{f:distribution} and \ref{f:prices}), fishing in both the northern and southern fishing grounds were equally profitable; thus, the distribution of fishing effort was equally distributed (Fig.
+\ref{f:effort}).
+
+Moving from an unmanaged period to a situation where one or both species were quota regulated by implementing MSY objectives while maintaining a LO management regime (i.e.
+main CFP objectives: exploiting at MSY and no discarding allowance) led to a very restrictive fishing activity (Fig.
+\ref{f:effort}).
+Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term.
+In addition to reconciling fishery and conservation objectives (such as the CFP), in agreement with \cite{Grafton2007}, setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig.
+\ref{f:effort}).
+However, as \cite{Prellezo2016a} concluded, since effort was drastically reduced to achieve the objectives of the CFP; the increase in net revenue gains were only experienced by some vessels.
+Consequently, a significant part of the fleet had to cease its activity (Fig.
+\ref{f:effort}).
+
 Nevertheless, the framework used made several assumptions: that there was perfect knowledge of the stochastic nature of catch rates, full compliance of fishers with the discarding policy, lack of transferability of individual quotas, and lack of interference competition among fishers \cite{Alzorriz2018}.
-Regarding fish populations, also important assumptions were established: constant mortality and almost neligible, constant recruitment along the years, equal conversion length to weight for the two species, constant migration that lead to a clear segregation of the youngers, age-dependant catchability assumed independent of age and equal for the two species and prices were age dependant but fixed over time.
-Also, we used monthly time steps in the model to keep computation time low, while the time scale at which fishing trips occur may be much smaller.
-Finally, we did not include the existence of any landing obligation discard flexibilities (CFP Article 15, \cite{CFP2013}) or year transfer swaps.
-We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under the landing obligation.
-Despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery fleet and examine whether potential harvest strategies can achieve the management objectives.
+IMPORTANT< UTILITY FUNCTION IS ANNUAL NET REV, NO SOCIAL DYNAMICS.
+Regarding fish populations, also important assumptions were established: negligible natural mortality, constant recruitment, constant growth, constant migration, and constant size-dependent prices.
+Individual choices on fishing behaviour were assumed to be made once a month.
+However, despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery and examine the effects of single species quota management.
 
-When the fishery was unmanaged, with unconstrained quota for both species (pre-manage period), the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$), thus resulted in a situation where the stocks were overfished, reaching total catches of more than 4 tons per species (Fig.
-\ref{f:catches}), which in terms of activity would be fishing around 3100 days per area and season (Fig.
-\ref{f:catchesbyarea} and \ref{f:meaneffort} a and c).
-Given the spatial and price distribution of the species' parameters (Fig.
-\ref{f:distribution} and  \ref{f:prices}), fishing in both the northern and southern fishing grounds would be equally profitable.
-Our results suggested that 80\% of the fleet  indistinctly choosed harvesting in both areas, with net profits of 130 thousand Euros/ year (Fig.
-\ref{f:effort} b and d).
-Fuel costs related to fishing choices were 40\% of the incurred gross revenues (Fig.
-\ref{f:effort} b an d).
+In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of any discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps.
+Allowing the latter could potentially remedy the underutilisation of quota in the two quota situation, given that this reuslts from the stochsticity in catches, which are uncorrelated.
 
-Moving from an unmanaged period to a situation where one or both species were regulated by quotas and forced to comply simulataneously with management objectives (i.e.
-exploiting at MSY and no discarding allowance) on the optimal effort allocation between fleets was analysed.
-For example, the underlying assumption of the two quota scenarios that fleets would switch targeted stocks to their relative abundance and management constraints was tested.
-Under both scenarios, our model results showed that simultaneously implementation of MSY and LO objectives led to a very restrictive fishing activity.
-Effort is constrained by some quota stock, as predicted in general terms by \cite{Ulrich2017}.
-Short-term reductions in effort would be needed while mid-term interactions between fleets and stocks that occur in mixed fisheries such as t
+ We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY.
+However, applying any kind of exemption or flexibility beyond the MSY would be negative \cite{Prellezo2016}.
+Consequently, a critical feature of management bodies is to ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation.
+Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}.
+The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance.
+With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2015, Batsleer2013}.
+In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations.
+However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.
+IF YOU WANT TO STUDY THIS, THEN DETAILED PARAMETRIZATION IS NEEDED THAT INCLUDES CASE DEPENDENT LH CHARATRISTICS AND MIGRATIONS
+ %Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation.
+The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis
 
-%Consequently, no comparisons could be made to verify the hypothesis stating that an effort regulation is more biologically robust as there are no incentives for discards and black landings.
-This hypothesis would be interesting to investigate in the model as, according to Faroese scientists, it seems to be one of the main advantages of this management system %(Jákupsstovu & Reinert, personal communication).
+Finally, a dynamic state variable model was used as a tool to create a quick quota allocation and fleet dynamics model.
+Many detailed fleet dynamics models have been developed over the past decades \cite{Abbott2011, Ono2017, Simons2015} and they could have been used to include spatial factors as another dimension in the MSE.
+However, following \cite{Dowling2011}, incorporating uncertainty in errors in human dynamic decision-making help us to understant and predict changes in human behaviour in response of adopting a major policy change such as the CFP \cite{CFP2013}.
+The effect of introducing a realistic multispecies fishery within this MSE framework still remains the next step to be taken.
+Modelling and predicting realistic human behavior with its associated uncertainty in an ever-changing socioecological system will remain a major challenge to address in the future \cite{Ono2017}.
 
-Achieving these complex objectives may not necessarily reduce harvest rates, but will change the underlying selectivity patterns of harvests.
-We contrast two management plans to achieve MSY from two stocks and compare their effectiveness based on a management strategy evaluation that uses a dynamic state variable model including errors in decision-making in its operating model.
-The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
-Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation.
-Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation.
-The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.
+\section{Conclusions}
 
-\begin{itemize}
-\item{Simultaneously implementation of MSY and LO objectives may lead to a very restrictive fishing activity, at least, in the short-term.
-Effort is constrained by some stock, as predicted, in general terms, by Ulrich et al.
-(2011).}
-\item{Drastic reduction of effort would increase profits if MSY and LO were contemplated.
-However, no flexibilities or year transfer swaps were considered.
-Be carefull if applying flexibilities, because as Prellezo 2016 (the results obtained from applying any kind of exemption or flexibility are, simply, negative) said, any increased in mortality beyond the MSY would be negative.}
-\item{once the status of the stock has been assessed relative to its MSY level, the calculation of a TAC comes automatically from the rules of the Management Procedure.
-Shifting areas toward complying TAC limitations produced changes in selectivity ant therefore, constant changes in MSY levels.
-(e.g.
-shifting selectivity toward smaller/younger fish will generally reduce MSY).
-if it is required that the TAC should correspond to the level where MSY can be obtained, it will be necessary to monitor continuously the stock abundance and to revise TACs on a regular basis.}
-\item{The  MSY under two quota species may be below what was achieved in one quota species because overall selectivity has shifted to smaller fish and low harvest of older fish; the impact of this change in selectivity on estimates of MSY is clearly seen in the results from Fig.
-\ref{f:selectivity} and \ref{f:yields}.
-While  the  potential  MSY  has decreased over time the stock biomass required to produce this MSY has increased.}
-\item{Catch quota management is a more important mechanism than the actual obligation to land.
-An accurate and monitoring of all catches is a primary step towards sustainable and responsible fisheries.}
-\item{The compliance of the fleets, the changes in behaviour and the availability and reliability of catch statistics (Catchpole and Riveiro Santos, 2015) .
-Data risks should not be overlooked (Hentati-Sundberg et al., 2014).
-Fish stock assessment provides the foundation for scientific advice to decision-makers on sustainable catch levels.
-A critical feature of stock assessment methodology is its dependence on accurate commercial data, which makes it susceptible to biased reporting (Walters and Maguire, 1996; Cotter et al., 2004; Bousquet et al., 2010).
-The reasons for misreporting and underreporting include e.g.
-IUU fishing (Agnew et al., 2009), discard obligation (Kelleher, 2005; Wielgus et al., 2010) and catches from recreational fisheries (Zeller et al., 2011).
-Such misreporting can be driven by fishing fleet overcapacity, resulting from restrictive TACs and/or effort regulations (Patterson, 1998; Beare et al., 2005; Dobby et al., 2008) that create economic incentives for non-compliance (Sumaila et al., 2006; Gallic and Cox, 2006; Osterblom et al., 2011).}
-
-\end{itemize}
-
-%When discarding was allowed, fishers did not avoid areas where catching undersized specimens was likely because there was no incentive to do so.
-Under all three implementations (scenarios) of the landing obligation, however, our model showed that the Basque trawling fleet was likely to move to areas with lower catch rates of quota species to avoid exceeding quota limitations.
-We showed that in the short term the landing obligation would likely result in a loss in profitability for the fleet.
-This loss in profitability was driven by the ``choke'' effect for species such as mackerel (scenario II and III), which was triggered by reductions in effort as a result of rapid quota exhaustion and by greater costs associated with fishing further from home ports (scenario IV).
-
-%Our results suggest that the Basque fleet is limited by hake, horse mackerel and mackerel quota.
-In scenarios II and III, the choke effect of mackerel is so strong that the effect of other species limits (hake under scenario II, horse mackerel under scenario III) is negligible.
-Consequently, under both these scenarios our model predicts that the fleet will avoid using OTB gear and will reduce fishing effort in sub-area 7 and 8 because of their high mackerel catch rates (Fig.
-\ref{fig:effort_modelval}).
-Thus, fishers will likely visit sub-area 6 and 8 (using PTB gear).
-Revenues under these two scenarios are predicted to result in a loss of profitability (60\% less than the status quo; Table ~\ref{t:net_rev}).
-These findings indicate that trading quotas (at rates undertaken in the past) does not alleviate the choking effect of the landing obligation.
-Therefore, new ways of swapping quotas would be needed under the landing obligation to alleviate the choke effect of any of the quota limiting species.
-Fleets with high discards rates of certain species will be likely to have larger incentives to increase their quotas for those species than fleets with low discard rates.
-
-%In scenario IV, the addition of inter-species quota-swapping flexibility to the landing obligation (e.g.
-9\% of the horse mackerel catch quota could go to increase mackerel quota) substantially decreased the mackerel choke effect and increased both fishing opportunities and net revenues, when compared to scenarios II and III.
-Under this scenario, revenue was 30\% lower than when discarding was allowed (scenario I).
-
-%Our results suggest that reductions in fishing opportunities under a landing obligation that are caused by choke species could be alleviated by the use of inter-species quota flexibility, and that there will be strong incentives to use this flexibility.
-Management bodies should ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation.
-Use of inter-species quota flexibility could cause serious impacts to stock abundance when quota from a high volume/low value species is transferred to a low volume/high value stock.
-Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries.
-
-%The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance.
-With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2005, Batsleer2013}.
-In our model, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations.
-However, if detection rates are low and the short–term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.
-
-%The landing obligation was implemented to incentivise individual businesses to improve selectivity and avoid unwanted catches.
-The landing obligation shifts the emphasis from prescribing the fishing gears that are permissible, to a results-based system where the onus is on the fishers to use fishing gear that will avoid unwanted catches.
-Therefore, fishers should be incentivised to develop alternative gear or adapt their fishing behaviour to mitigate the full impact of the landing obligation \cite{Alzorriz2016}.
-Discard ban policies may not necessarily reduce harvest rates but will change the underlying selectivity patterns of harvests \cite{Borges2016}.
-Our results reflect the short-term effects of a discard ban on selectivity patterns and net revenues implemented under three scenarios.
-If the landing obligation changes selectivity patterns as we predict, then we would also expect longer term changes in both fish populations and the economy of the fishery.
-Such analyses are the domain of bio-economic management strategy evaluations where the feedback between fleet and stock dynamics are modelled explicitly \cite{Andersen2010, Prellezo2016a, Punt2016}.
-In such models, effects of the changing selectivity of the fishing fleets on the reference points for sustainable exploitation can be evaluated.
+Achieving optimal yield from mixed fisheries is a pervasive problem in fisheries management (e.g., \cite{Farcas2016, Prellezo2016b, Salomon2014, Ulrich2017, Voss2014}).
+A multispecies fishery MSE framework based on two populations was developed in this study to examine different management scenarios.
+The results highlight pitfalls in managing mixed fisheries where erosion of production potential can take place, or unintentional underharvesting.
+We argue for the inclusion of spatial dynamics and fleet dynamics when designing a multispecies fishery management strategy evaluation.
+Such MSE should be done  when adopting a major policy change such as the CFP.
+Future studies aimed at building a realistic multispecies fishery MSE with significant discard constraints, and should consider expanding the current model to include the whole suite of species exploited and managed in a real mixed fishery.
+The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis.
 

### 2018-10-24 doc/MSE_MAR_18.tex
fixed issues with citations

@@
 
-Our results suggest that in mixed fisheries, constraining quota for a single-species brought risk of eroding the production potential of unconstrained species when species differ only in spatial distribution.
+Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species when species differ only in spatial distribution.
 Both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat.
@@
 This emphsizes the need for incorporating detailed fishing dynamics.
-Histrocially, focus of management strategy evaluation has been primarily on managing the population (biological) dynamics of single and multispecies fisheries \cite{Butterworth1997, DelaMare1996, Dichmont2006, Punt1992, Punt1999, Punt2002, Smith1999}.
-Few MSE studies have explicitly incorporated the direct effects of human dimension harvesting on the stocks, although is equally important \cite{Fulton2011, Hilborn1985}.
+Histrocially, focus of management strategy evaluation has been primarily on managing the population (biological) dynamics of single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999}.
+Few MSE studies have explicitly incorporated the direct effects of human dimension harvesting on the stocks, although is equally important \cite{Fulton2011}.
 Human behaviour and decision-making are intrinsic parts of conservation and natural resource management.
-MSE with explicit human (social) behaviour dimension is increasingly being used \cite{Andersen2010, Ono2017} as fisheries science takes other sources of uncertainty to a greater extent to identify management actions that ensure the sustainability of fisheries.
+MSE with explicit human (social) behaviour dimension is increasingly being used \cite{Andersen2010, Ono2018} as fisheries science takes other sources of uncertainty to a greater extent to identify management actions that ensure the sustainability of fisheries.
 
@@
 Finally, a dynamic state variable model was used as a tool to create a quick quota allocation and fleet dynamics model.
-Many detailed fleet dynamics models have been developed over the past decades \cite{Abbott2011, Ono2017, Simons2015} and they could have been used to include spatial factors as another dimension in the MSE.
+Many detailed fleet dynamics models have been developed over the past decades \cite{Andersen2010, Ono2018, Simons2015a} and they could have been used to include spatial factors as another dimension in the MSE.
 However, following \cite{Dowling2011}, incorporating uncertainty in errors in human dynamic decision-making help us to understant and predict changes in human behaviour in response of adopting a major policy change such as the CFP \cite{CFP2013}.
 The effect of introducing a realistic multispecies fishery within this MSE framework still remains the next step to be taken.
-Modelling and predicting realistic human behavior with its associated uncertainty in an ever-changing socioecological system will remain a major challenge to address in the future \cite{Ono2017}.
+Modelling and predicting realistic human behavior with its associated uncertainty in an ever-changing socioecological system will remain a major challenge to address in the future \cite{Ono2018}.
 

### 2019-04-04 doc/MSE_MAR_18.tex
small text fixes

@@
 \and
+\name{Adriaan Rijnsdorp}
+\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
+\email{adriaan.rijnsdorp@wur.nl}
+\and
 \name{Jan Jaap Poos}
-\address{Wageningen Marine Research, PO Box 68, 1970 AB IJmuiden, The Netherlands.}
+\address{Wageningen University, The Netherlands.}
 \email{janjaap.poos@wur.nl}}
-
 %Most worldwide fish stocks are overexploited, and so exploited beyond the maximum sustainable yield (MSY). 
@@
 Prices of fish were assumed to be dependent over fish age (subsection size dependent pricing).
-$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while  $ C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices.
+$G(p)$ represented the incurred fuel costs per season from the choice of fishing area $p$, while $C(p)$ represented the variable operating costs (crew share, gear maintenance and landing costs, Table~\ref{t:fishery}), which in turn depended on the change in cumulative catches and fish prices.
 The term $E_{p}[V (C_i^\prime, Q_i, F_i, s+1]$ denoted the expected future utility taken over all possible states resulting from choices $p$.
@@
 
-Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species when species differ only in spatial distribution.
-Both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat.
-The changes in production potential were caused by changing exploitation patterns in the fishery due to quota constraints in one species and the resulting spatial reallocation of fishing effort towards the nursery habitat of the other.
-The resulting change in age selectivity of the fishery decreased the production of the unmanaged stock (Fig.
+Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species.
+The only prerequisite for this to happen is a difference in the spatial ontogeny of the species: both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat.
+Changes in the quota constraints led to a spatial redistribution of the fishery.
+As a result of this spatial redistribution, the age-dependent exploitation for the species changed.
+For the unconstrained species, the quota resulted in a shift of the fishing pressure towards the younger ages.
+This in turn decreased the production of the unmanaged stock, as can be seen from the yield curves (Fig.
 \ref{f:selectivity}).
@@
 Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged.
-However, the stochastic nature of the catches combined with the individual quota result in individual fishers stopping when one of their quotas is reached.
-At the population level this results in overal cacthes that are lower than the TACs set based on the MSY principle.
+However, individual quota cobined with the stochastic fisheries catches resulted in lower than intended exploitation rates.
+The TACs were set so that the total TACs corresponded to the harvest rate that resulted in maximum sustainable yields.
+The TACs were divided in individual quotas for each fishing vessel.
+The lower than intended harvest rates were caused by the fact that despite their planning, fishers often exhausted one of the quotas before the other.
+If discarding is not allowed, then individual fishers must then stop fishing.
+At the population levels of the two species this resulted in overal cacthes that were lower than the TACs.
 Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere (xxxx,xxxx,xx), but thereit was attributed to a difference in the life-history characteristics of the different species.
@@
 Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term.
-In addition to reconciling fishery and conservation objectives (such as the CFP), in agreement with \cite{Grafton2007}, setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig.
+In addition to reconciling fishery and conservation objectives (such as the CFP), setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig.
 \ref{f:effort}).
@@
 The success of any new regulation also depends on achieving a balance between the costs and benefits of measures that promote voluntary compliance and those that punish non-compliance.
-With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Catchpole2015, Batsleer2013}.
+With poor enforcement the economic benefits of non-compliance outweigh the risks of detection \cite{Batsleer2013}.
 In this MSE framework, detection rates were assumed to be 100\% and fines were set sufficiently high so that no discarding occurred, reflecting full compliance with regulations.

### 2019-04-04 doc/MSE_MAR_18.tex
fixed title and intro

@@
 
-\title{MSE and dynamic state variable model with sigma function}
+\title{Modelling long-term effects of management strategies in mixed fisheries}
 
@@
 
-Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulation\cite{Punt2016}.
-Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}.
-Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable.
-As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy.
+Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulations \cite{Punt2016}.
+Such simulations should include all important processes of a fishery.
+Any fishery is inherently a socio-ecological system, with coupled dynamics of fishing fleets, exploited stocks, and their governance \cite{Punt2016, Rasemeyer2007}.
+Given the complexity of socio-ecologicals system the uncertainy about the dynamics and feedbacks are consideable.
+As a result, any forecasted success in achieving goals of management strategies may thus be small compared to system stochasticity, or worse, be counteracted by unintended consequences of the management strategy.
 The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
@@
 The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}.
-Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
-New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed.
-These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual.
-The effect of a choice on the utility depends on the economic environment, such as the home port of the vessel and the distance to fishing grounds, and the biological enviroment, such as the spatial distribution of the resources.
-Using such dynamic variable state models, effects of e.g.
+Simulation tools for such a response are available, borrowing methodology from state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}.
+State-dependent behaviour of individual fishing vessels, aggregated into fleet dynamics, have been developed using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995}.
+These models generally predict short term (within a fishing trip, or a quota year) effects of changes in management or the environment by determining which choices yields the best chance of increasing utility, while keeping track of the state of each individual.
+The effect of a choice on utility depends on the state of the individual, the economic environment, such as fuel prices, and the biological enviroment, such as the spatial distribution of resources.
+
+Using dynamic variable state models, effects of e.g.
 relocation costs of marine protected areas have been modelled \cite{Dowling2011}.
+Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995, Alzorriz2018}.
+Some of these studies have tried to shed light on the potential outcomes of the European fisheries management reform of 2013.
+That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commercially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
+Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted high costs in mixed fisheries in the short term as a result of the LO \cite{Alzorriz2018, Batsleer2013, Batsleer2015}.
+These costs result from removing the choice to discard part of the catch and in doing so adapt landings to quota \cite{Alzorriz2018}.
+This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.
 
-Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995}.
-These studies may shed light on the potential outcomes of the European fisheries management reform of 2013.
-That reform included using the use of the maximum sustainable yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}.
-Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted high costs in mixed fisheries in the short term as a result of the LO.
-These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}.
-This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.
-However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
+Meanwhile, the LO potentially improves selectivity of fisheries by providing incentives to move away from areas with high levels of unwanted catch and the use of more selective gears \cite{Alzorriz2016, Alzorriz2018}.
+The improved selectivity should lead to higher long term catches.
+These long term benefits to fisheries that could result from improved selectivity under the LO are ignored in previous studies on the adaptive response of fisheries \cite{Batsleer2015, Batsleer2013, Alzorriz2018}.
 Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
-Because the LO should provide incentives for the use of more selective gears and for fisheries to move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
-Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013, Ulrich2017}.
-Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.
-This will result in fishing effort reallocation and early closures of fisheries once quota have been reached.
-However, the long term benefits to fisheries that could result from improved selectivity are ignored in those studies.
-Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO.
-Because the LO should provide incentives for the use of more selective gears and for fisheries to move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches.
-To enable longer term prediction of the effects of changes in fisheries management, and thus complete the MSE, one needs to couple the fleet dynamic models that forecast the fleet response to management to biological dynamic models that forecast the fish population responses to the changing fleet response.
-We present an MSE that encompasses the individual harvester decision-making process and socioeconomic drivers on management effectiveness, i.e.
-if the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}.
-Such combination needs to be responsive to all the drivers that influence and motivate the actual fleet and to respect existing constraints \cite{Venables2009}.
-The drivers will include the (real or perceived) local abundance and catchability of the fished stocks and various resource costs, principally that of fuel.
-The constraints will include the obvious management-imposed constraints such as levels of total allowable catch, and economical constraints such as the contribution of the annual fines for exceeding catch quotas.
+In mixed fisheries, however, optimizing single species selectivity and achieving MSY while landing all commercial catches is challenging because of trade-offs in the catches of different species and size-classes \cite{Batsleer2013, Ulrich2017}.
+Forecasting whether any management will be effective requires understanding the response of fishers constrained by these trade-offs and the effects of the response on the development of fish populations \cite{Venables2009}.
+If the fleet dynamic model results in a short term movement of vessels to allocate effort both spatially and temporally, we also expect longer term changes to occur in fish populations and the economy of the fishery \cite{Alzorriz2018}.
 
-Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.
-To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences.
+We present an MSE framework that couples forecasts from a state-dependent behaviour model to a biological dynamic models that forecasts population effects resulting from the collective fishers behaviour.
+The model includes the (real or perceived) local biomass of two populations, governed by growth, mortality, and migration.
+These populations are size-structured, and are modelled in a spatially heterogeneous environment.
+The populations are harvested by a single fleet that has annual fishing rights for both populations, consisting of individuals who make monthly choices on fishing locations.
+These choices depend on the perceived spatial distributions of biomasses, but also on various costs, principally that of fuel.
+The constraints on the fleet  includes management-imposed constraints such as fish quotas.
+In the framework, these quotas are decided by a a management body that observes the total biomass of the two stocks and with the goal of constraining fishing mortality so that maximum sustainable yields are achieved.
+
+The MSE framework is usedt o understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 
-%Since maximizing effort does not maximize catch, the question is if there is an optimum fishing rate that would lead to a maximum yield.
-
 \section{Methods}
 
-Managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
+The managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
 In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.

### 2019-04-04 doc/MSE_MAR_18.tex
fixed typos and some language issues discussion

@@
 The results show that when accounting for quota on one specie and for quotas on both species, MSY landings in the short-term are 30\% to 50\% lower than when compared to an unmanaged situation simulations.
-Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situtation.
+Some years after the implementation, MSY landings are similar or 25\% to 50\% higher than the pre-manage situation.
 Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.4 times larger profits than before the CFP implementation.
-The results show taht for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}
+The results show that for both species, any of the two management plans would be beneficial to both the stock and part of the fleets over the mid-term compared with the projections over the short-term.}
 
@@
 Any fishery is inherently a socio-ecological system, with coupled dynamics of fishing fleets, exploited stocks, and their governance \cite{Punt2016, Rasemeyer2007}.
-Given the complexity of socio-ecologicals system the uncertainy about the dynamics and feedbacks are consideable.
+Given the complexity of socio-ecological system the uncertainy about the dynamics and feedbacks are considerable.
 As a result, any forecasted success in achieving goals of management strategies may thus be small compared to system stochasticity, or worse, be counteracted by unintended consequences of the management strategy.
-The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
+The robustness of management strategies to the uncertainty in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.
 
@@
 These models generally predict short term (within a fishing trip, or a quota year) effects of changes in management or the environment by determining which choices yields the best chance of increasing utility, while keeping track of the state of each individual.
-The effect of a choice on utility depends on the state of the individual, the economic environment, such as fuel prices, and the biological enviroment, such as the spatial distribution of resources.
+The effect of a choice on utility depends on the state of the individual, the economic environment, such as fuel prices, and the biological environment, such as the spatial distribution of resources.
 
@@
 
-The MSE framework is usedt o understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences.
+The MSE framework is used to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed combined with the LO where discards are not allowed, and to compare it in order to assess on the biological and economic consequences.
 If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding.
@@
 
-The managament strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
-In the fishery, there are three essential elements: (i) a collection of size strucured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.
+The management strategy evaluation framework was used to forecast the dynamics of a hypothetical mixed fishery on two species.
+In the fishery, there are three essential elements: (i) a collection of size structured fish stocks, whose dynamics are governed by annual reproduction, growth, migration, and mortality; (ii) a management body that evaluates the fishing pressure and aims to set annual quotas in accordance with fishing mortalities dictated by $H_{max}$ reference points (harvest at maximum sustainable yield); and (iii) a fleet of individual fishers who aim to make the best use of their annual quota.
 Within this mixed fishery, individual vessels make adaptive choices about fishing location and discarding that depend on the distribution of the resources, and the quota that hold.
@@
 The framework consists of an operating model (OM) and a management procedure (MP).
-The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
+The OM captured the key processes in the dynamics of the fish populations, the fisheries, and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.
 The OM thus includes individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance (Fig.
@@
 
-Mortality in the model resulted from fishery catches and natural causes such as predation, diseases, and scenesens.
+Mortality in the model resulted from fishery catches and natural causes such as predation, diseases, and senescence.
 The decrease in population numbers was thus the result of the catches ($C_i (a, y, s, p)$) and a natural mortality constant $M_i$ that described natural mortality as a fixed fraction of the population.
@@
 emigrants leave area $em$ and move to area $im$.
-The emigrated part of the population is then substracted from each of the areas, so that that population numbers per year and season remain unaffected by migration.
+The emigrated part of the population is then subtracted from each of the areas, so that that population numbers per year and season remain unaffected by migration.
 
@@
 
-The realized catches are the sum of all individual catches resulting from the Dynamic State Variable Model.
-The Dynamic State Variable model inputs consist of the expected individual catch rates, which are random variables.
+The realized catches are the sum of all individual catches resulting from the dynamic state variable model.
+The model inputs consist of the expected individual catch rates, which are random variables.
 These random variables were normally distributed, with means $\hat c_i (a, y, s, p)$ being a function of population size, age dependent catchability $q_i (a)$ in any year, season, and area,
@@
 
-To simulate a fleet of individual fishing vessels,we used a Dynamic State Variable Model \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
-The model was used to model location choice, same model structure as in \cite{Alzorriz2018}, but without the choice to discard one or more species and size classes while fishing.
+A dynamic state variable model was used to simulate a fleet of individual fishing vessels \cite{Alzorriz2018, Batsleer2015, ClarkandMangel2000, Dowling2011, Houston1999, Poos2010}.
+The model structure was equal to \cite{Alzorriz2018}.
 Each individual vessels had a set of choices, which include the choice to go fishing in a season, location choice within that season.
+No discarding was allowed in the model.
 The model had annual fines for exceeding landings quota as in \cite{Alzorriz2018}.
@@
 
-The dynamic state variable model was solved by iterating backwards in time, while finding the probability distribution choice in terms of location for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota.
+The dynamic state variable model was solved for each year by iterating backwards in time, while finding the probability distribution choice in terms of location for all possible states, combining the net revenue obtained from the sale of fish and costs of a fishing trip and the effect of the annual fines when exceeding annual quota.
 Further details for this procedure can be found in \cite{Alzorriz2018, Batsleer2016} and \cite{Dowling2011}.
@@
 Prices of fish were assumed to be fixed over time but influenced by the body weight of the individuals in the catch, as is commonly observed \cite{Zimmermann2011, Zimmermann2013}.
-Following \cite{Zimmermann2011}, therelationship between fish price and body weight was modelled as:
+Following \cite{Zimmermann2011}, the relationship between fish price and body weight was modelled as:
 \begin{equation}
@@
 Fishing at a harvest rate that is equal to $H_{i,max}$ should lead to maximum sustainable yields.
-The yield per recruit curve omited the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment.
-The HCR used by the management body resuslted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$.
+The yield per recruit curve omitted the potential effect of the feedback between adult biomass and recruitment, which anyway was absent in the population dynamics, that assumed a constant recruitment.
+The HCR used by the management body resulted in annual quotas such that the harvest rate in a year corresponded to $H_{i,max}$.
 These quotas were then divided equally over the individual vessels in the simulations.
@@
 The model was set up in three consecutive time windows.
-In the absence of a plusgroup and 6 ages, the first window of 10 years was enough to get to stable populations.
+The first window of 10 years was enough to get to stable populations.
 During this first period, new young individuals entered the population in the absence of fishing.
@@
 The consecutive 15 years, the fleet of 8000 vessels started fishing in the absence of any fisheries regulations.
-During this time window, the fishery was unmanaged, by meaned with unconstrained quota for both species (pre-manage period).
+During this time window, the fishery was unmanaged, with unlimited quota for both species (pre-manage period).
 During the last 15 years of projections, the management procedure was introduced, the fishery was constrained by setting MSY targets while discarding was not allowed (post-manage period).
 
-Two scenarios based on two management scenarios were examined in the model.
-The two management scenarios considered differ simply by the number of stocks forced to meet the objective of the plan.
-For this purpose, in a single-species catch quota context: quota on species 1 was was controlled by HCR in first scenario, while in both-species catch quota context: quotas on both species were controlled by HCRs in second scenario.
+Two management scenarios were examined in the model.
+The two scenarios differ in the number of stocks under quota.
+In the "single-species catch quota" scenario quota on species 1 was controlled by the HCR.
+In the "both-species catch quota"  quotas on both species were controlled by HCRs.
 The comparison among these scenarios allowed the evaluation of costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted from the response of individual effort allocation to meet the objectives of the landing obligation regarding the objectives of the CFP, specially the MSY.
@@
 \includegraphics[width=0.8\textwidth]{Figures/Selectivity.eps} 
-\caption{Modelled changes in harvest rates for both species in relation to the catch decision options made based in the availble individual quota: species 1 (a,c) and species 2 (b,d).
+\caption{Modelled changes in harvest rates for both species in relation to the catch decision options made based in the available individual quota: species 1 (a,c) and species 2 (b,d).
 In top panels only species 1 quota constrained the fishery (a,b), while in bottom panels both species quota constrained the fishery (c,d).
@@
 % YIELDS 
-Expected yields as a function of harvest show that during the post-manage period both species were underharvested to reach a sustainable harvesting (Fig.
-\ref{f:yields}; grey lines in a,c,d), except species 2 in the single-stock quota case scenario which was overharvested due to the abscence of any quota restriction (Fig.
+Expected yields as a function of harvest show that during the post-manage period both species were underharvested (Fig.
+\ref{f:yields}; grey lines in a,c,d), except species 2 in the single-stock quota case scenario which was overharvested due to the absence of any quota restriction (Fig.
 \ref{f:yields}; panel b).
 Indeed, higher sustainable yields achieved with lower values of fishing mortality.
-When quotas on both species, yields were 25\% higher than in the unmanaged situation; just on one specie quota, yield for specie 1 was 50\% higher than the unconstrined fishery value.
+When quotas on both species, yields were 25\% higher than in the unmanaged situation; just on one specie quota, yield for specie 1 was 50\% higher than the unconstrained fishery value.
 These differences in quota scenarios could be explained due to the already commented choke effect.
@@
 \ref{f:effort} b).
-After the initial 57 \% to 32 \% decline in cathes of species 1 and 2 due to the MP implementation, fishing in the southern (northern) area, during the post-manage period, showed in average 181.67\% and 12.87\% higher catches of species 1 and 2 than the pre-manage period, respectively (16.91\% and 38.12\% lower catches of species 1 and 2)(Fig.
+After the initial 57 \% to 32 \% decline in catches of species 1 and 2 due to the MP implementation, fishing in the southern (northern) area, during the post-manage period, showed in average 181.67\% and 12.87\% higher catches of species 1 and 2 than the pre-manage period, respectively (16.91\% and 38.12\% lower catches of species 1 and 2)(Fig.
 \ref{f:catchesbyarea} a and b).
@@
 Vessels indistinctly allocated the fishing effort in both fishing grounds (such in the pre-manage period) due to same model conditioning for both species.
-Therefore, after the initial 62 \% to 65 \% decline in cathes due to the MP implementation, fishing in the northern (southern) area, during the post-manage period, somulations showed in average 2.57\% and 54.54\% higher catches of species 1 and 2 than the pre-manage period, respectively (55.46\% and 4.20\%)(Fig.
+Therefore, after the initial 62 \% to 65 \% decline in cathes due to the MP implementation, fishing in the northern (southern) area, during the post-manage period, simulations showed in average 2.57\% and 54.54\% higher catches of species 1 and 2 than the pre-manage period, respectively (55.46\% and 4.20\%)(Fig.
 \ref{f:catchesbyarea} c and d).
@@
 
-Modelling results sugested that moving from an unmanaged period to a sustainable stock situation, where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY while it was not possible to discard, moderate and high reductions of fishing effort are expected (Fig.
+Modelling results suggested that moving from an unmanaged period to a sustainable stock situation, where one or both species were regulated by quotas and forced to keep fishing mortality below the MSY while it was not possible to discard, moderate and high reductions of fishing effort are expected (Fig.
 \ref{f:effort}, b and d).
@@
 \ref{f:meaneffort}, b and d).
-The relative annual contribution by vessel to the gross revenue was arouond 28 Euros, reduced to 17 Euros of net revenues due to the costs associated to fuel during the unmanage situation (Fig.
+The relative annual contribution by vessel to the gross revenue was around 28 Euros, reduced to 17 Euros of net revenues due to the costs associated to fuel during the unmanaged situation (Fig.
 \ref{f:meanecon}).
@@
 \includegraphics[width=0.8\textwidth]{Figures/Mean_Economics.eps} 
-\caption{Modelled main economics by vessel (euro/ year): median annual net revenues (a,c) and median grooss revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
+\caption{Modelled main economics by vessel (euro/ year): median annual net revenues (a,c) and median gross revenues (b,d) with the upper and lower limits of the box being the third and first quartile1 (75th and 25th percentile) respectively.
 MP year reflects the year where the management plan was introduced.}
@@
 Our results suggest that in mixed fisheries, constraining quota for a single-species can bring risk of eroding the production potential of unconstrained species.
-The only prerequisite for this to happen is a difference in the spatial ontogeny of the species: both species in this study were the same with respect to growth, recruitment, mortality, and price, but differed in the location of their nursery habitat.
-Changes in the quota constraints led to a spatial redistribution of the fishery.
-As a result of this spatial redistribution, the age-dependent exploitation for the species changed.
-For the unconstrained species, the quota resulted in a shift of the fishing pressure towards the younger ages.
+The only prerequisite for this to happen is a difference in the spatial ontogeny of the species: species differed only in the location of their nursery habitat.
+Growth, recruitment numbers, mortality, and price were the same for both species.
+Changes in quota constraints led to a spatial redistribution of the fishery and as a result of this  redistribution the age-dependent exploitation for both species changed.
+For the unconstrained species, the quota resulted in a shift of the fishing pressure towards younger ages.
 This in turn decreased the production of the unmanaged stock, as can be seen from the yield curves (Fig.

### 2019-04-10 doc/MSE_MAR_18.tex
worked on discussion

@@
 At the population levels of the two species this resulted in overal cacthes that were lower than the TACs.
-Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere (xxxx,xxxx,xx), but thereit was attributed to a difference in the life-history characteristics of the different species.
-Here it is only because of schochasticiyty in catches that is pervasive in all fisheries.DEPENDS ON QUOTA TRADING, AND IF THAT IS POSSIBLE/ HOW GOOD IS MARKET
+Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere \cite{Kempf2016, Rindorf2017, Hilborn2015}, but is generall attributed to differences in life-histories among fished species.
+In this study, the discrepancy between the intended annual quotas at MSY and realized catches is caused only by schochasticity in the system.
+This stochasticy in catch rates is observed in most fisheries  \cite{Sampson1988, VanOostenbrugge2004, Smith1980, Bernasconi2015}.
+It is caused by the complex migration biology and the effects of environmental factors on fish behaviour.
+A well functioning quota market would allow vessels to swap excess quota for constraining and thus bring realized catches closer to quotas.
 
-WHEN MOVING TO MSY HARVEST RATES THERE ARE FEW YEARS WITH REDUCED CATCHES LIKE CLASSICAL MODELS.
+The model aimed to evaluate the biological and economic effects of adopting a major policy change such as the CFP \cite{CFP2013}.
+Starting from an overfished mixed fishery under open access conditions but with a landing obligation, the response of individual effort allocation and the interannual population dynamics are explicitly modeled, while adopting MSY objectives.
+The analysis extended earlier approaches that used short-term state-dependent decision-making models (e.g.
+\cite{Alzorriz2018, Batsleer2016, Poos2010}) by including interannual population dynamics of fish stocks, and a management procedure that mimics the decisions of a management system.
+Such an MSE framework was thought of as a minimum realistic model \cite{Kell2007, Punt1995}.
+To keep the outcomes of the model simple, and to show the implications of the fishing effort allocation, the two fish populations modelled here are exactly the same apart from their spatial distribution difference (Fig.
+\ref{f:distribution}).
+As a result, if the landing obligation changes selectivity patterns in the short-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields.
+ERROR IN DECISION MAKING, NO DISCARING ALLOWED BY FISHERS, AND PERFECT COMPLIANCE TO THIS
 
-The forecasted effects of fisheries management in mixed fisheries result from individual choices by fishers on management and spatial heterogeneity in the resources rather than in differences in biological characestics of the resources.
-This emphsizes the need for incorporating detailed fishing dynamics.
-Histrocially, focus of management strategy evaluation has been primarily on managing the population (biological) dynamics of single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999}.
-Few MSE studies have explicitly incorporated the direct effects of human dimension harvesting on the stocks, although is equally important \cite{Fulton2011}.
-Human behaviour and decision-making are intrinsic parts of conservation and natural resource management.
-MSE with explicit human (social) behaviour dimension is increasingly being used \cite{Andersen2010, Ono2018} as fisheries science takes other sources of uncertainty to a greater extent to identify management actions that ensure the sustainability of fisheries.
-
-Our modeling MSE approach use a dynamic state variable model, including fishing location choices and heterogeneous populations, to evaluate the costs and benefits, both economic and in terms of risks to both stock and livelihoods, that resulted in response of adopting a major policy change such as the CFP \cite{CFP2013}.
-Starting from an overfished mixed fishery under open access conditions but with a landing obligation constraint, where discarding is not allowed, the response of individual effort allocation and the interannual dynamics for both fleet dynamics and quota allocation is counted when adopting MSY objectives while landing all catches.
-The analysis extended earlier approaches which used short-term state-dependent decision-making models (e.g.
-\cite{Alzorriz2018, Batsleer2016, Poos2010}) by including the consequent biomass of fish stocks and the management procedure to make its decision on how to respond to the state of the resource.
-The MSE framework, was thought of as a minimum realistic model \cite{Kell2007, Punt1995}, and included temporal and spatial key processes in the dynamics of two fish populations, the fisheries and a reevaluation of management targets.
-To keep the model simple, the two fish populations modelled here are exactly the same apart from their spatial distribution difference (Fig.
-\ref{f:distribution}).
-As a result, if the landing obligation changes selectivity patterns in the shor-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields.
-ERROR IN DECISION MAKING, NO DISCARING ALLOWED BY FISHERS, AND PERFECT COMPLIANCE TO THIS
+The forecasted effects of fisheries management in mixed fisheries result from spatial heterogeneity in resources and fishers adaptive behaviour.
+This emphasizes that human behaviour and decision-making are intrinsic parts of conservation and natural resource management and the need for incorporating detailed fishing dynamics in management strategy evaluation.
+Historically, the focus of management strategy evaluation has been primarily on population dynamics in single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999, Punt2016}.
+Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018}.
+Our MSE that links a dynamic state variable model for fisher behavior to a multi-species age structured model provides an alternative to those approaches.
 

### 2019-04-11 doc/MSE_MAR_18.tex
version ready to be checked by Adriaan

@@
 Meanwhile, when constraining both species, the production potential of both stocks was maintained, that is to say the yield curves were unchanged.
+The quota constraints were met by a reduction in fishing effort (Fig.
+\ref{f:effort}).
+as predicted in general terms by \cite{Ulrich2017}.
+This led to a reduction in fishing cost and an increase in net revenue over the mid-term because biomasses and yields increased.
 However, individual quota cobined with the stochastic fisheries catches resulted in lower than intended exploitation rates.
@@
 
+In the case study explored here, setting quotas based on MSY for all species resulted in mixed fisheries that were economically more viable in the long run compared to the unconstrained cased.
+It should be noted that this particular result may not hold true for mixed fisheries with unavoidable catches or bycatches of very low productive species.
+Fishing those species at their MSY levels would probably constrain effort to such an extent tat the net revenue will be reduced in the long-term \cite{Prellezo2016a,  Rindorf2017}.
+
 The model aimed to evaluate the biological and economic effects of adopting a major policy change such as the CFP \cite{CFP2013}.
@@
 As a result, if the landing obligation changes selectivity patterns in the short-term as predicted in \cite{Alzorriz2018, Batsleer2016}, the present MSE modeling approach allows the evaluation on the reference points for sustainable exploitation in longer terms by explicitly modelling how changes in exploitation rates impact fish populations and yields.
-ERROR IN DECISION MAKING, NO DISCARING ALLOWED BY FISHERS, AND PERFECT COMPLIANCE TO THIS
 
@@
 Historically, the focus of management strategy evaluation has been primarily on population dynamics in single and multispecies fisheries \cite{Butterworth2007, Dichmont2006b, Punt1999, Punt2009, Smith1999, Punt2016}.
-Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018}.
+Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012, Link2017}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018, Simons2015a}.
 Our MSE that links a dynamic state variable model for fisher behavior to a multi-species age structured model provides an alternative to those approaches.
 
-When the studied mixed fishery was unmanaged, with unconstrained quota for both species (pre-manage period) and no discarding allowance, the observed harvesting rates showed average values larger than the estimated average values that can be harvested in a sustainable manner ($H_{max}$).
-Overfishing occurs in both stocks when fishers exceed $H_{max}$, whereas rebuilding requires reducing exploitation below $H_{max}$ (Fig.
-\ref{f:yields}).
-Therefore, during the pre-manage period, a high exploitation rate caused a decline in total biomass and fishers mainly harvest the individuals that would normally be added to the population, with younger fish being targeted for.
-Our results suggested that given the spatial and price species distribution parameters (Fig.
-\ref{f:distribution} and \ref{f:prices}), fishing in both the northern and southern fishing grounds were equally profitable; thus, the distribution of fishing effort was equally distributed (Fig.
-\ref{f:effort}).
+The complexity of adding fishers behaviour in a minimum realistic model of mixed fisheries in an MSE framework requires making several simplifying assumptions.
+The details and possible consequences of these assumptions have been discussed in eg.
+\cite{Alzorriz2018}, and include full compliance of fishers with the discarding policy, absence of a quota market, monthly choices on fishing, and a lack of interference competition among fishers.
+Also, the utility function only comprises of decisions based on annual net revenues, ommitting any social dynamics or tradition.
+These social dynamics and traditions are potentially important \cite{Girardin2017, Schluter2012}, and could be incorporated in future applications.
+Important assumptions about the dynamics of the population dynamics included negligible natural mortality, constant recruitment, constant growth, and constant migration.
 
-Moving from an unmanaged period to a situation where one or both species were quota regulated by implementing MSY objectives while maintaining a LO management regime (i.e.
-main CFP objectives: exploiting at MSY and no discarding allowance) led to a very restrictive fishing activity (Fig.
-\ref{f:effort}).
-Fishing effort was constrained by one or two quota stocks, as predicted in general terms by \cite{Ulrich2017}, resulting in drastic reductions of effort over the short-term and mid-term.
-In addition to reconciling fishery and conservation objectives (such as the CFP), setting exploitation rates below $H_{max}$ resulted in a reduction in fishing cost and an increase in net revenue over the mid-term (Fig.
-\ref{f:effort}).
-However, as \cite{Prellezo2016a} concluded, since effort was drastically reduced to achieve the objectives of the CFP; the increase in net revenue gains were only experienced by some vessels.
-Consequently, a significant part of the fleet had to cease its activity (Fig.
-\ref{f:effort}).
-
-Nevertheless, the framework used made several assumptions: that there was perfect knowledge of the stochastic nature of catch rates, full compliance of fishers with the discarding policy, lack of transferability of individual quotas, and lack of interference competition among fishers \cite{Alzorriz2018}.
-IMPORTANT< UTILITY FUNCTION IS ANNUAL NET REV, NO SOCIAL DYNAMICS.
-Regarding fish populations, also important assumptions were established: negligible natural mortality, constant recruitment, constant growth, constant migration, and constant size-dependent prices.
-Individual choices on fishing behaviour were assumed to be made once a month.
-However, despite these assumptions, the minimum realistic model developed in this study allowed us to simulate the dynamics of a mixed fishery and examine the effects of single species quota management.
-
-In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of any discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps.
-Allowing the latter could potentially remedy the underutilisation of quota in the two quota situation, given that this reuslts from the stochsticity in catches, which are uncorrelated.
-
- We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY.
-However, applying any kind of exemption or flexibility beyond the MSY would be negative \cite{Prellezo2016}.
-Consequently, a critical feature of management bodies is to ensure that the use of different flexibilities will constrain catches to comply with the MSY objective in Article 2(2) of the CFP Basic Regulation.
+In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps.
+We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY \cite{Prellezo2016}.
 Achieving a balance between quota flexibility, over-exploitation risk, and administrative simplicity is critical for the profitability and sustainability of multi-species fisheries \cite{Alzorriz2018}.
@@
 However, if detection rates are low and the short-term benefits of non-compliance are high (as our results suggest) then non-compliance is likely to be a substantial problem in the absence of measures promoting voluntary compliance.
-IF YOU WANT TO STUDY THIS, THEN DETAILED PARAMETRIZATION IS NEEDED THAT INCLUDES CASE DEPENDENT LH CHARATRISTICS AND MIGRATIONS
- %Optimal effort at MSY generates effort reductions of 37\% to 62\% over the mid-term, thus there would be some vessels that could benefit up to 1.7 times larger profits than before the CFP implementation.
-The results highlighted the importance of accounting for technical interactions and their temporal dynamics in both quota allocation and fleet dynamics to design a realistic multispecies fishery management strategy analysis
-
-Finally, a dynamic state variable model was used as a tool to create a quick quota allocation and fleet dynamics model.
-Many detailed fleet dynamics models have been developed over the past decades \cite{Andersen2010, Ono2018, Simons2015a} and they could have been used to include spatial factors as another dimension in the MSE.
-However, following \cite{Dowling2011}, incorporating uncertainty in errors in human dynamic decision-making help us to understant and predict changes in human behaviour in response of adopting a major policy change such as the CFP \cite{CFP2013}.
-The effect of introducing a realistic multispecies fishery within this MSE framework still remains the next step to be taken.
-Modelling and predicting realistic human behavior with its associated uncertainty in an ever-changing socioecological system will remain a major challenge to address in the future \cite{Ono2018}.
 
@@
 
-%Although this may seem implausible too, a drastic reduction of effort would increase profits if MSY and LO were contemplated.
-However, no flexibilities or year transfer swaps were consider.
-Be carefull if applying flexibilities, because as Prellezo said, any increased in mortality beyond the MSY would be negative...
-%prellezo 2016, In the mid-term and without any consideration made in terms of the ecosystem functioning as a whole, the results obtained from applying any kind of exemption or flexibility are, simply, negative.
-Fishing beyond FMSY (even in one year) implies that there will be a penalty in the future.
-This penalty will come in the form of lower biomasses, that has the mixed effect of increasing the cost of fishing the same level of catches and reducing the total catch due to the lower abundances and the subsequent lower TACs.
-
-%The largest yield (or catch) that can be taken from a species' stock over an indefinite period.
-Fundamental to the notion of sustainable harvest, the concept of MSY aims to maintain the population size at the point of maximum growth rate by harvesting the individuals that would normally be added to the population, allowing the population to continue to be productive indefinitely.
-Under the assumption of logistic growth, resource limitation does not constrain individuals’ reproductive rates when populations are small, but because there are few individuals, the overall yield is small.
-At intermediate population densities, also represented by half the carrying capacity, individuals are able to breed to their maximum rate.
-At this point, called the maximum sustainable yield, there is a surplus of individuals that can be harvested because growth of the population is at its maximum point due to the large number of reproducing individuals.
-Above this point, density dependent factors increasingly limit breeding until the population reaches carrying capacity.
-At this point, there are no surplus individuals to be harvested and yield drops to zero.
-The maximum sustainable yield is usually higher than the optimum sustainable yield and maximum economic yield
-
-%There is an extensive literature on fisheries economic theory in which the Gordon-Schaefer equilibrium production model is central (Gordon 1954, Schaefer 1957, Clark 1983).
-This theory holds that there is an economic TRP, the Maximum Economic Yield (MEY), which occurs at the effort level yielding the greatest margin of revenue over cost from the resource.
-%For a linear cost curve, this inevitably occurs to the left of MSY on the fishing effort axis.
-Since Fmey occurs at lower levels of effort than Fmsy, the use of this economic Target Reference Point is less likely to result in biological overfishing than the use of Fmsy.
-
-%As a TRP, Fmey is responsive to any changes in the economic environment which affect either the value of fish, or the cost of fishing.
-It may also be dependent on changes in fish abundance, if market price increases with declining abundance and is independent of the availability of similar resources elsewhere.
-Subsidies or external economic considerations such as fuel taxes will also affect the location of an economic Reference Point (e.g.
-Panayotou 1988).
-
-%The effect of supply on fish prices may, under certain circumstances, result in higher total profit, or profit per unit catch, when total catch is reduced.
-This characteristic may be a consideration in setting target fishing levels or catches but is least likely to be effective in situations where fish prices are set by global markets, e.g.
-the tuna fishery for the canning industry.
-
-%The value of a unit weight of the landed catch may vary with the size of individual fish, and in multispecies fisheries with species composition.
-Both fish size and species composition are functions of fishing mortality, and based on purely economic criteria, may be used as target reference points.
-Even if the actual target F cannot be estimated, in theory, F could be adjusted in increments until the desirable target catch characteristics are achieved.
-
-%In considering TRPs based on economic criteria, it is important to be aware of the effect which the practice of discounting could have on reference points.
-In evaluating investment projects, including resource management, economists discount the future value of any commodity.
-Discount rates may be in the order of 10\%.
-In the case of a fishery where the population growth rate does not exceed the discount rate, then a strict application of economic theory would suggest that in the absence of other considerations (such as an economic value placed on recreational use of resources) the whole stock should be harvested now, and the proceeds of their sale invested.
-Long-lived species with slow growth rates, such as whales, clearly fall into this category.
-The blatant contradiction between this common economic approach, and the concept of sustainability, constitutes an unresolved paradox (Hilborn and Walters 1992).
-
 \begin{notes}[Acknowledgements]

### 2019-04-11 doc/MSE_MAR_18.tex
fixed a few typos

@@
 At the population levels of the two species this resulted in overal cacthes that were lower than the TACs.
-Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere \cite{Kempf2016, Rindorf2017, Hilborn2015}, but is generall attributed to differences in life-histories among fished species.
+Such lack of achieving MSY in mixed fisheries with multiple single-species quota has been described elsewhere \cite{Kempf2016, Rindorf2017, Hilborn2015}, but is generally attributed to differences in life-histories among fished species.
 In this study, the discrepancy between the intended annual quotas at MSY and realized catches is caused only by schochasticity in the system.
@@
 It is caused by the complex migration biology and the effects of environmental factors on fish behaviour.
-A well functioning quota market would allow vessels to swap excess quota for constraining and thus bring realized catches closer to quotas.
+A well-functioning quota market would allow vessels to swap excess quota for constraining and thus bring realized catches closer to quotas.
 
@@
 It should be noted that this particular result may not hold true for mixed fisheries with unavoidable catches or bycatches of very low productive species.
-Fishing those species at their MSY levels would probably constrain effort to such an extent tat the net revenue will be reduced in the long-term \cite{Prellezo2016a,  Rindorf2017}.
+Fishing those species at their MSY levels would probably constrain effort to such an extent that the net revenue will be reduced in the long-term \cite{Prellezo2016a,  Rindorf2017}.
 
 The model aimed to evaluate the biological and economic effects of adopting a major policy change such as the CFP \cite{CFP2013}.
-Starting from an overfished mixed fishery under open access conditions but with a landing obligation, the response of individual effort allocation and the interannual population dynamics are explicitly modeled, while adopting MSY objectives.
+Starting from an overfished mixed fishery under open access conditions but with a landing obligation, the response of individual effort allocation and the interannual population dynamics are explicitly modelled, while adopting MSY objectives.
 The analysis extended earlier approaches that used short-term state-dependent decision-making models (e.g.
@@
 Explicitly incorporating direct effects of human behaviour of those who harvest stocks is equally important \cite{Fulton2011,Milner-Gulland2012, Link2017}, but relatively few studies have attempted this \cite{Andersen2010, Ono2018, Simons2015a}.
-Our MSE that links a dynamic state variable model for fisher behavior to a multi-species age structured model provides an alternative to those approaches.
+Our MSE that links a dynamic state variable model for fisher behaviour to a multi-species age structured model provides an alternative to those approaches.
 
@@
 \cite{Alzorriz2018}, and include full compliance of fishers with the discarding policy, absence of a quota market, monthly choices on fishing, and a lack of interference competition among fishers.
-Also, the utility function only comprises of decisions based on annual net revenues, ommitting any social dynamics or tradition.
+Also, the utility function only comprises of decisions based on annual net revenues, omitting any social dynamics or tradition.
 These social dynamics and traditions are potentially important \cite{Girardin2017, Schluter2012}, and could be incorporated in future applications.
@@
 
-In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annaul quota swaps.
+In the study, the fishes were not allowed to discard, complying to a landings obligation.We did not include the use of discard flexibilities of the LO (CFP Article 15, \cite{CFP2013}) or annual quota swaps.
 We may thus have underestimated the flexibility of the fishery to respond to quota limitations imposed under both the landing obligation and exploiting at MSY \cite{Prellezo2016}.
